---
title: "Bioassay Analysis Report"
output:
  html_document:
    df_print: paged
    self_contained: true
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: flatly
    css: report_style.css
  word_document:
    toc: true
    fig_caption: true
    fig_width: 6
    fig_height: 4
  pdf_document:
    toc: true
header-includes:
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyfoot[C]{\thepage}
  - \fancyhead{}
  - \renewcommand{\headrulewidth}{0pt}
params:
  output_dir: NULL
  lang: "en"
  compact: false
  stats_env: !r new.env(parent = emptyenv())
---

<!--
================================================================================
Chunk pipeline (in execution order):

  setup                           - libraries, output_dir, i18n, source helpers
  load-data                       - read CSV, validate, compute %B/B0 for ELISA
  executive-summary               - top-of-report status box
  standards-table                 - display standard concentrations
  methods-section                 - methods paragraph with citations
  model-fitting                   - fit all selected weightings (LL.4 -> LL.3 ->
                                    interpolation fallback); defines:
                                    all_models, model_fit, R2, RMSE,
                                    conc_range, model_fits, ec20, ec80,
                                    classify_range(), flag_range()
  exec-summary-box                - rendered exec summary (needs model_fit)
  weight-comparison               - overlay plot of weightings (HTML)
  weighting-suitability           - Brown-Forsythe / variance ratio
  model-stability-assessment      - convergence / boundary diagnostics
  lloq-uloq                       - back-calculation accuracy -> formal LLOQ/ULOQ
  dose-response-plots             - main DRC plot
  model-stats                     - 4PL coefficient table
  save-model-stats                - write model_stats.json (partial)
  standard-backcalculation        - per-standard recovery table
  data-quality-overview           - quick data quality panel
  sample-analysis                 - MAIN SAMPLE QUANTIFICATION (defines:
                                    sample_results, replicate_stats,
                                    outlier_flags, replicate_summary,
                                    tissue-normalized columns if ELISA)
  sample-results-table            - rendered sample table + CSV writes
  outlier-results                 - outlier detection report block
  ci-method-note                  - CI method footnote
  save-sample-stats               - updates model_stats.json with mean_sample_cv
  summary-results-table           - detailed replicate summary
  sample-boxplot                  - variability visualization
  plate-heatmap                   - plate heatmap
  plate-positional-qc             - row/column bias check
  drc-with-samples                - DRC plot with sample points overlaid
  parallelism-assessment          - optional multi-curve comparison
  report-summary                  - high-level summary bullets
  report-interpretation           - QC-profile-based interpretation
  tissue-normalization-traceability - ELISA + tissue only
  exclusion-audit                 - formal exclusion audit table

Key cross-chunk dependencies:
  - model-fitting defines variables used by almost all later chunks
  - sample-analysis must run before sample-results-table, outlier-results,
    save-sample-stats, summary-results-table, sample-boxplot
  - tissue_weights (ELISA) is loaded inside sample-analysis
================================================================================
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# For DOCX/PDF output, force static PNG figures at high resolution
if (!knitr::is_html_output()) {
  knitr::opts_chunk$set(
    dev = "png",
    dpi = 300,
    fig.width = 6,
    fig.height = 4,
    fig.retina = 1
  )
}

# Load required libraries.
# kableExtra guard: install if missing so the Rmd is self-healing in
# environments where global.R's install step was skipped or failed.
if (!requireNamespace("kableExtra", quietly = TRUE)) {
  .rmd_repos <- c("https://packagemanager.posit.co/cran/latest",
                  "https://cloud.r-project.org")
  install.packages("kableExtra", repos = .rmd_repos, quiet = TRUE)
  if (!requireNamespace("kableExtra", quietly = TRUE)) {
    if (!requireNamespace("remotes", quietly = TRUE))
      install.packages("remotes", repos = .rmd_repos, quiet = TRUE)
    remotes::install_github("haozhu233/kableExtra")
  }
  rm(.rmd_repos)
}
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(drc)
library(scales)
library(knitr)
library(kableExtra)
library(stringr)
library(ggtext)
library(glue)
library(htmltools)
library(tibble)
library(jsonlite)

# %||% operator is defined in global.R (loaded before rendering)

# H4: compact-mode flag for chunk gating below.
is_compact <- isTRUE(params$compact)

# Validate output directory
output_dir <- normalizePath(params$output_dir, winslash = "/", mustWork = TRUE)

# Load language setting
lang <- params$lang %||% "en"

# Get script directory FIRST (needed before sourcing any files)
script_dir <- tryCatch({
  rmd_path <- knitr::current_input()
  if (!is.null(rmd_path) && rmd_path != "" && rmd_path != ".") {
    normalizePath(dirname(rmd_path), winslash = "/")
  } else {
    if (dir.exists("reports")) {
      normalizePath("reports", winslash = "/")
    } else {
      possible_reports_dir <- file.path(dirname(params$output_dir), "reports")
      if (dir.exists(possible_reports_dir)) {
        normalizePath(possible_reports_dir, winslash = "/")
      } else {
        "."
      }
    }
  }
}, error = function(e) {
  if (dir.exists("reports")) {
    return(normalizePath("reports", winslash = "/"))
  } else {
    "."
  }
})

# Source i18n translations (search multiple locations)
i18n_found <- FALSE
for (i18n_candidate in c(
  file.path(script_dir, "i18n.R"),
  file.path(dirname(script_dir), "i18n.R"),
  file.path(dirname(script_dir), "server", "i18n.R"),
  file.path(getwd(), "server", "i18n.R"),
  "../i18n.R",
  "../server/i18n.R",
  "i18n.R",
  file.path(dirname(output_dir), "i18n.R"),
  file.path(dirname(output_dir), "server", "i18n.R")
)) {
  if (file.exists(i18n_candidate)) {
    source(i18n_candidate)
    i18n_found <- TRUE
    break
  }
}
if (!i18n_found) {
  tr <- function(key, lang = "en", ...) paste0("[", key, "]")
}

# Check if all required function files exist
# M1.7: analysis_pipeline.R and report_sections.R are scaffolded empty now;
# subsequent M1.1-M1.6 extractions will populate them.
required_files <- c("report_constants.R", "report_functions.R",
                    "plot_functions.R", "analysis_pipeline.R",
                    "report_sections.R")
missing_files <- c()

for (file in required_files) {
  file_path <- file.path(script_dir, file)
  if (!file.exists(file_path)) {
    missing_files <- c(missing_files, file_path)
  }
}

if (length(missing_files) > 0) {
  stop("Missing required function files:\n", paste(missing_files, collapse = "\n"),
       "\n\nExpected in directory: ", script_dir,
       "\nActual files found: ", paste(list.files(script_dir, pattern = "\\.R$"), collapse = ", "),
       "\n\nMake sure all R files are in the 'reports/' folder.")
}

# Source function files
source(file.path(script_dir, "report_constants.R"))
source(file.path(script_dir, "report_functions.R"))
source(file.path(script_dir, "plot_functions.R"))
source(file.path(script_dir, "analysis_pipeline.R"))
source(file.path(script_dir, "report_sections.R"))

# Load assay configuration
assay_config <- tryCatch({
  load_assay_config(output_dir)
}, error = function(e) {
  # Fallback to RBA defaults
  list(
    assay_type = "rba",
    detection_method = "radioligand",
    toxin_standard_label = "Saxitoxin"
  )
})

# Load analysis settings
analysis_config <- tryCatch({
  cfg_file <- file.path(output_dir, "analysis_config.json")
  if (file.exists(cfg_file)) {
    jsonlite::fromJSON(cfg_file)
  } else {
    list(regression_weight = "none", quant_range_min = 20, quant_range_max = 80,
         ci_method = "t_dist", enable_outlier_detection = FALSE, outlier_min_n = 3,
         normality_assumption = "assume")
  }
}, error = function(e) {
  list(regression_weight = "none", quant_range_min = 20, quant_range_max = 80,
       ci_method = "t_dist", enable_outlier_detection = FALSE, outlier_min_n = 3,
       normality_assumption = "assume")
})
# Ensure backward compatibility for configs saved without normality_assumption
if (is.null(analysis_config$normality_assumption)) {
  analysis_config$normality_assumption <- "assume"
}
# Resolve configurable CV threshold (app-level override or fallback to global constant)
cv_limit <- analysis_config$cv_limit %||% QC_THRESHOLDS$cv_limit

# Helper: format numbers with scientific notation for extreme values
fmt_smart <- function(x, digits = 4) {
  if (is.na(x) || !is.finite(x)) return("NA")
  if (abs(x) >= 1e5 || (abs(x) > 0 & abs(x) < 1e-4)) {
    formatC(x, format = "e", digits = digits)
  } else {
    formatC(x, format = "f", digits = digits)
  }
}
```

```{css details-style, echo=FALSE, eval=knitr::is_html_output()}
/* GUI refresh §6.1 / §6.7: report-only typography overrides. The Plex
   font stack is loaded by app.R for the live Shiny app; the report
   ships as a self-contained HTML so we re-import the same fonts here
   so emailed / printed reports also pick up the design. */
@import url('https://fonts.googleapis.com/css2?family=IBM+Plex+Mono:wght@400;500;600&family=IBM+Plex+Sans:wght@400;500;600&family=IBM+Plex+Serif:wght@400;500&display=swap');

body {
  font-family: "IBM Plex Sans", -apple-system, BlinkMacSystemFont,
               "Segoe UI", sans-serif;
  color: #0f1722;
  font-size: 13px;
  line-height: 1.6;
}
.title { /* rmarkdown's auto title becomes the serif display heading */
  font-family: "IBM Plex Serif", Georgia, serif;
  font-weight: 500;
  font-size: 32px;
  letter-spacing: -0.01em;
  color: #0f1722;
  margin-bottom: 4px;
}
h1, h2, h3, h4 {
  font-family: "IBM Plex Sans", sans-serif;
  font-weight: 600;
  color: #0f1722;
  margin-top: 24px;
}
h1 { font-size: 22px; font-weight: 500; }
h2 { font-size: 17px; }
h3 { font-size: 14px; }
h4 { font-size: 13px; }
.mono, code, kbd, pre, tt {
  font-family: "IBM Plex Mono", ui-monospace, "SF Mono", monospace;
}

/* Section <details> chevron + sentence-case heading (§6.7) */
details > summary {
  cursor: pointer;
  font-weight: 500;
  color: #2a3340;
  margin-bottom: 8px;
}
details > summary:hover { color: #1f3a5f; }
details {
  margin-bottom: 12px;
}

/* GUI refresh §6.6: kable / kableExtra table styling. Mono numerics,
   tabular numerals, small-caps mono header, surface-2 hover. */
table.kable, table.dataframe, .kable_wrapper table {
  border-collapse: collapse;
  width: 100%;
  font-size: 12px;
}
table.kable thead th, table.dataframe thead th, .kable_wrapper thead th {
  background: #f1f3f5;
  color: #5b6573;
  font-family: "IBM Plex Mono", monospace;
  font-size: 11px;
  font-weight: 500;
  letter-spacing: 0.04em;
  text-transform: uppercase;
  padding: 6px 10px;
  border-bottom: 1px solid #e3e6ea;
  text-align: left;
}
table.kable td, table.dataframe td, .kable_wrapper td {
  padding: 6px 10px;
  border-bottom: 1px solid #e3e6ea;
  vertical-align: top;
}
/* Numeric-looking cells (mono + tabular) get right-aligned. heuristic:
   any <td> where rmarkdown's pandoc CSS class has been applied. */
table.kable td.r, table.dataframe td.r {
  font-family: "IBM Plex Mono", monospace;
  font-variant-numeric: tabular-nums;
  text-align: right;
}
table.kable tbody tr:hover, table.dataframe tbody tr:hover {
  background: #f1f3f5;
}

/* GUI refresh §6.3: executive summary card — quiet 1 px line, no
   tinted gradient (the .Rmd helper renders inline styling for
   email-safe fallback, this overrides for live HTML). */
.exec-summary-card, div[style*="linear-gradient(135deg, #E8F5E9"] {
  background: #ffffff !important;
  border: 1px solid #e3e6ea !important;
  border-left: 4px solid #009e73 !important;
  box-shadow: none !important;
}
```

```{r load-data, include=FALSE}
# Load and validate data
data_long_raw <- read.csv(file.path(output_dir, "long_data_output.csv"), stringsAsFactors = FALSE)

# Remove diagnostic control columns that confuse users
diagnostic_cols <- c("ControlValues_B0", "ControlValues_NSB", "ControlValues_Blank")
data_long <- data_long_raw %>%
  dplyr::select(-any_of(diagnostic_cols))

validate_input_data(data_long)

# Detect assay type from data if not specified in config
has_normalized <- "NormalizedValue" %in% names(data_long)
is_elisa <- isTRUE(assay_config$assay_type == "elisa") || has_normalized

# For ELISA, calculate proper B/B0 values following Cayman protocol
if (isTRUE(is_elisa)) {
  # Calculate B/B0 values using control wells (Blank, NSB, B0)
  # Use the raw data with diagnostic columns for calculation, but return clean data
  data_long <- tryCatch({
    data_with_bb0 <- calculate_elisa_bb0(data_long_raw, use_percent = TRUE)  # Use %B/B0 (0-100%)
    # Remove diagnostic columns from the result
    data_with_bb0 %>% dplyr::select(-any_of(diagnostic_cols))
  }, error = function(e) {
    warning("Could not calculate B/B0 values: ", e$message)
    warning("Falling back to existing NormalizedValue column")
    data_long$calculated_bb0 <- data_long$NormalizedValue
    data_long
  })
  
  response_var <- "calculated_bb0"
  
  # Update assay configuration
  if (!isTRUE(assay_config$assay_type == "elisa")) {
    assay_config$assay_type <- "elisa"
    assay_config$units <- "pg/mL"
    assay_config$analyte <- assay_config$analyte %||% "Analyte"
  }
  
  # Display control well summary for QC
  control_summary <- attr(data_long, "control_summary")
  if (!is.null(control_summary)) {
    cat("**Control Well QC:**\n")
    cat("- Blank average:", round(as.numeric(control_summary$blank_avg), 3), "\n")
    cat("- NSB average:", round(as.numeric(control_summary$nsb_avg), 3), "\n")
    cat("- B0 average:", round(as.numeric(control_summary$b0_avg), 3), "\n")
    cat("- Hierarchy valid:", control_summary$hierarchy_valid, "\n\n")
  }
  
} else {
  # RBA: Use existing response variable
  response_var <- if (has_normalized && "NormalizedValue" %in% names(data_long)) {
    "NormalizedValue"
  } else {
    "MeasurementValue"  
  }
}

# Get axis labels
labels <- get_axis_labels(assay_config, lang = lang)

# Load notes (optional)
notes_file <- file.path(output_dir, "notes.json")
notes_text <- "No notes added."
if (file.exists(notes_file)) {
  notes_obj <- jsonlite::fromJSON(notes_file, simplifyVector = FALSE)
  notes_text <- notes_obj$notes %||% "No notes added."
}

# Load sample processing configuration for ELISA
if (isTRUE(is_elisa)) {
  processing_config <- load_sample_processing_config(output_dir)
}
```


# `r tr(if (isTRUE(is_elisa)) "report_elisa" else "report_rba", lang)` `r tr("analysis_report", lang)`

```{r report-meta, results='asis'}
# Item 12: replace the former <h4 class="author"> + inline bold lines with a
# muted two-column metadata header.  HTML gets the styled .bs-report-meta
# block; DOCX/PDF falls back to compact bold lines.
.drc_ver_meta <- tryCatch(as.character(packageVersion("drc")), error = function(e) "?")
.assay_lbl    <- paste0(toupper(assay_config$assay_type %||% ""),
                        " — ",
                        assay_config$analyte %||%
                          assay_config$toxin_standard_label %||% "")
.app_ver_meta <- REPORT_INFO$version %||% "?"

if (is_html_out()) {
  cat(paste0(
    '<header class="bs-report-meta">\n',
    '<div class="meta-row">',
      '<span class="label">Authors</span>',
      '<span class="value">Kristof Moeller (IAEA, Monaco)',
        ' · Arnold Molina Porras (Universidad de Costa Rica)</span>',
    '</div>\n',
    '<div class="meta-row">',
      '<span class="label">Generated</span>',
      sprintf('<span class="value">%s · %s</span>',
              format(Sys.Date(), "%Y-%m-%d"), Sys.info()[["user"]]),
    '</div>\n',
    '<div class="meta-row">',
      '<span class="label">Assay</span>',
      sprintf('<span class="value">%s</span>', .assay_lbl),
    '</div>\n',
    '<div class="meta-row">',
      '<span class="label">Software</span>',
      sprintf('<span class="value">Bioassay Analysis App v%s · %s · drc v%s</span>',
              .app_ver_meta, R.version.string, .drc_ver_meta),
    '</div>\n',
    '</header>\n\n'
  ))
} else {
  cat(paste0(
    "**Authors:** Kristof Moeller (IAEA, Monaco) and Arnold Molina Porras (Universidad de Costa Rica)  \n",
    sprintf("**Generated:** %s · %s  \n", format(Sys.Date(), "%Y-%m-%d"), Sys.info()[["user"]]),
    sprintf("**Assay:** %s  \n", .assay_lbl),
    sprintf("**Software:** Bioassay Analysis App v%s · %s · drc v%s  \n\n",
            .app_ver_meta, R.version.string, .drc_ver_meta)
  ))
}
```

```{r dynamic-description, results='asis'}
# Dynamic description based on assay type
if (isTRUE(is_elisa)) {
  cat(tr("elisa_intro", lang, assay_config$analyte %||% "analyte"), "\n\n")
  cat("**", tr("elisa_method", lang), "**\n\n", sep = "")
} else {
  cat(tr("rba_intro", lang, assay_config$toxin_standard_label %||% "biotoxin"), "\n\n")
  cat("**", tr("rba_method", lang, assay_config$detection_method %||% "radioligand"), "**\n\n", sep = "")
}
```

## `r tr("analysis_notes", lang)`

```{r notes-section, results='asis'}
if (nzchar(trimws(notes_text)) && notes_text != "No notes added.") {
  cat(notes_text)
} else {
  cat("*", tr("no_notes", lang), "*", sep = "")
}
```

## `r tr("std_curve_config", lang)`

`r section_open(tr("std_concentrations", lang))`

```{r standards-table}
# Extract and display standard concentrations
standards_data <- data_long %>%
  filter(SampleType == "Standard", !is.na(StandardConc)) %>%
  arrange(desc(StandardConc)) %>%
  distinct(StandardConc)

# Format concentrations based on assay type
if (isTRUE(is_elisa)) {
  # ELISA: Use regular number formatting (no scientific notation)
  standard_table <- data.frame(
    StandardNumber = seq_len(nrow(standards_data)),
    Concentration = format(standards_data$StandardConc, digits = 4, nsmall = 1, big.mark = ",")
  )
  unit_label <- paste0("Concentration (", assay_config$units %||% "pg/mL", ")")
} else {
  # RBA: Use scientific notation  
  standard_table <- data.frame(
    StandardNumber = seq_len(nrow(standards_data)),
    Concentration = format(standards_data$StandardConc, scientific = TRUE, digits = 3)
  )
  unit_label <- "Concentration (mol/L)"
}

render_table(
  standard_table, 
  caption = tr("std_concentrations_table", lang),
  col_names = c(tr("col_standards", lang), unit_label)
)
```

`r section_close()`

## `r tr("methods_title", lang)`

```{r methods-section, results='asis'}
# Item 10: expanded Methods section covering model, weighting, LLOQ/ULOQ
# criteria, CI method, outlier rule, and software. Replaces the former single
# sentence. render_methods_section() lives in reports/report_sections.R.
render_methods_section(
  analysis_config = analysis_config,
  assay_config    = assay_config,
  is_elisa        = is_elisa,
  lang            = lang
)
```

## `r tr("drc_analysis", lang)`

```{r model-fitting, include=FALSE}
# M1.2: fit_all_models() in reports/analysis_pipeline.R handles the
# modelling work (high-variability filtering, weight setup, H2 auto pick,
# LL.4 -> LL.3 -> interpolation fallback chain, EC20/EC80, prediction grid).
# The two closure helpers below stay in the chunk so they capture the
# chunk-local environment (where bb0_lower / bb0_upper / ec20 / ec80 now
# live after list2env) rather than the function's local scope, preserving
# the legacy lookup behaviour for later chunks that call them.
#
# H4 cache: render_reports() in report_pipeline.R renders the same template
# up to four times per run (compact + full × html + docx). All variants share
# the same fit, so we serialise the first run's result to output_dir and
# short-circuit subsequent renders. Pipeline clears the file at run start.
.fit_cache_path <- file.path(output_dir, ".fit_all_models_cache.rds")
if (file.exists(.fit_cache_path)) {
  .fit_res <- readRDS(.fit_cache_path)
} else {
  .fit_res <- fit_all_models(
    data_long       = data_long,
    response_var    = response_var,
    analysis_config = analysis_config,
    is_elisa        = is_elisa,
    lang            = lang,
    cv_limit        = cv_limit,
    STATS_CONFIG    = STATS_CONFIG
  )
  saveRDS(.fit_res, .fit_cache_path)
}
list2env(.fit_res, envir = environment())

# classify_range(): uses different reference axes depending on assay type.
#
# ELISA: classifies by %B/B0 response value. The kit manufacturer
#   (e.g., Cayman) defines the reliable quantification range in terms of
#   %B/B0 bounds (default 20-80%). The response axis is bounded [0, 100],
#   so %B/B0 is the natural reference.
#
# RBA: classifies by estimated concentration vs EC20/EC80 concentrations,
#   following the AOAC SMPR recommendation for radioligand binding assays.
#   The response axis (CPM) is unbounded, so concentration is the natural
#   reference.
#
# Both approaches identify samples falling outside the linear portion of the
# dose-response curve; they just use different axes to do so.
classify_range <- function(response_val, conc_val, is_elisa_assay) {
  if (isTRUE(is_elisa_assay)) {
    if (is.na(response_val)) return("Unknown")
    if (response_val >= bb0_lower && response_val <= bb0_upper) return("Within Range")
    return("Out of Range")
  } else {
    if (is.na(conc_val) || is.na(ec20) || is.na(ec80)) return("Unknown")
    if (conc_val >= ec20 && conc_val <= ec80) return("Within Range")
    return("Out of Range")
  }
}

# Helper function: get LLOQ/ULOQ flag for a sample
flag_range <- function(response_val, conc_val, is_elisa_assay) {
  if (isTRUE(is_elisa_assay)) {
    if (is.na(response_val)) return("")
    if (response_val > bb0_upper) return("<LLOQ")
    if (response_val < bb0_lower) return(">ULOQ")
    return("")
  } else {
    if (is.na(conc_val) || is.na(ec20) || is.na(ec80)) return("")
    if (conc_val < ec20) return("<LLOQ")
    if (conc_val > ec80) return(">ULOQ")
    return("")
  }
}

# GUI refresh §6.2 / Q4(e): pre-compute LLOQ/ULOQ from
# determine_lloq_uloq() right after the fit so the KPI strip below
# can reference formal_lloq / formal_uloq. The lloq-uloq chunk
# further down still owns the rendering of the back-calculation
# table; it now reads .lloq_res from this chunk's environment
# instead of recomputing.
.lloq_res <- if (!drc_failed_completely && !is.null(model_fit)) {
  tryCatch(
    determine_lloq_uloq(
      standards_for_model = standards_for_model,
      model_fit           = model_fit,
      response_var        = response_var,
      is_elisa            = is_elisa
    ),
    error = function(e) NULL
  )
} else {
  NULL
}
formal_lloq <- if (!is.null(.lloq_res)) .lloq_res$formal_lloq else NA_real_
formal_uloq <- if (!is.null(.lloq_res)) .lloq_res$formal_uloq else NA_real_
```

```{r report-header, results='asis'}
# Condensed metadata block (left) + KPI strip (right) in a single flex row.
# Replaces the former stacked pre-fit executive summary box and post-fit
# green-gradient overall-status box.
render_report_header(
  R2              = R2,
  RMSE            = RMSE,
  formal_lloq     = formal_lloq,
  formal_uloq     = formal_uloq,
  assay_config    = assay_config,
  is_elisa        = is_elisa,
  data_long       = data_long,
  lang            = lang,
  control_summary = if (exists("control_summary")) control_summary else NULL
)
```

```{r exec-summary-accuracy, results='asis'}
# M1.6: one-line "Mean recovery X% ± Y% across N levels" appended after
# the status box. render_overall_accuracy() returns invisibly when the
# standard-recovery summary was not produced (e.g. compact mode).
render_overall_accuracy(
  recovery_summary = if (exists("recovery_summary")) recovery_summary else NULL
)
```

```{r variability-info, results='asis'}
# M1.6: bullet list of standards excluded for high CV.
render_high_variability_info(high_var_standards = high_var_standards,
                             is_elisa           = is_elisa,
                             assay_config       = assay_config,
                             lang               = lang)
```

```{r qc-traffic-light, results='asis'}
# M1.6: traffic-light QC summary card (R^2 / Hill / max CV / mean recovery).
# Compute + presentation are inseparable here so the helper does both;
# wrap whole block in a tryCatch that prints a one-line "could not be
# generated" message instead of crashing the report.
render_qc_traffic_light_section(model_fit          = model_fit,
                                standards_for_model = standards_for_model,
                                response_var       = response_var,
                                replicate_stats    = if (exists("replicate_stats")) replicate_stats else NULL,
                                R2                 = R2,
                                weight_desc        = weight_desc,
                                assay_config       = assay_config,
                                lang               = lang)
```

```{r weight-comparison, results='asis', fig.width=12, fig.height=7, warning=FALSE, message=FALSE, eval=!is_compact}
# Show comparison of multiple weighting methods if the user selected more than one
if (multi_weight_mode && length(all_models) > 1) {
  section_start(tr("weight_comparison_title", lang))
  emit_heading(tr("weight_comparison_title", lang), 4)
  cat(tr("weight_comparison_desc", lang), "\n\n")

  # --- Comparison table: R², RMSE, AIC, parameters (+ stability if available) ---
  comp_rows <- lapply(names(all_models), function(k) {
    m <- all_models[[k]]
    coefs <- m$coefs  # b (Hill), c (Bottom), d (Top), e (IC50)
    if (is.null(coefs)) {
      row <- data.frame(
        Method = m$label,
        R2 = NA, RMSE = NA, AIC = NA,
        Hill = NA, Bottom = NA, Top = NA, IC50 = NA,
        stringsAsFactors = FALSE
      )
    } else {
      row <- data.frame(
        Method = m$label,
        R2 = round(m$R2, 5),
        RMSE = round(m$RMSE, 3),
        AIC = if (!is.null(m$AIC) && !is.na(m$AIC)) round(m$AIC, 1) else NA,
        Hill = round(abs(coefs[1]), 4),
        Bottom = round(coefs[2], 2),
        Top = round(coefs[3], 2),
        IC50 = if (isTRUE(is_elisa)) format(round(coefs[4], 2), big.mark = ",") else format(coefs[4], scientific = TRUE, digits = 3),
        stringsAsFactors = FALSE
      )
    }
    # Append stability grade if already assessed
    if (!is.null(m$stability)) {
      grade_icon <- function(g) switch(g,
        good = "\U0001F7E2", acceptable = "\U0001F7E1",
        unstable = "\U0001F534", failed = "\u274C", "\u2753")
      row$Stability <- paste(grade_icon(m$stability$stability_grade),
                              tools::toTitleCase(m$stability$stability_grade))
    }
    row
  })
  comp_df <- do.call(rbind, comp_rows)
  col_names <- c(tr("qc_metric", lang), "R\u00B2", "RMSE", "AIC",
                  tr("hill_slope", lang), tr("bottom", lang), tr("top", lang), tr("ic50", lang))
  if ("Stability" %in% names(comp_df)) {
    col_names <- c(col_names, "Stability")
  }
  names(comp_df) <- col_names

  render_table(comp_df, caption = tr("weight_comparison_title", lang))

  # --- AIC recommendation note ---
  aic_vals <- sapply(all_models, function(m) if (!is.null(m$AIC)) m$AIC else NA_real_)
  if (any(!is.na(aic_vals))) {
    best_key <- names(which.min(aic_vals))
    best_label <- all_models[[best_key]]$label
    best_aic <- min(aic_vals, na.rm = TRUE)
    cat(sprintf("\n\n**AIC recommends: %s** (lowest AIC = %.1f). ", best_label, best_aic))
    cat(tr("weight_comparison_primary", lang, all_models[[primary_key]]$label), "\n\n")
  }

  # --- Overlay DRC curves ---
  curve_colors <- c("none" = "#000000", "inv_y" = "#0072B2", "inv_y2" = "#D55E00")
  curve_data_list <- list()
  for (k in names(all_models)) {
    if (is.null(all_models[[k]]$model)) next
    pred_df <- data.frame(conc = conc_range)
    pred_df$p <- predict(all_models[[k]]$model, newdata = pred_df)
    pred_df$method <- all_models[[k]]$label
    curve_data_list[[k]] <- pred_df
  }
  if (length(curve_data_list) > 0) {
  curve_data <- do.call(rbind, curve_data_list)
  plotted_keys <- names(curve_data_list)

  standards_plot <- data_long %>%
    dplyr::filter(!is.na(concentration), SampleType == "Standard")

  cmp_plot <- ggplot() +
    geom_point(data = standards_plot,
               aes(x = concentration, y = .data[[response_var]]),
               size = 3, alpha = 0.6, color = "grey40") +
    geom_line(data = curve_data,
              aes(x = conc, y = p, color = method),
              size = 1.1) +
    scale_color_manual(
      name = tr("col_weighting_label", lang),
      values = setNames(
        curve_colors[plotted_keys],
        sapply(all_models[plotted_keys], function(m) m$label)
      )
    ) +
    theme_classic() +
    theme(legend.position = "top",
          axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = labels$x_label, y = labels$y_label)

  if (isTRUE(is_elisa)) {
    cmp_plot <- cmp_plot +
      scale_x_log10(
        labels = function(x) format(x, digits = 3, big.mark = ","),
        breaks = scales::trans_breaks("log10", function(x) 10^x, n = 6)
      )
  } else {
    cmp_plot <- cmp_plot +
      scale_x_log10(
        limits = c(1e-12, 1e-5),
        breaks = 10^seq(-12, -5, by = 1),
        labels = scales::scientific_format(digits = 2)
      )
  }

  } # end if (length(curve_data_list) > 0)
}
```

```{r weight-comparison-plot, fig.width=12, fig.height=7, warning=FALSE, message=FALSE, eval=!is_compact}
# Separate chunk so render_plot() return value is visible to knitr
if (exists("cmp_plot") && multi_weight_mode && length(all_models) > 1) {
  render_plot(cmp_plot)
}
```

```{r weight-comparison-close, results='asis', eval=!is_compact}
if (multi_weight_mode && length(all_models) > 1) {
  cat(sprintf("\n\n*%s*\n\n",
      tr("weight_comparison_primary", lang, all_models[[primary_key]]$label)))
  section_end()
}

# Simple heteroscedasticity diagnostic for unweighted model
if ("none" %in% names(all_models) && !is.null(all_models[["none"]]$model)) {
  unw_fit <- all_models[["none"]]$model
  unw_resid <- residuals(unw_fit)
  conc_levels <- standards_for_model$StandardConc
  var_by_level <- tapply(unw_resid^2, conc_levels, mean, na.rm = TRUE)
  var_ratio <- max(var_by_level, na.rm = TRUE) / max(min(var_by_level[var_by_level > 0], na.rm = TRUE), 1e-10)
  if (var_ratio > 3) {
    cat(sprintf("\n\n> **Note:** Residual variance ratio across concentration levels = %.1f (>3), suggesting heteroscedasticity. Weighted regression (1/Y or 1/Y²) is recommended.\n\n", var_ratio))
  }
}
```

```{r weighting-suitability, results='asis'}
# H2: document the auto-weighting decision (if taken) at the top of the
# weighting-suitability section, before the full diagnostic block below.
if (!is.null(auto_weighting_result)) {
  weight_labels <- list(none = tr("weight_unweighted", lang), inv_y = "1/Y", inv_y2 = "1/Y²")
  chosen_label <- weight_labels[[auto_weighting_result$chosen_weight]] %||%
                  auto_weighting_result$chosen_weight
  if (isTRUE(auto_weighting_result$fallback)) {
    cat("\n\n", tr("weight_auto_fallback", lang), "\n\n", sep = "")
  } else {
    cat("\n\n", tr("weight_auto_decision", lang,
                    chosen_label,
                    auto_weighting_result$test_name %||% "Brown-Forsythe",
                    auto_weighting_result$statistic %||% NA_real_,
                    auto_weighting_result$p_value %||% NA_real_),
        "\n\n", sep = "")
  }
}

# Formal heteroscedasticity diagnostic (assess_heteroscedasticity)
tryCatch({
  # Determine which model to assess: prefer the unweighted model, otherwise
  # use the first available model as a reference
  hetero_model <- NULL
  hetero_label <- NULL

  if ("none" %in% names(all_models) && !is.null(all_models[["none"]]$model)) {
    hetero_model <- all_models[["none"]]$model
    hetero_label <- "Unweighted (LL.4/LL.3)"
  } else if (!drc_failed_completely) {
    # No unweighted model was fitted; use primary model as proxy
    hetero_model <- all_models[[primary_key]]$model
    hetero_label <- all_models[[primary_key]]$label
  }

  if (!is.null(hetero_model)) {
    hetero_result <- assess_heteroscedasticity(
      model = hetero_model,
      data_long = standards_for_model,
      response_var = response_var
    )

    cat(sprintf("\n\n### %s\n\n", tr("weighting_suitability_title", lang)))

    cat(sprintf("**Test:** %s\n\n", hetero_result$test_name %||% "N/A"))

    if (!is.na(hetero_result$statistic)) {
      cat(sprintf("**Statistic:** %s", fmt_smart(hetero_result$statistic, 3)))
      if (!is.na(hetero_result$p_value)) {
        cat(sprintf(" \u2003 **p-value:** %s", fmt_smart(hetero_result$p_value, 4)))
      }
      cat("\n\n")
    }

    if (!is.na(hetero_result$variance_ratio)) {
      cat(sprintf("**Variance ratio** (max/min across concentration levels): %.1f\n\n",
                  hetero_result$variance_ratio))
    }

    cat(sprintf("**Interpretation:** %s\n\n", hetero_result$interpretation))
    cat(sprintf("**Recommendation:** %s\n\n", hetero_result$recommendation))

    cat("*Note: This diagnostic is advisory only and does not change the selected model.*\n\n")
  }
}, error = function(e) {
  cat(sprintf("\n\n*Weighting suitability assessment could not be performed: %s*\n\n", e$message))
})
```

```{r model-stability-assessment, results='asis', eval=!is_compact}
# Convergence and stability diagnostics for all fitted models
tryCatch({
  if (!drc_failed_completely && length(all_models) > 0) {

    # Assess stability for each fitted model
    stability_results <- lapply(names(all_models), function(k) {
      assess_model_stability(all_models[[k]], all_models[[k]]$label)
    })
    names(stability_results) <- names(all_models)

    emit_heading(tr("model_stability_title", lang), 4)

    # Build summary table
    grade_icon <- function(g) {
      switch(g,
        good       = "\U0001F7E2 Good",
        acceptable = "\U0001F7E1 Acceptable",
        unstable   = "\U0001F534 Unstable",
        failed     = "\u274C Failed",
        "\u2753 Unknown"
      )
    }

    stability_df <- data.frame(
      Model = sapply(stability_results, function(s) s$model_name),
      Method = sapply(stability_results, function(s) s$fit_method %||% "N/A"),
      Converged = sapply(stability_results, function(s) if (s$converged) tr("yes_label", lang) else tr("no_label", lang)),
      SE_Available = sapply(stability_results, function(s) if (s$se_available) tr("yes_label", lang) else tr("no_label", lang)),
      Boundary = sapply(stability_results, function(s) if (s$boundary_flag) tr("yes_label", lang) else tr("no_label", lang)),
      Grade = sapply(stability_results, function(s) grade_icon(s$stability_grade)),
      stringsAsFactors = FALSE
    )
    names(stability_df) <- c(tr("col_model", lang), tr("col_fit_method", lang),
                              tr("col_converged", lang), tr("col_se_available", lang),
                              tr("col_boundary_hit", lang), tr("col_stability_grade", lang))

    render_table(stability_df, caption = tr("model_stability_title", lang))

    # Print detailed warnings / suspicious parameters per model
    for (k in names(stability_results)) {
      sr <- stability_results[[k]]
      has_details <- length(sr$suspicious_params) > 0 || length(sr$warnings) > 0

      if (has_details) {
        cat(sprintf("\n**%s:**\n\n", sr$model_name))
        for (w in sr$warnings) {
          cat(sprintf("- %s\n", w))
        }
        for (sp in sr$suspicious_params) {
          cat(sprintf("- %s\n", sp))
        }
        cat("\n")
      }
    }

    # Store stability info for downstream use (e.g., model comparison table)
    # Attach to all_models entries
    for (k in names(stability_results)) {
      all_models[[k]]$stability <- stability_results[[k]]
    }
  }
}, error = function(e) {
  cat(sprintf("\n\n*Model stability assessment could not be performed: %s*\n\n", e$message))
})
```

```{r lloq-uloq, results='asis'}
# Formal LLOQ/ULOQ determination by back-calculated standard accuracy.
# M1.3: compute lives in determine_lloq_uloq() in reports/analysis_pipeline.R;
# this chunk handles localisation, formatting, and rendering.
# GUI refresh §6.2 / Q4(e): determine_lloq_uloq() is now called once
# in the model-fitting chunk above (so the KPI strip can read its
# results); this chunk reuses .lloq_res from that call. Falls back
# to recomputing if the upstream chunk failed.
tryCatch({
  section_start(tr("lloq_uloq_title", lang))
  emit_heading(tr("lloq_uloq_title", lang), 4)
  cat(tr("lloq_uloq_desc", lang), "\n\n")

  if (!exists(".lloq_res") || is.null(.lloq_res)) {
    .lloq_res <- determine_lloq_uloq(
      standards_for_model = standards_for_model,
      model_fit           = model_fit,
      response_var        = response_var,
      is_elisa            = is_elisa
    )
  }
  backcalc_results <- .lloq_res$backcalc_results
  backcalc_summary <- .lloq_res$backcalc_summary
  passing_concs    <- .lloq_res$passing_concs

  # Add the localised accuracy label to the summary so the rendering code
  # below stays as-is.
  if (nrow(backcalc_summary) > 0) {
    backcalc_summary$accuracy <- ifelse(backcalc_summary$passes,
                                        tr("qc_green", lang),
                                        tr("qc_red", lang))
  }

  if (nrow(backcalc_summary) > 0) {
    # Display back-calculation table
    display_bc <- backcalc_summary %>%
      dplyr::select(nominal_conc, mean_backcalc, mean_recovery, cv_backcalc, accuracy)

    if (isTRUE(is_elisa)) {
      display_bc$nominal_conc <- format(display_bc$nominal_conc, digits = 4, big.mark = ",")
      display_bc$mean_backcalc <- format(display_bc$mean_backcalc, digits = 4, big.mark = ",")
    } else {
      display_bc$nominal_conc <- format(display_bc$nominal_conc, scientific = TRUE, digits = 3)
      display_bc$mean_backcalc <- format(display_bc$mean_backcalc, scientific = TRUE, digits = 3)
    }
    display_bc$mean_recovery <- paste0(round(as.numeric(backcalc_summary$mean_recovery), 1), "%")
    display_bc$cv_backcalc <- ifelse(is.na(backcalc_summary$cv_backcalc), "\u2014",
                                      paste0(round(backcalc_summary$cv_backcalc, 1), "%"))

    units_str <- if (isTRUE(is_elisa)) assay_config$units %||% "pg/mL" else "mol/L"
    names(display_bc) <- c(
      sprintf(tr("col_nominal", lang), units_str),
      sprintf(tr("col_backcalc", lang), units_str),
      tr("col_recovery", lang),
      "CV%",
      tr("col_accuracy", lang)
    )
    render_table(display_bc, caption = tr("backcalc_title", lang))

    # LLOQ/ULOQ already computed by determine_lloq_uloq(). Promote to the
    # outer chunk env (formal_lloq / formal_uloq are read by a later chunk
    # via exists() so we set them either way).
    if (length(passing_concs) >= 2) {
      formal_lloq <- .lloq_res$formal_lloq
      formal_uloq <- .lloq_res$formal_uloq
      cat(sprintf("\n\n**%s:** %s %s\n\n", tr("lloq_label", lang),
                  if (isTRUE(is_elisa)) format(formal_lloq, digits = 4, big.mark = ",") else format(formal_lloq, scientific = TRUE, digits = 3),
                  units_str))
      cat(sprintf("**%s:** %s %s\n\n", tr("uloq_label", lang),
                  if (isTRUE(is_elisa)) format(formal_uloq, digits = 4, big.mark = ",") else format(formal_uloq, scientific = TRUE, digits = 3),
                  units_str))
    } else {
      cat(tr("lloq_uloq_none", lang), "\n\n")
    }
  } else {
    cat(tr("lloq_uloq_none", lang), "\n\n")
  }

  section_end()
}, error = function(e) {
  cat(sprintf("\n\n*LLOQ/ULOQ determination failed: %s*\n\n", e$message))
})
```

`r section_open(tr("drc_analysis", lang), default_open = TRUE)`

```{r dose-response-plots, fig.width=12, fig.height=8, fig.cap="Dose-response curve showing fitted 4-parameter logistic model with observed standard concentrations"}
# Create unified dose-response curve plot
plot_data <- data_long %>% filter(!is.na(concentration))

drc_plot <- ggplot(plot_data, aes(x = concentration, y = .data[[response_var]], color = high_variability)) +
  geom_point(size = 3, aes(text = paste0(
    "Concentration: ", if (isTRUE(is_elisa)) format(concentration, digits = 4, big.mark = ",") else scales::scientific(concentration, digits = 2), "<br>",
    labels$y_label, ": ", round(.data[[response_var]], 2), "<br>",
    "Variability: ", high_variability
  ))) +
  geom_line(data = model_fits, aes(x = conc, y = p), color = "black", size = 1.2, inherit.aes = FALSE) +
  scale_color_manual(
    name = tr("legend_std_quality", lang),
    values = c(
      "Normal Variability" = "#2E86AB",
      "High Variability" = "#E63946"
    ),
    labels = c(
      "Normal Variability" = tr("variability_normal", lang),
      "High Variability"   = tr("variability_high", lang)
    )
  ) +
  theme_classic() +
  theme(
    legend.position = "top",
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 11),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    panel.grid.minor = element_line(color = "grey90", size = 0.25),
    panel.grid.major = element_line(color = "grey85", size = 0.5)
  )

# Set appropriate axis scaling based on assay type
if (isTRUE(is_elisa)) {
  # ELISA: Log scale for wide concentration ranges
  conc_range_factor <- max(plot_data$concentration, na.rm = TRUE) / min(plot_data$concentration, na.rm = TRUE)
  drc_plot <- drc_plot + 
    scale_x_log10(
      name = labels$x_label,
      labels = function(x) format(x, digits = 3, big.mark = ","),
      breaks = scales::trans_breaks("log10", function(x) 10^x, n = 6)
    )
} else {
  # RBA: Log scale with scientific notation
  drc_plot <- drc_plot + 
    scale_x_log10(
      name = labels$x_label,
      breaks = scales::trans_breaks("log10", function(x) 10^x),
      labels = scales::trans_format("log10", scales::math_format(10^.x))
    )
}

drc_plot <- drc_plot + 
  ylab(labels$y_label) +
  ggtitle(paste(tr("four_pl_coefficients", lang), "-", toupper(assay_config$assay_type)))

# Display plot
render_plot(drc_plot, tooltip = "text", plotly_layout = list(showlegend = TRUE))
```

`r section_close()`

## `r tr("model_parameters", lang)`

`r section_open(tr("four_pl_coefficients", lang), default_open = TRUE)`

```{r model-stats}
# Extract and display model coefficients with correct parameter order
tryCatch({
  
  coefficients <- summary(model_fit)$coefficients
  
  # Check if coefficients have expected structure
  if (is.matrix(coefficients) && nrow(coefficients) >= 4) {
    
    # DRC package LL.4() parameter order: b (Hill slope), c (Lower asymptote), d (Upper asymptote), e (IC50)
    actual_param_names <- c(tr("hill_slope", lang), tr("bottom", lang), tr("top", lang), tr("ic50", lang))
    
    # Ensure we don't exceed available rows
    n_params <- min(length(actual_param_names), nrow(coefficients))
    
    # Create results data frame with correct parameter mapping
    drc_results <- data.frame(
      Parameter = actual_param_names[1:n_params],
      Value = coefficients[1:n_params, "Estimate"],
      StdError = if ("Std. Error" %in% colnames(coefficients)) {
        coefficients[1:n_params, "Std. Error"]
      } else {
        rep(NA, n_params)
      },
      PValue = if ("Pr(>|t|)" %in% colnames(coefficients)) {
        coefficients[1:n_params, "Pr(>|t|)"]
      } else {
        rep(NA, n_params)
      }
    )
    
    render_table(
      drc_results,
      caption = tr("four_pl_coefficients", lang),
      col_names = c(tr("col_parameter", lang), tr("col_estimate", lang), tr("col_std_error", lang), tr("col_pvalue", lang))
    )
    
  } else {
    cat("Model coefficients structure unexpected. Available coefficients:\n")
    print(coefficients)
  }
  
}, error = function(e) {
  cat("Error extracting model coefficients:", e$message, "\n")
  cat("Model summary:\n")
  print(summary(model_fit))
})
```

`r section_close()`

**`r tr("model_fit_stats", lang)`**

- **R²** = `r round(R2, 3)`
- **RMSE** = `r round(RMSE, 3)` `r labels$y_unit`
- **`r tr("standards_used", lang)`** = `r nrow(standards_for_model)` / `r length(unique(standards_data$StandardConc))`

```{r save-model-stats, include=FALSE}
# Save model statistics for multiwavelength executive summary
model_stats <- list(
  r_squared = R2,
  rmse = RMSE,
  ic50 = coef(model_fit)[4],
  hill_slope = coef(model_fit)[1],
  n_standards = nrow(standards_for_model),
  n_unique_concs = length(unique(standards_data$StandardConc)),
  weight_method = weight_desc
)

# Note: mean_sample_cv is computed later in save-sample-stats chunk
# after replicate_stats is available (BUG_008 fix)

# write_json_safe is defined in global.R (loaded before rendering)
write_json_safe(model_stats, file.path(output_dir, "model_stats.json"))

# Env-handoff: assign into the environment passed from the Shiny host so
# the KPI strip can populate without reading the JSON file back from disk.
if (is.environment(params$stats_env)) {
  .se <- params$stats_env
  .se$model_stats <- model_stats
}
```

### `r tr("std_backcalc_title", lang)`

`r section_open("Click to expand standard back-calculation and recovery")`

```{r standard-backcalculation, results='asis', eval=!is_compact}
# M1.4: compute_standard_recovery() in reports/analysis_pipeline.R produces
# both data frames; this chunk now only formats and renders.
.recovery <- compute_standard_recovery(data_long       = data_long,
                                       model_fit       = model_fit,
                                       response_var    = response_var,
                                       is_elisa        = is_elisa)
standards_backcalc <- .recovery$standards_backcalc
recovery_summary   <- .recovery$recovery_summary

# Format for display
if (isTRUE(is_elisa)) {
  recovery_display <- recovery_summary %>%
    dplyr::mutate(
      `Nominal Conc.` = format(nominal, digits = 4, nsmall = 1, big.mark = ","),
      `Back-Calculated` = format(mean_backcalc, digits = 4, nsmall = 1, big.mark = ","),
      `SD` = ifelse(is.na(sd_backcalc), "—", format(sd_backcalc, digits = 3, nsmall = 1)),
      `Recovery (%)` = ifelse(is.na(mean_recovery), "—", format(mean_recovery, digits = 4, nsmall = 1))
    ) %>%
    dplyr::select(`Nominal Conc.`, n, `Back-Calculated`, SD, `Recovery (%)`)
} else {
  recovery_display <- recovery_summary %>%
    dplyr::mutate(
      `Nominal Conc.` = format(nominal, scientific = TRUE, digits = 3),
      `Back-Calculated` = format(mean_backcalc, scientific = TRUE, digits = 3),
      `SD` = ifelse(is.na(sd_backcalc), "—", format(sd_backcalc, scientific = TRUE, digits = 2)),
      `Recovery (%)` = ifelse(is.na(mean_recovery), "—", format(mean_recovery, digits = 4, nsmall = 1))
    ) %>%
    dplyr::select(`Nominal Conc.`, n, `Back-Calculated`, SD, `Recovery (%)`)
}

conc_unit <- if (isTRUE(is_elisa)) paste0(assay_config$units %||% "pg/mL") else "mol/L"
names(recovery_display) <- c(sprintf(tr("col_nominal", lang), conc_unit), tr("col_n", lang),
                              sprintf(tr("col_backcalc", lang), conc_unit),
                              paste0("SD (", conc_unit, ")"),
                              tr("col_recovery", lang))

render_table(recovery_display,
            caption = tr("std_backcalc_caption", lang))

# Report overall recovery statistics
mean_overall_recovery <- mean(recovery_summary$mean_recovery, na.rm = TRUE)
cat(sprintf("\n**%s**\n", tr("overall_recovery", lang, mean_overall_recovery)))
if (mean_overall_recovery >= 80 && mean_overall_recovery <= 120) {
  cat(tr("recovery_acceptable", lang), "\n\n")
} else {
  cat(tr("recovery_outside", lang), "\n\n")
}
```

`r section_close()`

## `r tr("sample_results", lang)`

```{r data-quality-overview, results='asis'}
total_wells <- nrow(data_long)
std_wells <- sum(data_long$SampleType == "Standard", na.rm = TRUE)
sample_wells <- sum(data_long$SampleType == "Sample", na.rm = TRUE)
blank_wells <- sum(data_long$SampleType %in% c("Blank", "NSB", "B0"), na.rm = TRUE)
excluded_wells <- sum(is.na(data_long$MeasurementValue) | !is.finite(data_long$MeasurementValue))

dq_lines <- c(
  "**Data Quality Overview**  ",
  sprintf("Total wells: %d | Standards: %d | Samples: %d | Controls: %d  ",
          total_wells, std_wells, sample_wells, blank_wells)
)
if (excluded_wells > 0) {
  dq_lines <- c(dq_lines, sprintf("**Excluded wells:** %d (missing or non-finite values)  ", excluded_wells))
}
if (exists("outlier_flags") && nrow(outlier_flags) > 0) {
  dq_lines <- c(dq_lines, sprintf("**Outliers flagged:** %d  ", nrow(outlier_flags)))
}
emit_styled_block(dq_lines,
  html_style = "background: #E3F2FD; padding: 12px; border-radius: 8px; border-left: 4px solid #1976D2; margin: 15px 0;")
```

```{r sample-analysis, include=FALSE}
# M1.1: quantify_samples() in reports/analysis_pipeline.R does the per-well
# delta-method prediction, optional outlier detection (Shapiro->MAD/Dixon's Q/
# Grubbs), replicate-group layered uncertainty (HP5/HP6), within-range and
# LLOQ/ULOQ flagging, and (for ELISA + tissue_weights.json) pg/g tissue.
# classify_range and flag_range are passed in so the function calls them
# without re-creating the chunk-scoped closures.
#
# H4 cache: as for fit_all_models, all variants in a single render_reports()
# run share the same per-well quantification; serialise the first render's
# result and short-circuit the rest. Pipeline clears the file at run start.
.samp_cache_path <- file.path(output_dir, ".quantify_samples_cache.rds")
if (file.exists(.samp_cache_path)) {
  .samp_res <- readRDS(.samp_cache_path)
} else {
  .samp_res <- quantify_samples(
    data_long           = data_long,
    response_var        = response_var,
    model_fit           = model_fit,
    standards_for_model = standards_for_model,
    analysis_config     = analysis_config,
    STATS_CONFIG        = STATS_CONFIG,
    is_elisa            = is_elisa,
    output_dir          = output_dir,
    classify_range      = classify_range,
    flag_range          = flag_range
  )
  saveRDS(.samp_res, .samp_cache_path)
}
list2env(.samp_res, envir = environment())
```

### `r tr("detailed_summary", lang)`

```{r sample-results-table, results='asis'}
# Wrap entire chunk in tryCatch for debugging
tryCatch({
  
  if (exists("replicate_stats") && !is.null(replicate_stats) && nrow(replicate_stats) > 0) {
    
    # Check if tissue weights were provided (robust to NA)
    has_tissue_weights <- FALSE
    if (exists("sample_results") && !is.null(sample_results) && "TissueWeight_mg" %in% names(sample_results)) {
      tissue_check <- tryCatch({
        any(!is.na(sample_results$TissueWeight_mg) & sample_results$TissueWeight_mg > 0, na.rm = TRUE)
      }, error = function(e) FALSE)
      has_tissue_weights <- isTRUE(tissue_check)
    }
    
    # Determine concentration column name based on assay type
    conc_col_name <- if (isTRUE(is_elisa)) "Mean Conc. (pg/mL)" else "Mean Conc. (mol/L)"
    
    # Create clean results table (include quant_status if available)
    base_cols <- c("Replicate", "sampleID", "n_replicates", "mean_conc_formatted", "ci_formatted", "cv_formatted")
    if ("quant_status" %in% names(replicate_stats)) {
      results_display <- replicate_stats %>%
        dplyr::select(dplyr::all_of(base_cols), quant_status) %>%
        dplyr::mutate(quant_status = dplyr::case_when(
          quant_status == "Interpolated"         ~ tr("status_interpolated", lang),
          quant_status == "Extrapolated (below)" ~ tr("status_extrapolated_below", lang),
          quant_status == "Extrapolated (above)" ~ tr("status_extrapolated_above", lang),
          quant_status == "Not estimable"        ~ tr("status_not_estimable", lang),
          TRUE ~ quant_status
        ))
      names(results_display) <- c(tr("col_replicate_group", lang), tr("col_sample_ids", lang), tr("col_n", lang), conc_col_name, tr("col_ci", lang), tr("col_cv", lang), tr("qc_status", lang))
    } else {
      results_display <- replicate_stats %>%
        dplyr::select(dplyr::all_of(base_cols))
      names(results_display) <- c(tr("col_replicate_group", lang), tr("col_sample_ids", lang), tr("col_n", lang), conc_col_name, tr("col_ci", lang), tr("col_cv", lang))
    }
    
    # Add tissue-based concentrations if available
    if (isTRUE(is_elisa) && isTRUE(has_tissue_weights)) {
      tissue_results <- tryCatch({
        sample_results %>%
          dplyr::filter(!is.na(TissueWeight_mg) & TissueWeight_mg > 0) %>%
          dplyr::group_by(Replicate) %>%
          dplyr::summarise(
            tissue_weight = dplyr::first(TissueWeight_mg),
            mean_tissue_conc = mean(concentration_pg_per_g, na.rm = TRUE),
            .groups = "drop"
          ) %>%
          dplyr::mutate(
            tissue_weight_formatted = format(tissue_weight, digits = 4, nsmall = 1),
            tissue_conc_formatted = format(mean_tissue_conc, digits = 4, nsmall = 2, big.mark = ",")
          )
      }, error = function(e) {
        warning("Could not process tissue weight data: ", e$message)
        NULL
      })
      
      if (!is.null(tissue_results) && nrow(tissue_results) > 0) {
        results_display <- results_display %>%
          dplyr::left_join(
            tissue_results %>% dplyr::select(Replicate, tissue_weight_formatted, tissue_conc_formatted), 
            by = c("Replicate Group" = "Replicate")
          )
        
        # Rename tissue columns if they exist
        if ("tissue_weight_formatted" %in% names(results_display)) {
          names(results_display)[names(results_display) == "tissue_weight_formatted"] <- "Tissue Mass [mg]"
        }
        if ("tissue_conc_formatted" %in% names(results_display)) {
          names(results_display)[names(results_display) == "tissue_conc_formatted"] <- "Conc. (pg/g tissue)"
        }
      }
    }
    
    # Generate table caption
    caption_text <- tr("sample_results_caption", lang, toupper(assay_config$assay_type))
    if (isTRUE(is_elisa) && isTRUE(has_tissue_weights)) {
      caption_text <- paste(caption_text, tr("with_tissue", lang))
    }
    
    render_table(results_display, caption = caption_text)

    # Task B (HP6): Add truncation footnote when any CIs were truncated to 0
    if ("ci_truncated" %in% names(replicate_stats)) {
      any_truncated <- any(replicate_stats$ci_truncated, na.rm = TRUE)
      if (isTRUE(any_truncated)) {
        truncated_groups <- replicate_stats$Replicate[replicate_stats$ci_truncated]
        if (is_html_out()) {
          cat(sprintf(
            "\n<small>\u2020 Lower bound truncated to 0 (original value was negative) for: %s</small>\n\n",
            paste(truncated_groups, collapse = ", ")
          ))
        } else {
          cat(sprintf(
            "\n*\u2020 Lower bound truncated to 0 (original value was negative) for: %s*\n\n",
            paste(truncated_groups, collapse = ", ")
          ))
        }
      }
    }

    # Save cleaned results files (Task B: include both raw and display CI bounds in CSV)
    if (exists("sample_results_clean") && !is.null(sample_results_clean)) {
      write.csv(sample_results_clean,
               file.path(output_dir, "unknown_results.csv"),
               row.names = FALSE)
    }

    if (exists("replicate_summary") && !is.null(replicate_summary)) {
      write.csv(replicate_summary,
               file.path(output_dir, "unknown_results_summary.csv"),
               row.names = FALSE)
    }
    
    cat("\n**", tr("output_files_created", lang), "**\n", sep = "")
    cat("- ", tr("individual_results", lang), "\n", sep = "")
    cat("- ", tr("summary_results", lang), "\n\n", sep = "")
    
    # Display quality assessment (robust to NA in cv_percent)
    high_cv_groups <- tryCatch({
      replicate_stats %>% 
        dplyr::filter(!is.na(cv_percent) & cv_percent > 30)
    }, error = function(e) {
      data.frame()
    })
    
    if (!is.null(high_cv_groups) && nrow(high_cv_groups) > 0) {
      cat("**", tr("quality_alert", lang), "**\n", sep = "")
      cat("- ", tr("high_cv_groups", lang, paste(high_cv_groups$Replicate, collapse = ", ")), "\n", sep = "")
      cat("- ", tr("check_preparation", lang), "\n\n", sep = "")
    } else {
      cat("**", tr("quality_pass", lang), "**\n\n", sep = "")
    }
    
  } else {
    cat("*", tr("no_samples_quantified", lang), "*\n\n", sep = "")
    
    if (exists("samples_to_analyze") && !is.null(samples_to_analyze) && nrow(samples_to_analyze) > 0) {
      cat("**Possible reasons:**\n")
      cat("- Response values outside the standard curve range\n")
      cat("- Model fitting issues\n")
      if (isTRUE(is_elisa)) {
        cat("- Invalid %B/B0 normalization\n")
        cat("- Control well hierarchy violations\n")
      }
      cat("- Mathematical prediction errors\n")
    } else {
      cat("**No sample data found for analysis.**\n")
      cat("Note: Only wells marked as 'Sample' are quantified.\n")
      cat("Blanks, NSB, TA, and Standards are excluded from quantification.\n")
    }
  }
  
}, error = function(e) {
  cat("\n**Error in sample results table:**\n")
  cat("Error message:", e$message, "\n")
  cat("This section will be skipped.\n\n")
})
```

```{r outlier-results, results='asis'}
if (isTRUE(analysis_config$enable_outlier_detection)) {
  section_start("Click to expand outlier detection results")
  emit_heading(tr("outlier_title", lang), 4)

  # Phase 1.5: clarify that flagged != removed from the raw table
  cat("> ", tr("outlier_flagged_not_removed_note", lang), "\n\n", sep = "")

  if (exists("outlier_flags") && nrow(outlier_flags) > 0) {
    cat(tr("outlier_desc", lang, analysis_config$outlier_min_n %||% 3), "\n\n")

    n_outliers <- nrow(outlier_flags)
    n_groups <- length(unique(outlier_flags$Replicate))
    cat(tr("outlier_found", lang, n_outliers, n_groups), "\n\n")

    outlier_display <- outlier_flags %>%
      dplyr::select(Well, Replicate, method)
    names(outlier_display) <- c("Well", "Replicate Group", "Method")

    render_table(outlier_display, caption = tr("outlier_title", lang))
  } else {
    cat(tr("outlier_none", lang), "\n\n")
  }

  # Method log (Shapiro-Wilk pre-test details)
  if (exists("outlier_method_log") && nrow(outlier_method_log) > 0) {
    cat("\n**Outlier detection method per replicate group (Shapiro-Wilk pre-test enabled):**\n\n")
    for (i in seq_len(nrow(outlier_method_log))) {
      row <- outlier_method_log[i, ]
      cat(sprintf("- %s: %s (Shapiro-Wilk p = %.4f)\n", row$Replicate, row$method_used, row$shapiro_p))
    }
    cat("\n")
  }

  section_end()
}
```

```{r ci-method-note, results='asis'}
if (isTRUE(analysis_config$ci_method == "bootstrap")) {
  cat(sprintf("\n*%s*\n\n", tr("ci_bootstrap_note", lang)))
  cat(sprintf("*Replicate-level CIs computed via percentile bootstrap (%d resamples).*\n\n",
              STATS_CONFIG$bootstrap_iterations))
} else {
  cat(sprintf("\n*%s*\n\n", tr("ci_tdist_note", lang)))
  cat("*Replicate-level CIs reflect measurement variability using Student's t-distribution (df = n-1).*\n\n")
}
cat(sprintf("*%s*\n\n", tr("ci_delta_method_note", lang)))
```

```{r ci-asymmetric-footnote, results='asis'}
cat(sprintf("\n*%s*\n\n", tr("ci_asymmetric_footnote", lang)))
```

```{r save-sample-stats, include=FALSE}
# BUG_008 fix: Update model_stats with mean_sample_cv now that replicate_stats exists
if (exists("model_stats") && exists("replicate_stats") &&
    !is.null(replicate_stats) && nrow(replicate_stats) > 0) {
  model_stats$mean_sample_cv <- mean(replicate_stats$cv_percent, na.rm = TRUE)
} else if (exists("model_stats")) {
  model_stats$mean_sample_cv <- NA_real_
}
# Re-save model_stats with the updated mean_sample_cv
if (exists("model_stats") && exists("write_json_safe")) {
  write_json_safe(model_stats, file.path(output_dir, "model_stats.json"))
}
# Env-handoff update: overwrite with the final mean_sample_cv-enriched copy
if (exists("model_stats") && is.environment(params$stats_env)) {
  .se <- params$stats_env
  .se$model_stats <- model_stats
}
```

```{=openxml}
<w:p>
  <w:pPr>
    <w:sectPr>
      <w:pgSz w:w="12240" w:h="15840"/>
      <w:pgMar w:top="1440" w:right="1440" w:bottom="1440" w:left="1440"
               w:header="720" w:footer="720"/>
    </w:sectPr>
  </w:pPr>
</w:p>
```

## `r tr("detailed_summary", lang)`

`r section_open(tr("detailed_summary", lang), default_open = TRUE)`

```{r table-legend, results='asis'}
render_table_legend()
```

```{r summary-results-table}
if (exists("replicate_summary") && !is.null(replicate_summary) && nrow(replicate_summary) > 0) {

  # Item 9: compute row-highlight indices BEFORE formatting (need raw numerics)
  .amber_rows <- which(
    (if ("range_flag" %in% names(replicate_summary))
       !is.na(replicate_summary$range_flag) &
       replicate_summary$range_flag %in% c(">ULOQ", "<LLOQ")
     else rep(FALSE, nrow(replicate_summary))) |
    (if ("quant_status" %in% names(replicate_summary))
       !is.na(replicate_summary$quant_status) &
       replicate_summary$quant_status %in% c("Extrapolated (above)", "Extrapolated (below)")
     else rep(FALSE, nrow(replicate_summary)))
  )
  .red_rows <- if ("cv_percent" %in% names(replicate_summary)) {
    which(!is.na(replicate_summary$cv_percent) &
          as.numeric(replicate_summary$cv_percent) > 30)
  } else {
    integer(0)
  }

  # B4: pre-extract flag/status columns before the mutate to avoid data-masking issues
  .rf_col <- if ("range_flag"   %in% names(replicate_summary)) replicate_summary[["range_flag"]]   else rep(NA_character_, nrow(replicate_summary))
  .qs_col <- if ("quant_status" %in% names(replicate_summary)) replicate_summary[["quant_status"]] else rep(NA_character_, nrow(replicate_summary))

  # Item 18: format columns; collapse SD + SE into a single "SD \u00b1 SE" column
  summary_display <- replicate_summary %>%
    dplyr::mutate(
      mean_concentration = if (isTRUE(is_elisa)) {
        format(as.numeric(mean_concentration), digits = 4, nsmall = 1, big.mark = ",")
      } else {
        format(as.numeric(mean_concentration), scientific = TRUE, digits = 3)
      },
      sd_se = if (isTRUE(is_elisa)) {
        paste0(
          ifelse(is.na(std_dev),   "\u2014",
                 format(as.numeric(std_dev),   digits = 3, nsmall = 1)),
          " \u00b1 ",
          ifelse(is.na(std_error), "\u2014",
                 format(as.numeric(std_error), digits = 3, nsmall = 1)))
      } else {
        paste0(
          ifelse(is.na(std_dev),   "\u2014",
                 format(as.numeric(std_dev),   scientific = TRUE, digits = 2)),
          " \u00b1 ",
          ifelse(is.na(std_error), "\u2014",
                 format(as.numeric(std_error), scientific = TRUE, digits = 2)))
      },
      ci_95 = if (isTRUE(is_elisa)) {
        paste0("[", format(as.numeric(ci_lower_95), digits = 3, nsmall = 1), " \u2013 ",
               format(as.numeric(ci_upper_95),      digits = 3, nsmall = 1), "]")
      } else {
        paste0("[", format(as.numeric(ci_lower_95), scientific = TRUE, digits = 2), " \u2013 ",
               format(as.numeric(ci_upper_95),      scientific = TRUE, digits = 2), "]")
      },
      cv_percent = ifelse(is.na(cv_percent), "\u2014",
                          format(as.numeric(cv_percent), digits = 3, nsmall = 1)),
      status_combined = dplyr::case_when(
        !is.na(.rf_col) & .rf_col != ""                                        ~ .rf_col,
        !is.na(.qs_col) & !(.qs_col %in% c("Quantified", "Within Range", "")) ~ .qs_col,
        TRUE                                                                    ~ "OK"
      ),
      mean_bb0 = if ("mean_response" %in% names(.)) {
        ifelse(is.na(mean_response), "\u2014",
               paste0(round(as.numeric(mean_response), 2), "%"))
      } else {
        NULL
      }
    ) %>%
    # Drop CI raw bounds, separate SD/SE (replaced by sd_se), and internal cols
    dplyr::select(-ci_lower_95, -ci_upper_95, -std_dev, -std_error,
                  -any_of(c("within_range", "range_flag", "quant_status", "mean_response",
                            "ci_lower_model",      "ci_upper_model",
                            "ci_lower_replicate",  "ci_upper_replicate",
                            "ci_lower_raw",        "ci_upper_raw",
                            "ci_truncated"))) %>%
    dplyr::relocate(sd_se,           .after  = mean_concentration) %>%
    dplyr::relocate(ci_95,           .before = cv_percent) %>%
    dplyr::relocate(status_combined, .after  = cv_percent) %>%
    { if ("mean_bb0" %in% names(.)) dplyr::relocate(., mean_bb0, .before = status_combined) else . }

  # Item 18 (HTML only): merge Sample IDs + n as a small grey line inside the
  # Group cell; drop those columns so they don't appear as separate columns.
  # NOTE: analysis_pipeline.R renames Replicate\u2192replicate_group and
  #       sampleID\u2192sample_ids; use those actual column names here.
  if (is_html_out()) {
    .has_n   <- "n_replicates" %in% names(summary_display)
    .has_sid <- "sample_ids"   %in% names(summary_display)
    # Pre-extract outside mutate to avoid data-masking lookup issues.
    .sids    <- if (.has_sid) summary_display[["sample_ids"]]
                else          character(nrow(summary_display))
    summary_display <- summary_display %>%
      dplyr::mutate(
        replicate_group = if (.has_n) {
          paste0(replicate_group,
                 '<br><span class="bs-replicate-meta">',
                 .sids,
                 ' \u00b7 n=', n_replicates, '</span>')
        } else {
          paste0(replicate_group,
                 '<br><span class="bs-replicate-meta">',
                 .sids,
                 '</span>')
        }
      ) %>%
      dplyr::select(-any_of(c("sample_ids", "n_replicates")))
    rm(.has_n, .has_sid, .sids)

    # B4: convert status_combined to .bs-status-pill HTML spans
    summary_display <- summary_display %>%
      dplyr::mutate(
        status_combined = dplyr::case_when(
          status_combined == ">ULOQ" ~ '<span class="bs-status-pill is-warn">&gt;Cal.&nbsp;high</span>',
          status_combined == "<LLOQ" ~ '<span class="bs-status-pill is-warn">&lt;Cal.&nbsp;low</span>',
          status_combined == "OK"    ~ '<span class="bs-status-pill is-pass">OK</span>',
          TRUE                       ~ paste0('<span class="bs-status-pill is-warn">', status_combined, '</span>')
        )
      )
  } else {
    # DOCX/PDF: drop n; keep sample_ids as its own column (no HTML tags)
    summary_display <- summary_display %>%
      dplyr::select(-any_of("n_replicates"))
  }

  # Build column names dynamically based on output format and available columns
  conc_unit <- if (isTRUE(is_elisa)) assay_config$units %||% "pg/mL" else "mol/L"

  if (is_html_out()) {
    base_names <- c("Group", "Sample Type",
                    paste0("Mean (", conc_unit, ")"),
                    paste0("SD \u00b1 SE (", conc_unit, ")"),
                    "95% CI", "CV [%]")
  } else {
    base_names <- c("Group", "Sample IDs", "Sample Type",
                    paste0("Mean (", conc_unit, ")"),
                    paste0("SD \u00b1 SE (", conc_unit, ")"),
                    "95% CI", "CV [%]")
  }

  if ("mean_bb0"        %in% names(summary_display)) base_names <- c(base_names, "%B/B0")
  if ("status_combined" %in% names(summary_display)) base_names <- c(base_names, "Status")

  # Tissue concentration columns if present
  if ("TissueWeight_mg" %in% names(summary_display)) {
    summary_display <- summary_display %>%
      dplyr::mutate(
        TissueWeight_mg = ifelse(
          is.na(TissueWeight_mg), "\u2014",
          format(TissueWeight_mg, digits = 4, nsmall = 1)),
        concentration_pg_per_g = ifelse(
          is.na(concentration_pg_per_g), "\u2014",
          format(concentration_pg_per_g, digits = 4, nsmall = 1, big.mark = ","))
      )
    base_names <- c(base_names, "Tissue Mass [mg]", "Conc. (pg/g)")
  }

  summary_display <- summary_display %>%
    dplyr::select(-any_of("total_amount_pg"))

  names(summary_display) <- base_names[1:ncol(summary_display)]

  # Item 9: row_highlight wires amber/red-left treatments via render_table()
  # escape=FALSE is required so the HTML span injected into replicate_group renders
  render_table(summary_display,
               caption       = tr("detailed_caption", lang),
               escape        = !is_html_out(),
               row_highlight = list(amber_rows    = .amber_rows,
                                    red_left_rows = .red_rows))
}
```

`r section_close()`

```{=openxml}
<w:p>
  <w:pPr>
    <w:sectPr>
      <w:pgSz w:w="15840" w:h="12240" w:orient="landscape"/>
      <w:pgMar w:top="1440" w:right="1440" w:bottom="1440" w:left="1440"
               w:header="720" w:footer="720"/>
    </w:sectPr>
  </w:pPr>
</w:p>
```

## `r tr("sample_variability", lang)`

```{r sample-boxplot, fig.width=10, fig.height=7, results='asis', warning=FALSE, message=FALSE}
if (exists("sample_results") && !is.null(sample_results) && nrow(sample_results) > 0) {
  
  sample_plot_data <- sample_results %>%
    dplyr::filter(!is.na(estimated_concentration), is.finite(estimated_concentration)) %>%
    dplyr::mutate(
      Replicate_wrapped = stringr::str_wrap(Replicate, width = 15),
      within_range = mapply(classify_range, response_value, estimated_concentration,
                            MoreArgs = list(is_elisa_assay = is_elisa)),
      range_flag = mapply(flag_range, response_value, estimated_concentration,
                          MoreArgs = list(is_elisa_assay = is_elisa))
    )
  
  if (nrow(sample_plot_data) > 0) {

    # Dynamic font sizing based on number of groups
    n_groups <- length(unique(sample_plot_data$Replicate_wrapped))
    x_font_size <- if (n_groups > 20) 7 else if (n_groups > 10) 9 else 11

    # For very large experiments, add panel column before creating ggplot
    use_faceting <- n_groups > 30
    if (use_faceting) {
      sample_plot_data$panel <- ceiling(as.numeric(factor(sample_plot_data$Replicate)) / 15)
    }

    # Dynamic figure height: taller when faceted
    fig_h <- if (use_faceting) {
      n_panels <- max(sample_plot_data$panel)
      max(7, n_panels * 4.5)
    } else {
      7
    }

    # Create boxplot
    sample_box <- ggplot(sample_plot_data,
                        aes(x = factor(Replicate_wrapped),
                            y = estimated_concentration)) +
      geom_boxplot(outlier.shape = NA) +
      geom_jitter(aes(color = within_range,
                     text = paste0(
                       "Replicate: ", Replicate, "<br>",
                       "Sample ID: ", SampleID, "<br>",
                       "Concentration: ",
                       if (isTRUE(is_elisa)) {
                         format(estimated_concentration, digits = 4, big.mark = ",")
                       } else {
                         scales::scientific(estimated_concentration, digits = 2)
                       }, "<br>",
                       "Range Status: ", within_range
                     )),
                 width = 0.15, size = 2) +
      xlab(tr("col_replicate_group", lang)) +
      ylab(if (isTRUE(is_elisa)) {
        paste0("Estimated Concentration (", assay_config$units %||% "pg/mL", ")")
      } else {
        "Estimated Concentration (mol/L)"
      }) +
      theme_classic() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1, size = x_font_size),
        axis.title.x = element_text(margin = margin(t = 12)),
        axis.title.y = element_text(margin = margin(r = 12)),
        legend.position = "top",
        panel.spacing.y = unit(1.5, "lines"),
        strip.text = element_blank()
      ) +
      labs(color = "Within linear range?")

    # Add log scale if appropriate
    conc_range <- max(sample_plot_data$estimated_concentration, na.rm = TRUE) /
                  min(sample_plot_data$estimated_concentration, na.rm = TRUE)

    if (!isTRUE(is_elisa) || conc_range > 100) {
      sample_box <- sample_box +
        scale_y_log10(labels = if (isTRUE(is_elisa)) {
          function(x) format(x, big.mark = ",")
        } else {
          scales::scientific_format(digits = 2)
        })
    }

    # Add faceting for large experiments
    if (use_faceting) {
      sample_box <- sample_box + facet_wrap(~panel, scales = "free_x", ncol = 1)
    }

    # Section wrapper + description
    section_start("Sample variability visualization")
    cat(sprintf("*%s*\n\n", tr("sample_variability_desc", lang)))
  }
}
```

```{r sample-boxplot-render, fig.width=10, fig.height=7, warning=FALSE, message=FALSE}
# Separate chunk so render_plot() return value is visible to knitr
if (exists("sample_box") && !is.null(sample_box)) {
  fig_h <- if (exists("use_faceting") && isTRUE(use_faceting)) {
    n_panels <- max(sample_plot_data$panel)
    max(7, n_panels * 4.5)
  } else { 7 }
  render_plot(sample_box, tooltip = "text", height = fig_h * 96,
              plotly_layout = list(margin = list(b = 120, l = 80)))
}
```

```{r sample-boxplot-close, results='asis'}
if (exists("sample_box") && !is.null(sample_box)) {
  section_end()
}
```

## `r tr("heatmap_title", lang)`

`r section_open(tr("heatmap_title", lang), default_open = TRUE)`

```{r plate-heatmap, fig.width=10, fig.height=6, results='asis', warning=FALSE, message=FALSE}
tryCatch({
  cat(tr("heatmap_desc", lang), "\n\n")

  # Reconstruct plate layout from long data
  heatmap_data <- data_long %>%
    dplyr::filter(!is.na(MeasurementValue)) %>%
    dplyr::mutate(
      Row = factor(Row, levels = LETTERS[1:8]),
      Column = as.numeric(gsub("[^0-9]", "", as.character(Column)))
    ) %>%
    dplyr::filter(!is.na(Row) & !is.na(Column))

  if (nrow(heatmap_data) > 0) {
    heatmap_plot <- ggplot(heatmap_data, aes(x = factor(Column), y = Row, fill = MeasurementValue)) +
      geom_tile(color = "white", size = 0.5) +
      geom_text(aes(label = sprintf("%.3f", MeasurementValue)), size = 2, color = "black") +
      scale_fill_gradient2(low = "#2166AC", mid = "#F7F7F7", high = "#B2182B",
                           midpoint = median(heatmap_data$MeasurementValue, na.rm = TRUE),
                           name = "Value") +
      scale_y_discrete(limits = rev(levels(factor(heatmap_data$Row, levels = LETTERS[1:8])))) +
      labs(x = "Column", y = "Row", title = tr("heatmap_title", lang)) +
      theme_minimal() +
      theme(
        panel.grid = element_blank(),
        axis.text = element_text(size = 10, face = "bold"),
        plot.title = element_text(hjust = 0.5)
      )

    # Static heatmap for all outputs (no plotly conversion needed)
    print(heatmap_plot)
  }
}, error = function(e) {
  cat(sprintf("\n\n*Heatmap could not be generated: %s*\n\n", e$message))
})
```

`r section_close()`

```{r plate-positional-qc, results='asis', eval=!is_compact}
section_start("Click to expand plate positional quality")
emit_heading("Plate Positional Quality", 4)
positional <- if (exists("data_long")) assess_plate_positional(data_long) else NULL
if (!is.null(positional)) {
  if (positional$flagged_rows > 0 || positional$flagged_cols > 0) {
    cat(sprintf("**Warning:** %d row(s) and %d column(s) deviate >2 SD from plate mean, suggesting possible edge effects.\n\n",
                positional$flagged_rows, positional$flagged_cols))
  } else {
    cat("No significant row or column bias detected (all within 2 SD of plate mean).\n\n")
  }
  row_display <- positional$row_stats
  row_display$Mean <- round(row_display$Mean, 3)
  row_display$SD   <- round(row_display$SD,   3)
  # Empty-flag cells: show a pass pill for HTML, blank for DOCX/PDF
  if (is_html_out())
    row_display$Flag <- ifelse(row_display$Flag == "",
                               '<span class="bs-status-pill is-pass">OK</span>',
                               row_display$Flag)
  names(row_display) <- c("Row", "Mean Response", "SD", "Flag")
  render_table(row_display, caption = "Response by Row", escape = !is_html_out())
}
section_end()
```

## `r tr("drc_with_samples", lang)`

`r section_open(tr("drc_with_samples", lang), default_open = TRUE)`

```{r drc-with-samples, fig.width=12, fig.height=8, warning=FALSE, message=FALSE, fig.cap="Dose-response curve with sample concentration estimates overlaid on the fitted model"}
if (exists("sample_results") && !is.null(sample_results) && nrow(sample_results) > 0) {
  
  standards_plot_data <- data_long %>%
    dplyr::filter(!is.na(concentration), SampleType == "Standard") %>%
    # Task 3: pre-compute hover text for the standards layer so plotly's
    # tooltip="text" picks it up. Inline aes(text = paste0(...)) would also
    # work but a mutate keeps the response-column dynamic lookup
    # (.data[[response_var]]) outside the aesthetic mapping.
    dplyr::mutate(
      .response_for_hover = .data[[response_var]],
      .hover_text = paste0(
        "Standard: ", SampleID, "<br>",
        "Concentration: ",
        if (isTRUE(is_elisa)) {
          format(concentration, digits = 4, big.mark = ",")
        } else {
          scales::scientific(concentration, digits = 2)
        }, "<br>",
        labels$y_label, ": ", round(.response_for_hover, 2), "<br>",
        "Variability: ", high_variability
      )
    )

  unknown_plot_data <- sample_results %>%
    dplyr::filter(!is.na(concentration_diluted), is.finite(concentration_diluted)) %>%
    dplyr::mutate(
      falls_within_range = mapply(classify_range, response_value, estimated_concentration,
                                  MoreArgs = list(is_elisa_assay = is_elisa))
    )

  if (nrow(unknown_plot_data) > 0) {

    combined_drc <- ggplot() +
      geom_line(data = model_fits, aes(x = conc, y = p),
               color = "black", size = 1.2) +
      geom_point(data = standards_plot_data,
                aes(x = concentration, y = .data[[response_var]],
                    color = high_variability,
                    text = .hover_text),
                size = 3, alpha = 0.7) +
      geom_point(data = unknown_plot_data,
                aes(x = concentration_diluted, y = response_value,
                    shape = falls_within_range,
                    text = paste0(
                      "Sample: ", SampleID, "<br>",
                      "Replicate: ", Replicate, "<br>",
                      "Dilution: 1:", DilutionFactor, "<br>",
                      "Actual Concentration: ",
                      if (isTRUE(is_elisa)) {
                        format(estimated_concentration, digits = 4, big.mark = ",")
                      } else {
                        scales::scientific(estimated_concentration, digits = 2)
                      }, "<br>",
                      "Diluted Concentration: ",
                      if (isTRUE(is_elisa)) {
                        format(concentration_diluted, digits = 4, big.mark = ",")
                      } else {
                        scales::scientific(concentration_diluted, digits = 2)
                      }, "<br>",
                      labels$y_label, ": ", round(response_value, 2), "<br>",
                      if ("concentration_pg_per_g" %in% names(unknown_plot_data) &&
                          !is.na(concentration_pg_per_g)) {
                        paste0("Conc. (pg/g tissue): ",
                               format(concentration_pg_per_g, digits = 4, nsmall = 1, big.mark = ","), "<br>")
                      } else { "" },
                      "Range Status: ", falls_within_range
                    )),
                size = 3, color = "#E69F00") +
      scale_shape_manual(name = "Sample Status",
                        values = c("Within Range" = 16, "Out of Range" = 17, "Unknown" = 18)) +
      # GUI refresh §6.5: keep variability cue on standards (still
      # useful) but adopt the Okabe-Ito hexes verbatim.
      scale_color_manual(name = tr("plot_std_variability", lang),
                        values = c("Normal Variability" = "#0072b2",
                                  "High Variability"   = "#d55e00"),
                        labels = c("Normal Variability" = tr("variability_normal", lang),
                                   "High Variability"   = tr("variability_high", lang))) +
      # GUI refresh §6.5: theme_minimal() with the design's gridline
      # color and a hairline panel border. Replaces the previous
      # theme_classic() + theme() chain with a single declarative
      # block.
      theme_minimal() +
      theme(
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "#e3e6ea", linewidth = 0.4),
        panel.border     = element_rect(color = "#e3e6ea", fill = NA, linewidth = 0.5),
        legend.position  = "top",
        legend.box       = "vertical",
        axis.text.x      = element_text(angle = 45, hjust = 1)
      ) +
      labs(x = labels$x_label, y = labels$y_label)
    
    if (isTRUE(is_elisa)) {
      combined_drc <- combined_drc + 
        scale_x_log10(
          labels = function(x) format(x, digits = 3, big.mark = ","),
          breaks = scales::trans_breaks("log10", function(x) 10^x, n = 6)
        )
    } else {
      combined_drc <- combined_drc + 
        scale_x_log10(
          limits = c(1e-12, 1e-5),
          breaks = 10^seq(-12, -5, by = 1),
          labels = scales::scientific_format(digits = 2)
        )
    }
    
    # Render plot - interactive for HTML, static for DOCX/PDF
    render_plot(combined_drc, tooltip = "text")
  }
}
```

`r section_close()`

## `r tr("parallelism_title", lang)`

```{r parallelism-assessment, results='asis', eval=!is_compact}
# Task C (HP7): Optional parallelism / relative potency assessment
tryCatch({
  if (exists("model_fit") && !is.null(model_fit) && exists("data_long") && exists("response_var")) {
    parallelism_result <- assess_parallelism(data_long, model_fit, response_var)

    if (isTRUE(parallelism_result$applicable)) {
      cat("### Test Results\n\n")

      # Slope comparison
      if (!is.na(parallelism_result$slope_test$p_value)) {
        if (isTRUE(parallelism_result$slope_test$parallel)) {
          cat(sprintf("**Slope comparison:** Slopes are not significantly different (t = %.3f, p = %.4f). Curves appear parallel.\n\n",
                      parallelism_result$slope_test$statistic,
                      parallelism_result$slope_test$p_value))
        } else {
          cat(sprintf("**Slope comparison:** Slopes differ significantly (t = %.3f, p = %.4f). Curves are **not** parallel.\n\n",
                      parallelism_result$slope_test$statistic,
                      parallelism_result$slope_test$p_value))
        }
      }

      # Potency ratio
      if (!is.na(parallelism_result$potency_ratio$estimate)) {
        cat(sprintf("**Relative potency (EC50 ratio):** %.4f [95%% CI: %.4f \u2013 %.4f]\n\n",
                    parallelism_result$potency_ratio$estimate,
                    parallelism_result$potency_ratio$ci_lower,
                    parallelism_result$potency_ratio$ci_upper))
      }

      cat(sprintf("*%d curve groups were compared.*\n\n", parallelism_result$curves_found))

    } else {
      cat(sprintf("*%s*\n\n", tr("parallelism_not_applicable", lang, parallelism_result$reason)))
    }
  } else {
    cat(sprintf("*%s*\n\n", tr("parallelism_no_model", lang)))
  }
}, error = function(e) {
  cat(sprintf("*Parallelism assessment could not be completed: %s*\n\n", e$message))
})
```

## Summary

```{r report-summary, results='asis'}
assay_type_label <- toupper(assay_config$assay_type %||% "Unknown")
analyte_label <- assay_config$analyte %||% assay_config$toxin_standard_label %||% "Unknown"
cat(sprintf("- **Assay:** %s \u2014 %s\n", assay_type_label, analyte_label))

n_std_used <- if (exists("model_stats") && !is.null(model_stats)) model_stats$n_standards %||% NA_integer_ else NA_integer_
n_std_total <- if (exists("model_stats") && !is.null(model_stats)) model_stats$n_unique_concs %||% NA_integer_ else NA_integer_
if (!is.na(n_std_used) && !is.na(n_std_total)) {
  cat(sprintf("- **Standards:** %d used (after exclusions) out of %d\n", n_std_used, n_std_total))
}

weight_info <- if (exists("model_stats") && !is.null(model_stats)) model_stats$weight_method %||% "none" else "none"
cat(sprintf("- **Model:** 4PL logistic with %s weighting\n", weight_info))

if (exists("replicate_stats") && !is.null(replicate_stats) && nrow(replicate_stats) > 0) {
  n_total <- nrow(replicate_stats)
  n_quantified <- if ("within_range" %in% names(replicate_stats)) sum(replicate_stats$within_range == "Within Range", na.rm = TRUE) else NA_integer_
  n_flagged <- if (!is.na(n_quantified)) n_total - n_quantified else NA_integer_
  if (!is.na(n_quantified)) {
    cat(sprintf("- **Samples analyzed:** %d total, %d within range, %d flagged\n", n_total, n_quantified, n_flagged))
  } else {
    cat(sprintf("- **Samples analyzed:** %d total\n", n_total))
  }
}

if (exists("formal_lloq") && exists("formal_uloq") && !is.na(formal_lloq) && !is.na(formal_uloq)) {
  conc_unit <- if (isTRUE(is_elisa)) assay_config$units %||% "pg/mL" else "mol/L"
  lloq_fmt <- if (isTRUE(is_elisa)) format(formal_lloq, digits = 4, big.mark = ",") else format(formal_lloq, scientific = TRUE, digits = 3)
  uloq_fmt <- if (isTRUE(is_elisa)) format(formal_uloq, digits = 4, big.mark = ",") else format(formal_uloq, scientific = TRUE, digits = 3)
  cat(sprintf("- **LLOQ:** %s %s, **ULOQ:** %s %s\n", lloq_fmt, conc_unit, uloq_fmt, conc_unit))
}
```

## Interpretation & Recommendations

```{r report-interpretation, results='asis'}
tryCatch({
  qc_issues <- c()

  # Use assay-specific QC profile for thresholds (M7)
  interp_qc_profile <- get_qc_profile(assay_config$assay_type %||% "rba")

  r2_val <- if (exists("model_stats") && !is.null(model_stats)) model_stats$r_squared %||% NA_real_ else NA_real_
  if (!is.na(r2_val) && is.finite(r2_val) && r2_val < interp_qc_profile$r2_threshold) {
    qc_issues <- c(qc_issues, sprintf("Low R\u00b2 (%.3f < %.2f threshold)", r2_val, interp_qc_profile$r2_threshold))
  }

  hill_val <- if (exists("model_stats") && !is.null(model_stats)) model_stats$hill_slope %||% NA_real_ else NA_real_
  if (!is.na(hill_val) && is.finite(hill_val)) {
    abs_hill <- abs(hill_val)
    abs_range <- sort(abs(interp_qc_profile$hill_slope_range))
    if (abs_hill < abs_range[1] || abs_hill > abs_range[2]) {
      qc_issues <- c(qc_issues, sprintf("Unusual Hill slope (%.2f; expected |slope| in %.1f to %.1f)",
                                          hill_val, abs_range[1], abs_range[2]))
    }
  }

  if (exists("replicate_stats") && !is.null(replicate_stats) && "cv_percent" %in% names(replicate_stats)) {
    max_cv <- max(replicate_stats$cv_percent, na.rm = TRUE)
    if (is.finite(max_cv) && max_cv > interp_qc_profile$cv_limit) {
      qc_issues <- c(qc_issues, sprintf("High sample CV (max %.1f%% > %d%% limit)", max_cv, interp_qc_profile$cv_limit))
    }
  }

  if (length(qc_issues) == 0) {
    cat("All quality criteria met. Results are suitable for reporting.\n\n")
  } else {
    cat(sprintf("**Quality warnings detected:** %s. Review flagged items before use.\n\n",
                paste(qc_issues, collapse = ", ")))
  }

  # M6: Quantification status counts
  if (exists("sample_results") && !is.null(sample_results) && nrow(sample_results) > 0 &&
      "quantification_status" %in% names(sample_results)) {
    # Phase 1.4: explain why two apparently overlapping range indicators exist
    cat(tr("range_indicators_explanation", lang), "\n\n", sep = "")
    status_counts <- table(sample_results$quantification_status)
    cat("**Quantification status breakdown (per well):**\n\n")
    for (s in names(status_counts)) {
      cat(sprintf("- %s: %d\n", s, status_counts[[s]]))
    }
    cat("\n")
    if (any(grepl("Extrapolated", names(status_counts)))) {
      cat("*Note: Extrapolated values are outside the standard curve range and should be interpreted with caution.*\n\n")
    }
  }

  cat(sprintf("\n*QC profile: %s (R\u00b2 >= %.2f, CV <= %d%%, Hill slope [%.1f, %.1f], Recovery [%d%%, %d%%])*\n\n",
              toupper(assay_config$assay_type %||% "rba"),
              interp_qc_profile$r2_threshold,
              interp_qc_profile$cv_limit,
              interp_qc_profile$hill_slope_range[1],
              interp_qc_profile$hill_slope_range[2],
              interp_qc_profile$recovery_range[1],
              interp_qc_profile$recovery_range[2]))
}, error = function(e) {
  cat(sprintf("\n\n*Interpretation section failed: %s*\n\n", e$message))
})
```

```{r tissue-normalization-traceability, results='asis', eval=!is_compact}
# M8: Tissue Normalization Traceability (ELISA + tissue weights only)
render_tissue_normalization_section(
  is_elisa          = isTRUE(is_elisa),
  tissue_weights    = if (exists("tissue_weights")) tissue_weights else NULL,
  processing_config = if (exists("processing_config")) processing_config else list(),
  replicate_stats   = if (exists("replicate_stats")) replicate_stats else NULL,
  lang              = lang
)
```

### Exclusion Audit

`r section_open(tr("exclusion_audit_caption", lang))`

```{r exclusion-audit, results='asis', eval=!is_compact}
# M5: Formal Exclusion Audit Table
render_exclusion_audit_section(
  data_long          = if (exists("data_long")) data_long else NULL,
  outlier_flags      = if (exists("outlier_flags")) outlier_flags else NULL,
  high_var_standards = if (exists("high_var_standards")) high_var_standards else NULL,
  lang               = lang
)
```

`r section_close()`

```{r abbreviations, results='asis'}
# Item 11: collapsed Abbreviations block above the report footer.
render_abbreviations_block()
```

---

**`r tr("report_generated", lang)`** `r Sys.Date()`
**`r tr("contact", lang)`** kr.moeller@iaea.org
**`r tr("feedback", lang)`** [`r tr("online_form", lang)`](https://forms.office.com/e/q8eqJfp4QM)

*`r tr("automated_analysis", lang)`*