---
title: "Bioassay Analysis Report"
author: "Kristof Moeller (IAEA, Monaco) and Arnold Molina Porras (Universidad de Costa Rica)"
output:
  html_document:
    df_print: paged
    self_contained: true
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: flatly
    css: report_style.css
  word_document:
    toc: true
  pdf_document:
    toc: true
header-includes:
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyfoot[C]{\thepage}
  - \fancyhead{}
  - \renewcommand{\headrulewidth}{0pt}
params:
  output_dir: NULL
  lang: "en"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load required libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(drc)
library(scales)
library(knitr)
library(kableExtra)
library(stringr)
library(ggtext)
library(glue)
library(htmltools)
library(tibble)
library(jsonlite)

# %||% operator is defined in global.R (loaded before rendering)

# Validate output directory
output_dir <- normalizePath(params$output_dir, winslash = "/", mustWork = TRUE)

# Load language setting
lang <- params$lang %||% "en"

# Get script directory FIRST (needed before sourcing any files)
script_dir <- tryCatch({
  rmd_path <- knitr::current_input()
  if (!is.null(rmd_path) && rmd_path != "" && rmd_path != ".") {
    normalizePath(dirname(rmd_path), winslash = "/")
  } else {
    if (dir.exists("reports")) {
      normalizePath("reports", winslash = "/")
    } else {
      possible_reports_dir <- file.path(dirname(params$output_dir), "reports")
      if (dir.exists(possible_reports_dir)) {
        normalizePath(possible_reports_dir, winslash = "/")
      } else {
        "."
      }
    }
  }
}, error = function(e) {
  if (dir.exists("reports")) {
    return(normalizePath("reports", winslash = "/"))
  } else {
    "."
  }
})

# Source i18n translations (search multiple locations)
i18n_found <- FALSE
for (i18n_candidate in c(
  file.path(script_dir, "i18n.R"),
  file.path(dirname(script_dir), "i18n.R"),
  "../i18n.R",
  "i18n.R",
  file.path(dirname(output_dir), "i18n.R")
)) {
  if (file.exists(i18n_candidate)) {
    source(i18n_candidate)
    i18n_found <- TRUE
    break
  }
}
if (!i18n_found) {
  tr <- function(key, lang = "en", ...) paste0("[", key, "]")
}

# Check if all required function files exist
required_files <- c("report_constants.R", "report_functions.R", "plot_functions.R")
missing_files <- c()

for (file in required_files) {
  file_path <- file.path(script_dir, file)
  if (!file.exists(file_path)) {
    missing_files <- c(missing_files, file_path)
  }
}

if (length(missing_files) > 0) {
  stop("Missing required function files:\n", paste(missing_files, collapse = "\n"),
       "\n\nExpected in directory: ", script_dir,
       "\nActual files found: ", paste(list.files(script_dir, pattern = "\\.R$"), collapse = ", "),
       "\n\nMake sure all R files are in the 'reports/' folder.")
}

# Source function files
source(file.path(script_dir, "report_constants.R"))
source(file.path(script_dir, "report_functions.R"))
source(file.path(script_dir, "plot_functions.R"))

# Load assay configuration
assay_config <- tryCatch({
  load_assay_config(output_dir)
}, error = function(e) {
  # Fallback to RBA defaults
  list(
    assay_type = "rba",
    detection_method = "radioligand",
    toxin_standard_label = "Saxitoxin"
  )
})

# Load analysis settings
analysis_config <- tryCatch({
  cfg_file <- file.path(output_dir, "analysis_config.json")
  if (file.exists(cfg_file)) {
    jsonlite::fromJSON(cfg_file)
  } else {
    list(regression_weight = "none", quant_range_min = 20, quant_range_max = 80,
         ci_method = "t_dist", enable_outlier_detection = FALSE, outlier_min_n = 3,
         normality_assumption = "assume")
  }
}, error = function(e) {
  list(regression_weight = "none", quant_range_min = 20, quant_range_max = 80,
       ci_method = "t_dist", enable_outlier_detection = FALSE, outlier_min_n = 3,
       normality_assumption = "assume")
})
# Ensure backward compatibility for configs saved without normality_assumption
if (is.null(analysis_config$normality_assumption)) {
  analysis_config$normality_assumption <- "assume"
}
# Resolve configurable CV threshold (app-level override or fallback to global constant)
cv_limit <- analysis_config$cv_limit %||% QC_THRESHOLDS$cv_limit
```

```{r load-data, include=FALSE}
# Load and validate data
data_long_raw <- read.csv(file.path(output_dir, "long_data_output.csv"), stringsAsFactors = FALSE)

# Remove diagnostic control columns that confuse users
diagnostic_cols <- c("ControlValues_B0", "ControlValues_NSB", "ControlValues_Blank")
data_long <- data_long_raw %>%
  dplyr::select(-any_of(diagnostic_cols))

validate_input_data(data_long)

# Detect assay type from data if not specified in config
has_normalized <- "NormalizedValue" %in% names(data_long)
is_elisa <- isTRUE(assay_config$assay_type == "elisa") || has_normalized

# For ELISA, calculate proper B/B0 values following Cayman protocol
if (isTRUE(is_elisa)) {
  # Calculate B/B0 values using control wells (Blank, NSB, B0)
  # Use the raw data with diagnostic columns for calculation, but return clean data
  data_long <- tryCatch({
    data_with_bb0 <- calculate_elisa_bb0(data_long_raw, use_percent = TRUE)  # Use %B/B0 (0-100%)
    # Remove diagnostic columns from the result
    data_with_bb0 %>% dplyr::select(-any_of(diagnostic_cols))
  }, error = function(e) {
    warning("Could not calculate B/B0 values: ", e$message)
    warning("Falling back to existing NormalizedValue column")
    data_long$calculated_bb0 <- data_long$NormalizedValue
    data_long
  })
  
  response_var <- "calculated_bb0"
  
  # Update assay configuration
  if (!isTRUE(assay_config$assay_type == "elisa")) {
    assay_config$assay_type <- "elisa"
    assay_config$units <- "pg/mL"
    assay_config$analyte <- assay_config$analyte %||% "Analyte"
  }
  
  # Display control well summary for QC
  control_summary <- attr(data_long, "control_summary")
  if (!is.null(control_summary)) {
    cat("**Control Well QC:**\n")
    cat("- Blank average:", round(as.numeric(control_summary$blank_avg), 3), "\n")
    cat("- NSB average:", round(as.numeric(control_summary$nsb_avg), 3), "\n")
    cat("- B0 average:", round(as.numeric(control_summary$b0_avg), 3), "\n")
    cat("- Hierarchy valid:", control_summary$hierarchy_valid, "\n\n")
  }
  
} else {
  # RBA: Use existing response variable
  response_var <- if (has_normalized && "NormalizedValue" %in% names(data_long)) {
    "NormalizedValue"
  } else {
    "MeasurementValue"  
  }
}

# Get axis labels
labels <- get_axis_labels(assay_config)

# Load notes (optional)
notes_file <- file.path(output_dir, "notes.json")
notes_text <- "No notes added."
if (file.exists(notes_file)) {
  notes_obj <- jsonlite::fromJSON(notes_file, simplifyVector = FALSE)
  notes_text <- notes_obj$notes %||% "No notes added."
}

# Load sample processing configuration for ELISA
if (isTRUE(is_elisa)) {
  processing_config <- load_sample_processing_config(output_dir)
}
```

# `r tr(if (isTRUE(is_elisa)) "report_elisa" else "report_rba", lang)` `r tr("analysis_report", lang)`

**`r tr("assay_type", lang)`** `r toupper(assay_config$assay_type)`
**`r tr("analysis_date", lang)`** `r Sys.Date()`
**`r tr("analyst", lang)`** `r Sys.info()[["user"]]`

```{r software-version-info, results='asis'}
cat(sprintf("**Software:** Bioassay Analysis App v%s  \n", REPORT_INFO$version))
cat(sprintf("**R version:** %s  \n", R.version.string))
cat(sprintf("**drc package:** v%s  \n\n", as.character(packageVersion("drc"))))
```

```{r dynamic-description, results='asis'}
# Dynamic description based on assay type
if (isTRUE(is_elisa)) {
  cat(tr("elisa_intro", lang, assay_config$analyte %||% "analyte"), "\n\n")
  cat("**", tr("elisa_method", lang), "**\n\n", sep = "")
} else {
  cat(tr("rba_intro", lang, assay_config$toxin_standard_label %||% "biotoxin"), "\n\n")
  cat("**", tr("rba_method", lang, assay_config$detection_method %||% "radioligand"), "**\n\n", sep = "")
}
```

## `r tr("analysis_notes", lang)`

```{r notes-section, results='asis'}
if (nzchar(trimws(notes_text)) && notes_text != "No notes added.") {
  cat(notes_text)
} else {
  cat("*", tr("no_notes", lang), "*", sep = "")
}
```

## `r tr("std_curve_config", lang)`

```{r standards-table}
# Extract and display standard concentrations
standards_data <- data_long %>%
  filter(SampleType == "Standard", !is.na(StandardConc)) %>%
  arrange(desc(StandardConc)) %>%
  distinct(StandardConc)

# Format concentrations based on assay type
if (isTRUE(is_elisa)) {
  # ELISA: Use regular number formatting (no scientific notation)
  standard_table <- data.frame(
    StandardNumber = seq_len(nrow(standards_data)),
    Concentration = format(standards_data$StandardConc, digits = 4, nsmall = 1, big.mark = ",")
  )
  unit_label <- paste0("Concentration (", assay_config$units %||% "pg/mL", ")")
} else {
  # RBA: Use scientific notation  
  standard_table <- data.frame(
    StandardNumber = seq_len(nrow(standards_data)),
    Concentration = format(standards_data$StandardConc, scientific = TRUE, digits = 3)
  )
  unit_label <- "Concentration (mol/L)"
}

render_table(
  standard_table, 
  caption = tr("std_concentrations_table", lang),
  col_names = c("Standard", unit_label)
)
```

## `r tr("methods_title", lang)`

```{r methods-section, results='asis'}
cat(tr("methods_drc_citation", lang))

# Describe the regression weighting used by the primary model
selected_weights <- analysis_config$regression_weight
if (is.null(selected_weights) || length(selected_weights) == 0) selected_weights <- "none"
weight_options_labels <- list(none = "Unweighted", inv_y = "1/Y weighted", inv_y2 = "1/Y\u00B2 weighted")
primary_weight_key <- selected_weights[[1]]
weight_label <- weight_options_labels[[primary_weight_key]] %||% primary_weight_key
cat(sprintf("4-parameter log-logistic model (LL.4) fitted with %s regression. ", weight_label))

if (isTRUE(analysis_config$enable_outlier_detection)) {
  cat(tr("methods_outlier_citation", lang))
}

use_shapiro_methods <- isTRUE(analysis_config$normality_assumption == "test_shapiro")
if (use_shapiro_methods) {
  cat("Non-normality fallback: MAD-based detection (Leys et al., 2013). ")
}
cat("\n\n")
```

## `r tr("drc_analysis", lang)`

```{r model-fitting, include=FALSE}
# Add concentration column and identify high variability standards
data_long <- data_long %>%
  mutate(concentration = ifelse(SampleType == "Standard", StandardConc, NA_real_))

high_var_standards <- identify_high_variability_standards(data_long, cv_limit = cv_limit)

# Mark high variability in data
data_long <- data_long %>%
  mutate(
    high_variability = ifelse(
      SampleType == "Standard" & concentration %in% high_var_standards$StandardConc,
      "High Variability", 
      "Normal Variability"
    )
  )

# Prepare standards for modeling (excluding high variability)
standards_for_model <- data_long %>%
  filter(
    SampleType == "Standard",
    !is.na(concentration),
    high_variability == "Normal Variability",
    !is.na(.data[[response_var]])  # Exclude NA response values
  )

# Build list of weightings to fit
weight_options <- list(
  none = list(weights = NULL, label = "Unweighted"),
  inv_y = list(label = "1/Y weighted"),
  inv_y2 = list(label = "1/Y\u00B2 weighted")
)

# Compute weight vectors for the standard data
resp_vals <- standards_for_model[[response_var]]
weight_options$inv_y$weights <- ifelse(resp_vals > 0, 1 / resp_vals, 1)
weight_options$inv_y2$weights <- ifelse(resp_vals > 0, 1 / (resp_vals^2), 1)

# Defensive check: weight vectors must match filtered standards (BUG_003 guard)
stopifnot(length(weight_options$inv_y$weights) == nrow(standards_for_model))
stopifnot(length(weight_options$inv_y2$weights) == nrow(standards_for_model))

# Which weightings did the user select?
selected_weights <- analysis_config$regression_weight
if (is.null(selected_weights) || length(selected_weights) == 0) selected_weights <- "none"
# Backward compat: if a single string, wrap in vector
if (is.character(selected_weights)) selected_weights <- as.character(selected_weights)

multi_weight_mode <- length(selected_weights) > 1

# Fit ALL selected models and store results
all_models <- list()
for (wt_key in selected_weights) {
  wt_info <- weight_options[[wt_key]]
  if (is.null(wt_info)) next

  # DRC fallback chain: LL.4() -> LL.3() -> log-linear interpolation
  fml <- as.formula(paste(response_var, "~ concentration"))
  fit <- NULL
  fit_method <- "LL.4"

  # Tier 1: Try LL.4() (4-parameter logistic)
  fit <- tryCatch({
    if (isTRUE(is_elisa)) {
      if (is.null(wt_info$weights)) {
        drc::drm(fml, data = standards_for_model, fct = drc::LL.4(),
                 lowerl = c(NA, 0, NA, NA), upperl = c(NA, NA, 100, NA))
      } else {
        drc::drm(fml, data = standards_for_model, fct = drc::LL.4(),
                 weights = wt_info$weights,
                 lowerl = c(NA, 0, NA, NA), upperl = c(NA, NA, 100, NA))
      }
    } else {
      if (is.null(wt_info$weights)) {
        drc::drm(fml, data = standards_for_model, fct = drc::LL.4())
      } else {
        drc::drm(fml, data = standards_for_model, fct = drc::LL.4(),
                 weights = wt_info$weights)
      }
    }
  }, error = function(e) {
    warning("LL.4() failed for weight '", wt_key, "': ", e$message, ". Trying LL.3()...")
    NULL
  })

  # Tier 2: Try LL.3() (3-parameter, fixed bottom = 0)
  if (is.null(fit)) {
    fit_method <- "LL.3"
    fit <- tryCatch({
      if (is.null(wt_info$weights)) {
        drc::drm(fml, data = standards_for_model, fct = drc::LL.3())
      } else {
        drc::drm(fml, data = standards_for_model, fct = drc::LL.3(),
                 weights = wt_info$weights)
      }
    }, error = function(e) {
      warning("LL.3() also failed for weight '", wt_key, "': ", e$message, ". Falling back to log-linear interpolation.")
      NULL
    })
  }

  if (!is.null(fit)) {
    resid <- residuals(fit)
    resp_vals <- standards_for_model[[response_var]]

    # Use weighted R² when weights are applied
    if (!is.null(wt_info$weights)) {
      w <- wt_info$weights
      ss_res <- sum(w * resid^2)
      ss_tot <- sum(w * (resp_vals - weighted.mean(resp_vals, w))^2)
    } else {
      ss_res <- sum(resid^2)
      ss_tot <- sum((resp_vals - mean(resp_vals))^2)
    }

    if (ss_tot > 0) {
      r2 <- 1 - ss_res / ss_tot
    } else {
      r2 <- NA_real_
      warning(tr("r2_identical_warning", lang))
    }
    rmse <- sqrt(mean(resid^2))
    coefs <- coef(fit)

    conv_info <- tryCatch({
      s <- summary(fit)
      se_vals <- s$coefficients[, "Std. Error"]
      list(converged = TRUE, se_available = all(!is.na(se_vals)),
           max_se = max(se_vals, na.rm = TRUE))
    }, error = function(e) list(converged = TRUE, se_available = FALSE, max_se = NA_real_))

    lof_result <- tryCatch({
      lof <- modelFit(fit)
      list(f_stat = lof[2, "F value"], p_value = lof[2, "p value"])
    }, error = function(e) NULL)

    all_models[[wt_key]] <- list(
      model = fit, label = paste0(wt_info$label, " (", fit_method, ")"),
      R2 = r2, RMSE = rmse, coefs = coefs,
      fit_method = fit_method,
      convergence = conv_info,
      lack_of_fit = lof_result
    )
  } else {
    # Tier 3: Log-linear interpolation fallback
    tryCatch({
      std_means <- standards_for_model %>%
        dplyr::group_by(concentration) %>%
        dplyr::summarise(mean_resp = mean(.data[[response_var]], na.rm = TRUE), .groups = "drop") %>%
        dplyr::filter(is.finite(mean_resp), concentration > 0) %>%
        dplyr::arrange(concentration)
      if (nrow(std_means) >= 2) {
        interp_fun <- approxfun(log10(std_means$concentration), std_means$mean_resp, rule = 2)
        warning("Using log-linear interpolation fallback for weight '", wt_key, "'.")
        all_models[[wt_key]] <- list(
          model = NULL,
          interp_fun = interp_fun,
          interp_data = std_means,
          label = paste0(wt_info$label, " (log-linear interpolation)"),
          R2 = NA_real_, RMSE = NA_real_, coefs = NULL,
          fit_method = "interpolation"
        )
      } else {
        warning("Insufficient data for interpolation fallback for weight '", wt_key, "'.")
      }
    }, error = function(e) {
      warning("All fitting methods failed for weight '", wt_key, "': ", e$message)
    })
  }
}

# Primary model = first successfully fitted weighting (used for sample calculation)
drc_failed_completely <- (length(all_models) == 0)
if (drc_failed_completely) {
  warning("No dose-response models could be fitted. Report will show standards data only.")
}
if (!drc_failed_completely) {
  primary_key <- names(all_models)[[1]]
  model_fit <- all_models[[primary_key]]$model
  weight_desc <- all_models[[primary_key]]$label
  R2 <- all_models[[primary_key]]$R2
  RMSE <- all_models[[primary_key]]$RMSE
} else {
  model_fit <- NULL
  weight_desc <- "None (all models failed)"
  R2 <- NA_real_
  RMSE <- NA_real_
}

# Generate predictions for plotting - use appropriate concentration range
if (isTRUE(is_elisa)) {
  conc_min <- min(standards_for_model$concentration, na.rm = TRUE) * 0.1
  conc_max <- max(standards_for_model$concentration, na.rm = TRUE) * 10
  conc_range <- exp(seq(log(conc_min), log(conc_max), length.out = 1000))
} else {
  conc_range <- seq(1e-12, 1e-5, length.out = 1000)
}

model_fits <- data.frame(conc = conc_range)
model_fits$p <- predict(model_fit, newdata = model_fits)

# Calculate EC20 and EC80 immediately after model fitting so they are available
# for within-range classification in all downstream plots and tables.
# For ELISA: classify by %B/B0 response value (20-80% is within range)
# For RBA: classify by concentration (EC20-EC80 range)
if (isTRUE(is_elisa)) {
  bb0_lower <- analysis_config$quant_range_min %||% 20  # Lower bound of reliable %B/B0 range
  bb0_upper <- analysis_config$quant_range_max %||% 80  # Upper bound of reliable %B/B0 range
  # Still compute EC20/EC80 concentrations for reference/reporting
  ec20 <- tryCatch({
    ed_result <- ED(model_fit, respLev = 80, type = "absolute", display = FALSE)
    as.numeric(ed_result[1, 1])
  }, error = function(e) NA_real_)
  ec80 <- tryCatch({
    ed_result <- ED(model_fit, respLev = 20, type = "absolute", display = FALSE)
    as.numeric(ed_result[1, 1])
  }, error = function(e) NA_real_)
  if (!is.na(ec20) && !is.na(ec80) && ec20 > ec80) {
    temp <- ec20; ec20 <- ec80; ec80 <- temp
  }
} else {
  # RBA: use concentration-based EC20/EC80
  ec20 <- tryCatch({
    ed_result <- ED(model_fit, respLev = 80, type = "absolute", display = FALSE)
    as.numeric(ed_result[1, 1])
  }, error = function(e) NA_real_)
  ec80 <- tryCatch({
    ed_result <- ED(model_fit, respLev = 20, type = "absolute", display = FALSE)
    as.numeric(ed_result[1, 1])
  }, error = function(e) NA_real_)
  if (!is.na(ec20) && !is.na(ec80) && ec20 > ec80) {
    temp <- ec20; ec20 <- ec80; ec80 <- temp
  }
}

# Helper function: classify a sample as within range or out of range
# For ELISA: based on %B/B0 response value
# For RBA: based on estimated concentration vs EC20/EC80
classify_range <- function(response_val, conc_val, is_elisa_assay) {
  if (isTRUE(is_elisa_assay)) {
    if (is.na(response_val)) return("Unknown")
    if (response_val >= bb0_lower && response_val <= bb0_upper) return("Within Range")
    return("Out of Range")
  } else {
    if (is.na(conc_val) || is.na(ec20) || is.na(ec80)) return("Unknown")
    if (conc_val >= ec20 && conc_val <= ec80) return("Within Range")
    return("Out of Range")
  }
}

# Helper function: get LLOQ/ULOQ flag for a sample
flag_range <- function(response_val, conc_val, is_elisa_assay) {
  if (isTRUE(is_elisa_assay)) {
    if (is.na(response_val)) return("")
    if (response_val > bb0_upper) return("<LLOQ")
    if (response_val < bb0_lower) return(">ULOQ")
    return("")
  } else {
    if (is.na(conc_val) || is.na(ec20) || is.na(ec80)) return("")
    if (conc_val < ec20) return("<LLOQ")
    if (conc_val > ec80) return(">ULOQ")
    return("")
  }
}
```

```{r exec-summary-box, results='asis'}
# Executive summary box — only shown if model fitting succeeded
if (!drc_failed_completely && exists("R2") && !is.na(R2)) {

  overall_status <- if (R2 >= 0.99) "\U0001F7E2 Excellent"
                    else if (R2 >= 0.95) "\U0001F7E1 Acceptable"
                    else "\U0001F534 Review needed"

  if (knitr::is_html_output()) {
    cat('<div style="background: linear-gradient(135deg, #E8F5E9 0%, #F1F8E9 100%); ',
        'padding: 16px 20px; border-radius: 8px; border-left: 4px solid #4CAF50; ',
        'margin: 15px 0 25px 0; font-size: 14px;">\n\n')
  } else {
    cat("---\n\n")
  }

  cat(sprintf("**Curve fit:** R\u00B2 = %.4f | RMSE = %.3f | %s\n\n", R2, RMSE, overall_status))

  primary_model_entry <- if (!drc_failed_completely) all_models[[primary_key]] else NULL
  if (!is.null(primary_model_entry$convergence) && !primary_model_entry$convergence$se_available) {
    cat("**Warning:** Standard errors could not be computed \u2014 interpret parameters with caution.\n\n")
  }
  if (!is.null(primary_model_entry$lack_of_fit) && !is.null(primary_model_entry$lack_of_fit$p_value) && primary_model_entry$lack_of_fit$p_value < 0.05) {
    cat(sprintf("**Lack-of-fit:** F = %.2f, p = %.4f \u2014 model may not adequately describe the data.\n\n",
                primary_model_entry$lack_of_fit$f_stat, primary_model_entry$lack_of_fit$p_value))
  }

  cat(sprintf("**Standards:** %d wells used (of %d unique concentrations) | Method: %s\n\n",
              nrow(standards_for_model), length(unique(standards_data$StandardConc)), weight_desc))

  if (isTRUE(is_elisa) && exists("control_summary") && !is.null(control_summary)) {
    cat(sprintf("**Controls:** Blank=%.3f | NSB=%.3f | B0=%.3f | Hierarchy: %s\n\n",
                as.numeric(control_summary$blank_avg),
                as.numeric(control_summary$nsb_avg),
                as.numeric(control_summary$b0_avg),
                if (control_summary$hierarchy_valid) "Valid" else "**INVALID**"))
  }

  if (knitr::is_html_output()) {
    cat('</div>\n\n')
  } else {
    cat("---\n\n")
  }
}
```

```{r exec-summary-accuracy, results='asis'}
if (exists("recovery_summary") && !is.null(recovery_summary) && nrow(recovery_summary) > 0) {
  mean_recovery <- mean(recovery_summary$mean_recovery, na.rm = TRUE)
  sd_recovery <- sd(recovery_summary$mean_recovery, na.rm = TRUE)
  n_levels <- sum(!is.na(recovery_summary$mean_recovery))
  cat(sprintf("**Overall accuracy:** Mean recovery %.1f%% \u00b1 %.1f%% across %d standard levels\n\n",
              mean_recovery, sd_recovery, n_levels))
}
```

```{r variability-info, results='asis'}
if (nrow(high_var_standards) > 0) {
  cat("**", tr("high_var_standards", lang), "**\n\n", sep = "")
  
  for (i in seq_len(nrow(high_var_standards))) {
    if (isTRUE(is_elisa)) {
      cat(sprintf("- Concentration %.1f %s: %.1f%% CV\n", 
                  high_var_standards$StandardConc[i], 
                  assay_config$units %||% "pg/mL",
                  high_var_standards$cv[i]))
    } else {
      cat(sprintf("- Concentration %.2e mol/L: %.1f%% CV\n", 
                  high_var_standards$StandardConc[i], 
                  high_var_standards$cv[i]))
    }
  }
  cat("\n")
} else {
  cat(tr("all_std_acceptable", lang), "\n\n")
}
```

```{r qc-traffic-light, results='asis'}
# Traffic-light QC summary card
tryCatch({
  # Gather metrics
  coefs <- coef(model_fit)
  hill_slope <- abs(coefs[1])  # b parameter (Hill slope)

  # Standard back-calculation for recovery
  std_backcalc <- tryCatch({
    pred_conc <- predict(model_fit, newdata = data.frame(concentration = standards_for_model$concentration), type = "response")
    data.frame(
      nominal = standards_for_model$concentration,
      predicted_response = pred_conc,
      observed_response = standards_for_model[[response_var]]
    )
  }, error = function(e) NULL)

  mean_recovery <- NA_real_
  if (!is.null(std_backcalc)) {
    # Back-calculate concentrations from observed responses
    backcalc_conc <- tryCatch({
      sapply(std_backcalc$observed_response, function(resp) {
        tryCatch({
          ed <- ED(model_fit, respLev = resp, type = "absolute", display = FALSE)
          as.numeric(ed[1, 1])
        }, error = function(e) NA_real_)
      })
    }, error = function(e) rep(NA_real_, nrow(std_backcalc)))

    recovery_pct <- (backcalc_conc / std_backcalc$nominal) * 100
    recovery_pct <- recovery_pct[is.finite(recovery_pct) & recovery_pct > 0 & recovery_pct < 500]
    if (length(recovery_pct) > 0) mean_recovery <- mean(recovery_pct, na.rm = TRUE)
  }

  max_cv <- if (exists("replicate_stats") && nrow(replicate_stats) > 0 && "cv_percent" %in% names(replicate_stats)) {
    max(replicate_stats$cv_percent, na.rm = TRUE)
  } else NA_real_

  # Define thresholds and classify
  qc_items <- list(
    list(metric = tr("qc_r2", lang), value = round(R2, 4),
         status = if (R2 >= 0.99) "green" else if (R2 >= 0.95) "amber" else "red"),
    list(metric = tr("qc_hill", lang), value = round(hill_slope, 3),
         status = if (hill_slope >= 0.8 && hill_slope <= 1.5) "green"
                  else if (hill_slope >= 0.5 && hill_slope <= 2.0) "amber" else "red")
  )
  if (!is.na(max_cv) && is.finite(max_cv)) {
    qc_items[[length(qc_items) + 1]] <- list(
      metric = tr("qc_max_cv", lang), value = paste0(round(max_cv, 1), "%"),
      status = if (max_cv <= 15) "green" else if (max_cv <= 30) "amber" else "red")
  }
  if (!is.na(mean_recovery) && is.finite(mean_recovery)) {
    qc_items[[length(qc_items) + 1]] <- list(
      metric = tr("qc_recovery", lang), value = paste0(round(mean_recovery, 1), "%"),
      status = if (mean_recovery >= 90 && mean_recovery <= 110) "green"
               else if (mean_recovery >= 80 && mean_recovery <= 120) "amber" else "red")
  }

  # Build traffic light table
  status_icon <- function(s) switch(s, green = "\U0001F7E2", amber = "\U0001F7E1", red = "\U0001F534", "\u2753")
  status_label <- function(s) switch(s, green = tr("qc_green", lang), amber = tr("qc_amber", lang), red = tr("qc_red", lang), "?")

  qc_df <- data.frame(
    Metric = sapply(qc_items, function(x) x$metric),
    Value = sapply(qc_items, function(x) as.character(x$value)),
    Status = sapply(qc_items, function(x) paste(status_icon(x$status), status_label(x$status))),
    stringsAsFactors = FALSE
  )
  names(qc_df) <- c(tr("qc_metric", lang), tr("qc_value", lang), tr("qc_status", lang))

  cat(sprintf("\n\n### %s\n\n", tr("qc_card_title", lang)))
  cat(sprintf("*%s: %s*\n\n", "Regression", weight_desc))
  render_table(qc_df, caption = tr("qc_card_title", lang))

}, error = function(e) {
  cat("\n\n*QC summary could not be generated.*\n\n")
})
```

```{r weight-comparison, results='asis', fig.width=12, fig.height=7, warning=FALSE, message=FALSE}
# Show comparison of multiple weighting methods if the user selected more than one
if (multi_weight_mode && length(all_models) > 1) {
  cat(sprintf("\n\n### %s\n\n", tr("weight_comparison_title", lang)))
  cat(tr("weight_comparison_desc", lang), "\n\n")

  # --- Comparison table: R², RMSE, parameters (+ stability if available) ---
  comp_rows <- lapply(names(all_models), function(k) {
    m <- all_models[[k]]
    coefs <- m$coefs  # b (Hill), c (Bottom), d (Top), e (IC50)
    row <- data.frame(
      Method = m$label,
      R2 = round(m$R2, 5),
      RMSE = round(m$RMSE, 3),
      Hill = round(abs(coefs[1]), 4),
      Bottom = round(coefs[2], 2),
      Top = round(coefs[3], 2),
      IC50 = if (isTRUE(is_elisa)) format(round(coefs[4], 2), big.mark = ",") else format(coefs[4], scientific = TRUE, digits = 3),
      stringsAsFactors = FALSE
    )
    # Append stability grade if already assessed
    if (!is.null(m$stability)) {
      grade_icon <- function(g) switch(g,
        good = "\U0001F7E2", acceptable = "\U0001F7E1",
        unstable = "\U0001F534", failed = "\u274C", "\u2753")
      row$Stability <- paste(grade_icon(m$stability$stability_grade),
                              tools::toTitleCase(m$stability$stability_grade))
    }
    row
  })
  comp_df <- do.call(rbind, comp_rows)
  col_names <- c(tr("qc_metric", lang), "R\u00B2", "RMSE",
                  "Hill slope", "Bottom", "Top", "IC50")
  if ("Stability" %in% names(comp_df)) {
    col_names <- c(col_names, "Stability")
  }
  names(comp_df) <- col_names

  render_table(comp_df, caption = tr("weight_comparison_title", lang))

  # --- Overlay DRC curves ---
  curve_colors <- c("none" = "#000000", "inv_y" = "#0072B2", "inv_y2" = "#D55E00")
  curve_data_list <- list()
  for (k in names(all_models)) {
    pred_df <- data.frame(conc = conc_range)
    pred_df$p <- predict(all_models[[k]]$model, newdata = pred_df)
    pred_df$method <- all_models[[k]]$label
    curve_data_list[[k]] <- pred_df
  }
  curve_data <- do.call(rbind, curve_data_list)

  standards_plot <- data_long %>%
    dplyr::filter(!is.na(concentration), SampleType == "Standard")

  cmp_plot <- ggplot() +
    geom_point(data = standards_plot,
               aes(x = concentration, y = .data[[response_var]]),
               size = 3, alpha = 0.6, color = "grey40") +
    geom_line(data = curve_data,
              aes(x = conc, y = p, color = method),
              size = 1.1) +
    scale_color_manual(
      name = "Weighting",
      values = setNames(
        curve_colors[names(all_models)],
        sapply(all_models, function(m) m$label)
      )
    ) +
    theme_classic() +
    theme(legend.position = "top",
          axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = labels$x_label, y = labels$y_label,
         title = tr("weight_comparison_title", lang))

  if (isTRUE(is_elisa)) {
    cmp_plot <- cmp_plot +
      scale_x_log10(
        labels = function(x) format(x, digits = 3, big.mark = ","),
        breaks = scales::trans_breaks("log10", function(x) 10^x, n = 6)
      )
  } else {
    cmp_plot <- cmp_plot +
      scale_x_log10(
        limits = c(1e-12, 1e-5),
        breaks = 10^seq(-12, -5, by = 1),
        labels = scales::scientific_format(digits = 2)
      )
  }

  if (knitr::is_html_output()) {
    print(plotly::ggplotly(cmp_plot))
  } else {
    print(cmp_plot)
  }

  cat(sprintf("\n\n*%s*\n\n",
      tr("weight_comparison_primary", lang, all_models[[primary_key]]$label)))
}

# Simple heteroscedasticity diagnostic for unweighted model
if ("none" %in% names(all_models) && !is.null(all_models[["none"]]$model)) {
  unw_fit <- all_models[["none"]]$model
  unw_resid <- residuals(unw_fit)
  conc_levels <- standards_for_model$StandardConc
  var_by_level <- tapply(unw_resid^2, conc_levels, mean, na.rm = TRUE)
  var_ratio <- max(var_by_level, na.rm = TRUE) / max(min(var_by_level[var_by_level > 0], na.rm = TRUE), 1e-10)
  if (var_ratio > 3) {
    cat(sprintf("\n\n> **Note:** Residual variance ratio across concentration levels = %.1f (>3), suggesting heteroscedasticity. Weighted regression (1/Y or 1/Y²) is recommended.\n\n", var_ratio))
  }
}
```

```{r weighting-suitability, results='asis'}
# Formal heteroscedasticity diagnostic (assess_heteroscedasticity)
tryCatch({
  # Determine which model to assess: prefer the unweighted model, otherwise
  # use the first available model as a reference
  hetero_model <- NULL
  hetero_label <- NULL

  if ("none" %in% names(all_models) && !is.null(all_models[["none"]]$model)) {
    hetero_model <- all_models[["none"]]$model
    hetero_label <- "Unweighted (LL.4/LL.3)"
  } else if (!drc_failed_completely) {
    # No unweighted model was fitted; use primary model as proxy
    hetero_model <- all_models[[primary_key]]$model
    hetero_label <- all_models[[primary_key]]$label
  }

  if (!is.null(hetero_model)) {
    hetero_result <- assess_heteroscedasticity(
      model = hetero_model,
      data_long = standards_for_model,
      response_var = response_var
    )

    cat("\n\n### Weighting Suitability Assessment\n\n")

    cat(sprintf("**Test:** %s\n\n", hetero_result$test_name %||% "N/A"))

    if (!is.na(hetero_result$statistic)) {
      cat(sprintf("**Statistic:** %.3f", hetero_result$statistic))
      if (!is.na(hetero_result$p_value)) {
        cat(sprintf("&emsp;**p-value:** %.4f", hetero_result$p_value))
      }
      cat("\n\n")
    }

    if (!is.na(hetero_result$variance_ratio)) {
      cat(sprintf("**Variance ratio** (max/min across concentration levels): %.1f\n\n",
                  hetero_result$variance_ratio))
    }

    cat(sprintf("**Interpretation:** %s\n\n", hetero_result$interpretation))
    cat(sprintf("**Recommendation:** %s\n\n", hetero_result$recommendation))

    cat("*Note: This diagnostic is advisory only and does not change the selected model.*\n\n")
  }
}, error = function(e) {
  cat(sprintf("\n\n*Weighting suitability assessment could not be performed: %s*\n\n", e$message))
})
```

```{r model-stability-assessment, results='asis'}
# Convergence and stability diagnostics for all fitted models
tryCatch({
  if (!drc_failed_completely && length(all_models) > 0) {

    # Assess stability for each fitted model
    stability_results <- lapply(names(all_models), function(k) {
      assess_model_stability(all_models[[k]], all_models[[k]]$label)
    })
    names(stability_results) <- names(all_models)

    cat("\n\n### Model Stability Assessment\n\n")

    # Build summary table
    grade_icon <- function(g) {
      switch(g,
        good       = "\U0001F7E2 Good",
        acceptable = "\U0001F7E1 Acceptable",
        unstable   = "\U0001F534 Unstable",
        failed     = "\u274C Failed",
        "\u2753 Unknown"
      )
    }

    stability_df <- data.frame(
      Model = sapply(stability_results, function(s) s$model_name),
      Method = sapply(stability_results, function(s) s$fit_method %||% "N/A"),
      Converged = sapply(stability_results, function(s) if (s$converged) "Yes" else "No"),
      SE_Available = sapply(stability_results, function(s) if (s$se_available) "Yes" else "No"),
      Boundary = sapply(stability_results, function(s) if (s$boundary_flag) "Yes" else "No"),
      Grade = sapply(stability_results, function(s) grade_icon(s$stability_grade)),
      stringsAsFactors = FALSE
    )
    names(stability_df) <- c("Model", "Fit Method", "Converged", "SEs Available",
                              "Boundary Hit", "Stability Grade")

    render_table(stability_df, caption = "Model Stability Assessment")

    # Print detailed warnings / suspicious parameters per model
    for (k in names(stability_results)) {
      sr <- stability_results[[k]]
      has_details <- length(sr$suspicious_params) > 0 || length(sr$warnings) > 0

      if (has_details) {
        cat(sprintf("\n**%s:**\n\n", sr$model_name))
        for (w in sr$warnings) {
          cat(sprintf("- %s\n", w))
        }
        for (sp in sr$suspicious_params) {
          cat(sprintf("- %s\n", sp))
        }
        cat("\n")
      }
    }

    # Store stability info for downstream use (e.g., model comparison table)
    # Attach to all_models entries
    for (k in names(stability_results)) {
      all_models[[k]]$stability <- stability_results[[k]]
    }
  }
}, error = function(e) {
  cat(sprintf("\n\n*Model stability assessment could not be performed: %s*\n\n", e$message))
})
```

```{r lloq-uloq, results='asis'}
# Formal LLOQ/ULOQ determination by back-calculated standard accuracy
tryCatch({
  cat(sprintf("\n\n### %s\n\n", tr("lloq_uloq_title", lang)))
  cat(tr("lloq_uloq_desc", lang), "\n\n")

  # Back-calculate each standard point
  backcalc_results <- data.frame(
    nominal_conc = standards_for_model$concentration,
    observed_response = standards_for_model[[response_var]],
    stringsAsFactors = FALSE
  )

  backcalc_results$backcalc_conc <- sapply(seq_len(nrow(backcalc_results)), function(i) {
    tryCatch({
      resp <- backcalc_results$observed_response[i]
      ed <- ED(model_fit, respLev = resp, type = "absolute", display = FALSE)
      as.numeric(ed[1, 1])
    }, error = function(e) NA_real_)
  })

  backcalc_results$recovery <- (backcalc_results$backcalc_conc / backcalc_results$nominal_conc) * 100

  # Aggregate by nominal concentration
  backcalc_summary <- backcalc_results %>%
    group_by(nominal_conc) %>%
    summarise(
      n = n(),
      mean_backcalc = mean(backcalc_conc, na.rm = TRUE),
      mean_recovery = mean(recovery, na.rm = TRUE),
      cv_backcalc = ifelse(n() >= 2, (sd(backcalc_conc, na.rm = TRUE) / mean(backcalc_conc, na.rm = TRUE)) * 100, NA_real_),
      .groups = "drop"
    ) %>%
    filter(is.finite(mean_recovery) & mean_recovery > 0) %>%
    mutate(
      passes = (mean_recovery >= 80 & mean_recovery <= 120) & (is.na(cv_backcalc) | cv_backcalc < 20),
      accuracy = ifelse(passes, tr("qc_green", lang), tr("qc_red", lang))
    )

  if (nrow(backcalc_summary) > 0) {
    # Display back-calculation table
    display_bc <- backcalc_summary %>%
      dplyr::select(nominal_conc, mean_backcalc, mean_recovery, cv_backcalc, accuracy)

    if (isTRUE(is_elisa)) {
      display_bc$nominal_conc <- format(display_bc$nominal_conc, digits = 4, big.mark = ",")
      display_bc$mean_backcalc <- format(display_bc$mean_backcalc, digits = 4, big.mark = ",")
    } else {
      display_bc$nominal_conc <- format(display_bc$nominal_conc, scientific = TRUE, digits = 3)
      display_bc$mean_backcalc <- format(display_bc$mean_backcalc, scientific = TRUE, digits = 3)
    }
    display_bc$mean_recovery <- paste0(round(as.numeric(backcalc_summary$mean_recovery), 1), "%")
    display_bc$cv_backcalc <- ifelse(is.na(backcalc_summary$cv_backcalc), "\u2014",
                                      paste0(round(backcalc_summary$cv_backcalc, 1), "%"))

    units_str <- if (isTRUE(is_elisa)) assay_config$units %||% "pg/mL" else "mol/L"
    names(display_bc) <- c(
      sprintf(tr("col_nominal", lang), units_str),
      sprintf(tr("col_backcalc", lang), units_str),
      tr("col_recovery", lang),
      "CV%",
      tr("col_accuracy", lang)
    )
    render_table(display_bc, caption = tr("backcalc_title", lang))

    # Determine LLOQ and ULOQ from passing standards
    passing_concs <- backcalc_summary %>% filter(passes) %>% pull(nominal_conc)
    if (length(passing_concs) >= 2) {
      formal_lloq <- min(passing_concs)
      formal_uloq <- max(passing_concs)
      cat(sprintf("\n\n**%s:** %s %s\n\n", tr("lloq_label", lang),
                  if (isTRUE(is_elisa)) format(formal_lloq, digits = 4, big.mark = ",") else format(formal_lloq, scientific = TRUE, digits = 3),
                  units_str))
      cat(sprintf("**%s:** %s %s\n\n", tr("uloq_label", lang),
                  if (isTRUE(is_elisa)) format(formal_uloq, digits = 4, big.mark = ",") else format(formal_uloq, scientific = TRUE, digits = 3),
                  units_str))
    } else {
      cat(tr("lloq_uloq_none", lang), "\n\n")
    }
  } else {
    cat(tr("lloq_uloq_none", lang), "\n\n")
  }

}, error = function(e) {
  cat(sprintf("\n\n*LLOQ/ULOQ determination failed: %s*\n\n", e$message))
})
```

```{r dose-response-plots, fig.width=12, fig.height=8, fig.cap="Dose-response curve showing fitted 4-parameter logistic model with observed standard concentrations"}
# Create unified dose-response curve plot
plot_data <- data_long %>% filter(!is.na(concentration))

drc_plot <- ggplot(plot_data, aes(x = concentration, y = .data[[response_var]], color = high_variability)) +
  geom_point(size = 3, aes(text = paste0(
    "Concentration: ", if (isTRUE(is_elisa)) format(concentration, digits = 4, big.mark = ",") else scales::scientific(concentration, digits = 2), "<br>",
    labels$y_label, ": ", round(.data[[response_var]], 2), "<br>",
    "Variability: ", high_variability
  ))) +
  geom_line(data = model_fits, aes(x = conc, y = p), color = "black", size = 1.2, inherit.aes = FALSE) +
  scale_color_manual(
    name = "Standard Quality",
    values = c(
      "Normal Variability" = "#2E86AB",
      "High Variability" = "#E63946"
    )
  ) +
  theme_classic() +
  theme(
    legend.position = "top",
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 11),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    panel.grid.minor = element_line(color = "grey90", size = 0.25),
    panel.grid.major = element_line(color = "grey85", size = 0.5)
  )

# Set appropriate axis scaling based on assay type
if (isTRUE(is_elisa)) {
  # ELISA: Log scale for wide concentration ranges
  conc_range_factor <- max(plot_data$concentration, na.rm = TRUE) / min(plot_data$concentration, na.rm = TRUE)
  drc_plot <- drc_plot + 
    scale_x_log10(
      name = labels$x_label,
      labels = function(x) format(x, digits = 3, big.mark = ","),
      breaks = scales::trans_breaks("log10", function(x) 10^x, n = 6)
    )
} else {
  # RBA: Log scale with scientific notation
  drc_plot <- drc_plot + 
    scale_x_log10(
      name = labels$x_label,
      breaks = scales::trans_breaks("log10", function(x) 10^x),
      labels = scales::trans_format("log10", scales::math_format(10^.x))
    )
}

drc_plot <- drc_plot + 
  ylab(labels$y_label) +
  ggtitle(paste(tr("four_pl_coefficients", lang), "-", toupper(assay_config$assay_type)))

# Display plot
if (knitr::is_html_output()) {
  plotly::ggplotly(drc_plot, tooltip = "text") %>%
    plotly::layout(showlegend = TRUE)
} else {
  print(drc_plot)
}
```

## `r tr("model_parameters", lang)`

```{r model-stats}
# Extract and display model coefficients with correct parameter order
tryCatch({
  
  coefficients <- summary(model_fit)$coefficients
  
  # Check if coefficients have expected structure
  if (is.matrix(coefficients) && nrow(coefficients) >= 4) {
    
    # DRC package LL.4() parameter order: b (Hill slope), c (Lower asymptote), d (Upper asymptote), e (IC50)
    actual_param_names <- c("Hill Slope", "Bottom", "Top", "IC50")
    
    # Ensure we don't exceed available rows
    n_params <- min(length(actual_param_names), nrow(coefficients))
    
    # Create results data frame with correct parameter mapping
    drc_results <- data.frame(
      Parameter = actual_param_names[1:n_params],
      Value = coefficients[1:n_params, "Estimate"],
      StdError = if ("Std. Error" %in% colnames(coefficients)) {
        coefficients[1:n_params, "Std. Error"]
      } else {
        rep(NA, n_params)
      },
      PValue = if ("Pr(>|t|)" %in% colnames(coefficients)) {
        coefficients[1:n_params, "Pr(>|t|)"]
      } else {
        rep(NA, n_params)
      }
    )
    
    render_table(
      drc_results,
      caption = tr("four_pl_coefficients", lang),
      col_names = c(tr("col_parameter", lang), tr("col_estimate", lang), tr("col_std_error", lang), tr("col_pvalue", lang))
    )
    
  } else {
    cat("Model coefficients structure unexpected. Available coefficients:\n")
    print(coefficients)
  }
  
}, error = function(e) {
  cat("Error extracting model coefficients:", e$message, "\n")
  cat("Model summary:\n")
  print(summary(model_fit))
})
```

**`r tr("model_fit_stats", lang)`**

- **R²** = `r round(R2, 3)`
- **RMSE** = `r round(RMSE, 3)` `r labels$y_unit`
- **`r tr("standards_used", lang)`** = `r nrow(standards_for_model)` / `r length(unique(standards_data$StandardConc))`

```{r save-model-stats, include=FALSE}
# Save model statistics for multiwavelength executive summary
model_stats <- list(
  r_squared = R2,
  rmse = RMSE,
  ic50 = coef(model_fit)[4],
  hill_slope = coef(model_fit)[1],
  n_standards = nrow(standards_for_model),
  n_unique_concs = length(unique(standards_data$StandardConc)),
  weight_method = weight_desc
)

# Note: mean_sample_cv is computed later in save-sample-stats chunk
# after replicate_stats is available (BUG_008 fix)

# write_json_safe is defined in global.R (loaded before rendering)
write_json_safe(model_stats, file.path(output_dir, "model_stats.json"))
```

### `r tr("std_backcalc_title", lang)`

<!-- Back-calculation promoted to visible section (critical QC data) -->

```{r standard-backcalculation, results='asis'}
# Back-calculate concentrations for each standard well using the fitted model
# and compute % recovery vs nominal concentration
standards_backcalc <- data_long %>%
  dplyr::filter(SampleType == "Standard", !is.na(concentration), !is.na(.data[[response_var]])) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    backcalc_conc = tryCatch({
      pred <- ED(model_fit, respLev = .data[[response_var]], type = "absolute", display = FALSE)
      as.numeric(pred[1, 1])
    }, error = function(e) NA_real_),
    recovery_pct = if (!is.na(backcalc_conc) && concentration > 0) {
      (backcalc_conc / concentration) * 100
    } else {
      NA_real_
    }
  ) %>%
  dplyr::ungroup()

# Summarise by standard concentration level (mean of replicates)
recovery_summary <- standards_backcalc %>%
  dplyr::group_by(concentration) %>%
  dplyr::summarise(
    n = n(),
    nominal = first(concentration),
    mean_backcalc = mean(backcalc_conc, na.rm = TRUE),
    sd_backcalc = sd(backcalc_conc, na.rm = TRUE),
    mean_recovery = mean(recovery_pct, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  dplyr::arrange(desc(nominal))

# Format for display
if (isTRUE(is_elisa)) {
  recovery_display <- recovery_summary %>%
    dplyr::mutate(
      `Nominal Conc.` = format(nominal, digits = 4, nsmall = 1, big.mark = ","),
      `Back-Calculated` = format(mean_backcalc, digits = 4, nsmall = 1, big.mark = ","),
      `SD` = ifelse(is.na(sd_backcalc), "—", format(sd_backcalc, digits = 3, nsmall = 1)),
      `Recovery (%)` = ifelse(is.na(mean_recovery), "—", format(mean_recovery, digits = 4, nsmall = 1))
    ) %>%
    dplyr::select(`Nominal Conc.`, n, `Back-Calculated`, SD, `Recovery (%)`)
} else {
  recovery_display <- recovery_summary %>%
    dplyr::mutate(
      `Nominal Conc.` = format(nominal, scientific = TRUE, digits = 3),
      `Back-Calculated` = format(mean_backcalc, scientific = TRUE, digits = 3),
      `SD` = ifelse(is.na(sd_backcalc), "—", format(sd_backcalc, scientific = TRUE, digits = 2)),
      `Recovery (%)` = ifelse(is.na(mean_recovery), "—", format(mean_recovery, digits = 4, nsmall = 1))
    ) %>%
    dplyr::select(`Nominal Conc.`, n, `Back-Calculated`, SD, `Recovery (%)`)
}

conc_unit <- if (isTRUE(is_elisa)) paste0(assay_config$units %||% "pg/mL") else "mol/L"
names(recovery_display) <- c(paste0("Nominal (", conc_unit, ")"), "n",
                              paste0("Back-Calc. (", conc_unit, ")"), 
                              paste0("SD (", conc_unit, ")"),
                              "Recovery (%)")

render_table(recovery_display,
            caption = tr("std_backcalc_caption", lang))

# Report overall recovery statistics
mean_overall_recovery <- mean(recovery_summary$mean_recovery, na.rm = TRUE)
cat(sprintf("\n**%s**\n", tr("overall_recovery", lang, mean_overall_recovery)))
if (mean_overall_recovery >= 80 && mean_overall_recovery <= 120) {
  cat(tr("recovery_acceptable", lang), "\n\n")
} else {
  cat(tr("recovery_outside", lang), "\n\n")
}
```

## `r tr("sample_results", lang)`

```{r data-quality-overview, results='asis'}
total_wells <- nrow(data_long)
std_wells <- sum(data_long$SampleType == "Standard", na.rm = TRUE)
sample_wells <- sum(data_long$SampleType == "Sample", na.rm = TRUE)
blank_wells <- sum(data_long$SampleType %in% c("Blank", "NSB", "B0"), na.rm = TRUE)
excluded_wells <- sum(is.na(data_long$MeasurementValue) | !is.finite(data_long$MeasurementValue))

cat('<div style="background: #E3F2FD; padding: 12px; border-radius: 8px; border-left: 4px solid #1976D2; margin: 15px 0;">\n')
cat("**Data Quality Overview**  \n")
cat(sprintf("Total wells: %d | Standards: %d | Samples: %d | Controls: %d  \n",
            total_wells, std_wells, sample_wells, blank_wells))
if (excluded_wells > 0) {
  cat(sprintf("**Excluded wells:** %d (missing or non-finite values)  \n", excluded_wells))
}
if (exists("outlier_flags") && nrow(outlier_flags) > 0) {
  cat(sprintf("**Outliers flagged:** %d  \n", nrow(outlier_flags)))
}
cat('</div>\n\n')
```

```{r sample-analysis, include=FALSE}
# Only analyze actual samples (exclude standards and control wells)
samples_to_analyze <- data_long %>%
  filter(
    SampleType == "Sample",  # Only samples, not Standards, Blanks, NSB, TA, etc.
    !is.na(.data[[response_var]]),
    is.finite(.data[[response_var]])
  )

if (nrow(samples_to_analyze) > 0) {
  
  # Use DRC's ED function to get predictions with proper confidence intervals
  # This accounts for model uncertainty via the delta method
  
  # For each sample, predict concentration with CI using DRC
  sample_predictions_list <- lapply(seq_len(nrow(samples_to_analyze)), function(i) {
    response_value <- samples_to_analyze[[response_var]][i]
    dilution <- samples_to_analyze$DilutionFactor[i]
    
    tryCatch({
      # Use ED function to get inverse prediction with confidence interval
      # The "delta" method properly propagates model uncertainty
      pred_result <- ED(model_fit, respLev = response_value, type = "absolute", 
                       interval = "delta", level = 0.95, display = FALSE)
      
      # Extract estimates and CI (these are at the diluted level - as measured)
      estimate_diluted <- as.numeric(pred_result[1, 1])
      ci_lower_diluted <- as.numeric(pred_result[1, 3])
      ci_upper_diluted <- as.numeric(pred_result[1, 4])
      
      # Apply dilution correction to get actual sample concentration
      data.frame(
        Well = samples_to_analyze$Well[i],
        SampleID = samples_to_analyze$SampleID[i],
        SampleType = samples_to_analyze$SampleType[i],
        Replicate = samples_to_analyze$Replicate[i],
        DilutionFactor = dilution,
        response_value = response_value,
        # For plotting on the curve (diluted level)
        concentration_diluted = estimate_diluted,
        # For reporting (actual sample concentration)
        estimated_concentration = estimate_diluted / dilution,
        ci_lower_individual = ci_lower_diluted / dilution,
        ci_upper_individual = ci_upper_diluted / dilution,
        stringsAsFactors = FALSE
      )
    }, error = function(e) {
      # If prediction fails, return NA
      data.frame(
        Well = samples_to_analyze$Well[i],
        SampleID = samples_to_analyze$SampleID[i],
        SampleType = samples_to_analyze$SampleType[i],
        Replicate = samples_to_analyze$Replicate[i],
        DilutionFactor = dilution,
        response_value = response_value,
        concentration_diluted = NA_real_,
        estimated_concentration = NA_real_,
        ci_lower_individual = NA_real_,
        ci_upper_individual = NA_real_,
        stringsAsFactors = FALSE
      )
    })
  })
  
  # Combine all predictions
  sample_results <- do.call(rbind, sample_predictions_list)
  
  # Outlier detection (applied before stats if enabled)
  outlier_flags <- data.frame(Well = character(), Replicate = character(),
                               is_outlier = logical(), method = character(),
                               stringsAsFactors = FALSE)

  if (isTRUE(analysis_config$enable_outlier_detection)) {
    outlier_min <- analysis_config$outlier_min_n %||% 3
    use_shapiro <- isTRUE(analysis_config$normality_assumption == "test_shapiro")
    outlier_method_log <- data.frame(Replicate = character(), method_used = character(),
                                      shapiro_p = numeric(), stringsAsFactors = FALSE)

    outlier_groups <- sample_results %>%
      filter(!is.na(estimated_concentration)) %>%
      group_by(Replicate) %>%
      filter(n() >= outlier_min) %>%
      ungroup()

    if (nrow(outlier_groups) > 0) {
      for (rep_grp in unique(outlier_groups$Replicate)) {
        grp_data <- outlier_groups %>% filter(Replicate == rep_grp)
        n_grp <- nrow(grp_data)
        vals <- grp_data$estimated_concentration

        # Determine if MAD-based detection should be used (Shapiro-Wilk pre-test)
        use_mad <- FALSE
        shapiro_p <- NA_real_
        if (use_shapiro && n_grp >= 3) {
          shap_test <- tryCatch(shapiro.test(vals), error = function(e) list(p.value = 1))
          shapiro_p <- shap_test$p.value
          if (shapiro_p < 0.05) {
            use_mad <- TRUE
          }
        }

        if (use_mad) {
          # MAD-based outlier detection (robust, non-parametric)
          med_val <- median(vals)
          mad_val <- mad(vals, constant = 1.4826)
          if (mad_val > 0) {
            mad_scores <- abs(vals - med_val) / mad_val
            outlier_idx <- which(mad_scores > 3)
            if (length(outlier_idx) > 0) {
              for (oi in outlier_idx) {
                outlier_flags <- rbind(outlier_flags, data.frame(
                  Well = grp_data$Well[oi], Replicate = rep_grp,
                  is_outlier = TRUE, method = "MAD", stringsAsFactors = FALSE))
              }
            }
          }
          if (use_shapiro) {
            outlier_method_log <- rbind(outlier_method_log, data.frame(
              Replicate = rep_grp, method_used = "MAD",
              shapiro_p = shapiro_p, stringsAsFactors = FALSE))
          }
        } else if (n_grp >= 3 && n_grp <= 5) {
          # Dixon's Q-test
          sorted_vals <- sort(vals)
          range_val <- max(sorted_vals) - min(sorted_vals)
          if (range_val > 0) {
            q_low <- (sorted_vals[2] - sorted_vals[1]) / range_val
            q_high <- (sorted_vals[n_grp] - sorted_vals[n_grp - 1]) / range_val
            # Critical values for Dixon's Q (alpha=0.05)
            q_crit <- c(`3` = 0.970, `4` = 0.829, `5` = 0.710)
            crit <- q_crit[as.character(n_grp)]
            if (q_low > crit) {
              outlier_flags <- rbind(outlier_flags, data.frame(
                Well = grp_data$Well[which.min(vals)], Replicate = rep_grp,
                is_outlier = TRUE, method = "Dixon Q", stringsAsFactors = FALSE))
            }
            if (q_high > crit) {
              outlier_flags <- rbind(outlier_flags, data.frame(
                Well = grp_data$Well[which.max(vals)], Replicate = rep_grp,
                is_outlier = TRUE, method = "Dixon Q", stringsAsFactors = FALSE))
            }
          }
          if (use_shapiro) {
            outlier_method_log <- rbind(outlier_method_log, data.frame(
              Replicate = rep_grp, method_used = "Dixon Q",
              shapiro_p = shapiro_p, stringsAsFactors = FALSE))
          }
        } else if (n_grp >= 6) {
          # Grubbs' test (simplified)
          mean_val <- mean(vals)
          sd_val <- sd(vals)
          if (sd_val > 0) {
            g_stats <- abs(vals - mean_val) / sd_val
            # Approximate Grubbs critical value using t-distribution
            t_crit <- qt(1 - 0.05 / (2 * n_grp), n_grp - 2)
            g_crit <- ((n_grp - 1) / sqrt(n_grp)) * sqrt(t_crit^2 / (n_grp - 2 + t_crit^2))
            outlier_idx <- which(g_stats > g_crit)
            if (length(outlier_idx) > 0) {
              for (oi in outlier_idx) {
                outlier_flags <- rbind(outlier_flags, data.frame(
                  Well = grp_data$Well[oi], Replicate = rep_grp,
                  is_outlier = TRUE, method = "Grubbs", stringsAsFactors = FALSE))
              }
            }
          }
          if (use_shapiro) {
            outlier_method_log <- rbind(outlier_method_log, data.frame(
              Replicate = rep_grp, method_used = "Grubbs",
              shapiro_p = shapiro_p, stringsAsFactors = FALSE))
          }
        }
      }
    }
  }

  # Mark outliers in sample_results
  if (nrow(outlier_flags) > 0) {
    sample_results$is_outlier <- sample_results$Well %in% outlier_flags$Well
  } else {
    sample_results$is_outlier <- FALSE
  }

  # Calculate replicate group statistics (excluding flagged outliers from mean/CI)
  stats_data <- sample_results %>%
    filter(!is.na(estimated_concentration) & !is_outlier)

  replicate_stats <- stats_data %>%
    group_by(Replicate) %>%
    summarise(
      sampleID = paste(unique(SampleID), collapse = ","),
      sample_type = first(SampleType),
      n_replicates = n(),
      mean_conc = mean(estimated_concentration, na.rm = TRUE),
      sd_conc = sd(estimated_concentration, na.rm = TRUE),
      se_conc = sd_conc / sqrt(n_replicates),
      # Confidence intervals - method depends on config, combined with model uncertainty
      ci_lower = {
        # Replicate-variability CI
        ci_replicate_lower <- if (isTRUE(analysis_config$ci_method == "bootstrap") && n_replicates >= 3) {
          set.seed(42)
          boot_vals <- replicate(1000, mean(sample(estimated_concentration, replace = TRUE)))
          set.seed(NULL)
          quantile(boot_vals, 0.025)
        } else if (n_replicates >= 2 && !is.na(sd_conc)) {
          mean_conc - qt(0.975, df = n_replicates - 1) * se_conc
        } else {
          mean_conc
        }
        # Model-uncertainty envelope from per-well delta-method CIs
        ci_model_lower <- if (any(!is.na(ci_lower_individual))) {
          min(ci_lower_individual, na.rm = TRUE)
        } else {
          ci_replicate_lower
        }
        # Use the wider (more conservative) of the two lower bounds
        min(ci_replicate_lower, ci_model_lower)
      },
      ci_upper = {
        # Replicate-variability CI
        ci_replicate_upper <- if (isTRUE(analysis_config$ci_method == "bootstrap") && n_replicates >= 3) {
          set.seed(42)
          boot_vals <- replicate(1000, mean(sample(estimated_concentration, replace = TRUE)))
          set.seed(NULL)
          quantile(boot_vals, 0.975)
        } else if (n_replicates >= 2 && !is.na(sd_conc)) {
          mean_conc + qt(0.975, df = n_replicates - 1) * se_conc
        } else {
          mean_conc
        }
        # Model-uncertainty envelope from per-well delta-method CIs
        ci_model_upper <- if (any(!is.na(ci_upper_individual))) {
          max(ci_upper_individual, na.rm = TRUE)
        } else {
          ci_replicate_upper
        }
        # Use the wider (more conservative) of the two upper bounds
        max(ci_replicate_upper, ci_model_upper)
      },
      cv_percent = (sd_conc / mean_conc) * 100,
      .groups = "drop"
    ) %>%
    mutate(
      ci_lower = pmax(ci_lower, 0),
      mean_conc_formatted = if (isTRUE(is_elisa)) {
        format(mean_conc, digits = 4, nsmall = 1, big.mark = ",")
      } else {
        format(mean_conc, scientific = TRUE, digits = 3)
      },
      ci_formatted = if (isTRUE(is_elisa)) {
        paste0("[", format(ci_lower, digits = 3, nsmall = 1), " - ",
               format(ci_upper, digits = 3, nsmall = 1), "]")
      } else {
        paste0("[", format(ci_lower, scientific = TRUE, digits = 2), " - ",
               format(ci_upper, scientific = TRUE, digits = 2), "]")
      },
      cv_formatted = format(cv_percent, digits = 3, nsmall = 1)
    )
  
  # Add within-range classification and LLOQ/ULOQ flags per replicate group
  # by looking at the mean response value of each group
  replicate_range <- sample_results %>%
    filter(!is.na(estimated_concentration)) %>%
    group_by(Replicate) %>%
    summarise(
      mean_response = mean(response_value, na.rm = TRUE),
      mean_conc_for_flag = mean(estimated_concentration, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(
      within_range = mapply(classify_range, mean_response, mean_conc_for_flag,
                            MoreArgs = list(is_elisa_assay = is_elisa)),
      range_flag = mapply(flag_range, mean_response, mean_conc_for_flag,
                          MoreArgs = list(is_elisa_assay = is_elisa))
    )
  
  replicate_stats <- replicate_stats %>%
    left_join(replicate_range %>% dplyr::select(Replicate, mean_response, within_range, range_flag), by = "Replicate")
  
  # Add tissue-weight-based concentration if applicable (ELISA only)
  if (isTRUE(is_elisa)) {
    # Load processing config for extraction volume
    processing_config <- tryCatch({
      load_sample_processing_config(output_dir)
    }, error = function(e) {
      list(extraction_volume_ul = 500)
    })
    
    # Load tissue weights if available
    tissue_weights_file <- file.path(output_dir, "tissue_weights.json")
    tissue_weights <- NULL
    if (file.exists(tissue_weights_file)) {
      tissue_weights <- tryCatch({
        jsonlite::fromJSON(tissue_weights_file, simplifyVector = FALSE)
      }, error = function(e) NULL)
    }
    
    if (!is.null(tissue_weights) && length(tissue_weights) > 0) {
      # Parse tissue weights — supports both old format (flat numeric)
      # and new format (list with weight + extraction_uL per group)
      tw_df <- data.frame(
        Replicate = names(tissue_weights),
        stringsAsFactors = FALSE
      )
      tw_df$TissueWeight_mg <- sapply(tissue_weights, function(x) {
        if (is.list(x)) as.numeric(x$weight) else as.numeric(x)
      })
      tw_df$Extraction_uL <- sapply(tissue_weights, function(x) {
        if (is.list(x) && !is.null(x$extraction_uL)) as.numeric(x$extraction_uL)
        else processing_config$extraction_volume_ul %||% 500
      })
      tw_df <- tw_df %>% filter(!is.na(TissueWeight_mg) & TissueWeight_mg > 0)

      if (nrow(tw_df) > 0) {
        replicate_stats <- replicate_stats %>%
          left_join(tw_df, by = "Replicate") %>%
          mutate(
            effective_ext_uL = ifelse(!is.na(Extraction_uL), Extraction_uL,
                                      processing_config$extraction_volume_ul %||% 500),
            total_amount_pg = mean_conc * (effective_ext_uL / 1000),
            concentration_pg_per_g = ifelse(
              !is.na(TissueWeight_mg) & TissueWeight_mg > 0,
              total_amount_pg / (TissueWeight_mg / 1000),
              NA_real_
            )
          )

        # Also add to sample_results for individual well reporting
        sample_results <- sample_results %>%
          left_join(tw_df, by = "Replicate") %>%
          mutate(
            effective_ext_uL = ifelse(!is.na(Extraction_uL), Extraction_uL,
                                      processing_config$extraction_volume_ul %||% 500),
            total_amount_pg = estimated_concentration * (effective_ext_uL / 1000),
            concentration_pg_per_g = ifelse(
              !is.na(TissueWeight_mg) & TissueWeight_mg > 0,
              total_amount_pg / (TissueWeight_mg / 1000),
              NA_real_
            )
          )
      }
    }
  }
  
  # Save detailed results (cleaned up columns)
  sample_results_clean <- sample_results %>%
    dplyr::rename(
      concentration = estimated_concentration
    )
  
  # Save replicate summary
  replicate_summary <- replicate_stats %>%
    dplyr::select(
      Replicate, sampleID, sample_type, n_replicates,
      mean_conc, sd_conc, se_conc, ci_lower, ci_upper, cv_percent,
      any_of(c("mean_response", "within_range", "range_flag", "TissueWeight_mg", "concentration_pg_per_g"))
    ) %>%
    dplyr::rename(
      replicate_group = Replicate,
      sample_ids = sampleID,
      mean_concentration = mean_conc,
      std_dev = sd_conc,
      std_error = se_conc,
      ci_lower_95 = ci_lower,
      ci_upper_95 = ci_upper,
      cv_percent = cv_percent
    )
    
} else {
  sample_results <- data.frame()
  replicate_stats <- data.frame()
  sample_results_clean <- data.frame()
  replicate_summary <- data.frame()
}
```

```{r sample-results-table, results='asis'}
# Wrap entire chunk in tryCatch for debugging
tryCatch({
  
  if (exists("replicate_stats") && !is.null(replicate_stats) && nrow(replicate_stats) > 0) {
    
    # Check if tissue weights were provided (robust to NA)
    has_tissue_weights <- FALSE
    if (exists("sample_results") && !is.null(sample_results) && "TissueWeight_mg" %in% names(sample_results)) {
      tissue_check <- tryCatch({
        any(!is.na(sample_results$TissueWeight_mg) & sample_results$TissueWeight_mg > 0, na.rm = TRUE)
      }, error = function(e) FALSE)
      has_tissue_weights <- isTRUE(tissue_check)
    }
    
    # Determine concentration column name based on assay type
    conc_col_name <- if (isTRUE(is_elisa)) "Mean Conc. (pg/mL)" else "Mean Conc. (mol/L)"
    
    # Create clean results table
    results_display <- replicate_stats %>%
      dplyr::select(Replicate, sampleID, n_replicates, mean_conc_formatted, ci_formatted, cv_formatted)
    
    # Rename columns
    names(results_display) <- c("Replicate Group", "Sample ID", "n", conc_col_name, "95% CI", "CV [%]")
    
    # Add tissue-based concentrations if available
    if (isTRUE(is_elisa) && isTRUE(has_tissue_weights)) {
      tissue_results <- tryCatch({
        sample_results %>%
          dplyr::filter(!is.na(TissueWeight_mg) & TissueWeight_mg > 0) %>%
          dplyr::group_by(Replicate) %>%
          dplyr::summarise(
            tissue_weight = dplyr::first(TissueWeight_mg),
            mean_tissue_conc = mean(concentration_ng_per_g, na.rm = TRUE),
            .groups = "drop"
          ) %>%
          dplyr::mutate(
            tissue_weight_formatted = format(tissue_weight, digits = 4, nsmall = 1),
            tissue_conc_formatted = format(mean_tissue_conc, digits = 4, nsmall = 2, big.mark = ",")
          )
      }, error = function(e) {
        warning("Could not process tissue weight data: ", e$message)
        NULL
      })
      
      if (!is.null(tissue_results) && nrow(tissue_results) > 0) {
        results_display <- results_display %>%
          dplyr::left_join(
            tissue_results %>% dplyr::select(Replicate, tissue_weight_formatted, tissue_conc_formatted), 
            by = c("Replicate Group" = "Replicate")
          )
        
        # Rename tissue columns if they exist
        if ("tissue_weight_formatted" %in% names(results_display)) {
          names(results_display)[names(results_display) == "tissue_weight_formatted"] <- "Tissue Mass [mg]"
        }
        if ("tissue_conc_formatted" %in% names(results_display)) {
          names(results_display)[names(results_display) == "tissue_conc_formatted"] <- "Conc. (ng/g tissue)"
        }
      }
    }
    
    # Generate table caption
    caption_text <- tr("sample_results_caption", lang, toupper(assay_config$assay_type))
    if (isTRUE(is_elisa) && isTRUE(has_tissue_weights)) {
      caption_text <- paste(caption_text, tr("with_tissue", lang))
    }
    
    render_table(results_display, caption = caption_text)
    
    # Save cleaned results files
    if (exists("sample_results_clean") && !is.null(sample_results_clean)) {
      write.csv(sample_results_clean, 
               file.path(output_dir, "unknown_results.csv"),
               row.names = FALSE)
    }
    
    if (exists("replicate_summary") && !is.null(replicate_summary)) {
      write.csv(replicate_summary,
               file.path(output_dir, "unknown_results_summary.csv"), 
               row.names = FALSE)
    }
    
    cat("\n**", tr("output_files_created", lang), "**\n", sep = "")
    cat("- ", tr("individual_results", lang), "\n", sep = "")
    cat("- ", tr("summary_results", lang), "\n\n", sep = "")
    
    # Display quality assessment (robust to NA in cv_percent)
    high_cv_groups <- tryCatch({
      replicate_stats %>% 
        dplyr::filter(!is.na(cv_percent) & cv_percent > 30)
    }, error = function(e) {
      data.frame()
    })
    
    if (!is.null(high_cv_groups) && nrow(high_cv_groups) > 0) {
      cat("**", tr("quality_alert", lang), "**\n", sep = "")
      cat("- ", tr("high_cv_groups", lang, paste(high_cv_groups$Replicate, collapse = ", ")), "\n", sep = "")
      cat("- ", tr("check_preparation", lang), "\n\n", sep = "")
    } else {
      cat("**", tr("quality_pass", lang), "**\n\n", sep = "")
    }
    
  } else {
    cat("*", tr("no_samples_quantified", lang), "*\n\n", sep = "")
    
    if (exists("samples_to_analyze") && !is.null(samples_to_analyze) && nrow(samples_to_analyze) > 0) {
      cat("**Possible reasons:**\n")
      cat("- Response values outside the standard curve range\n")
      cat("- Model fitting issues\n")
      if (isTRUE(is_elisa)) {
        cat("- Invalid %B/B0 normalization\n")
        cat("- Control well hierarchy violations\n")
      }
      cat("- Mathematical prediction errors\n")
    } else {
      cat("**No sample data found for analysis.**\n")
      cat("Note: Only wells marked as 'Sample' are quantified.\n")
      cat("Blanks, NSB, TA, and Standards are excluded from quantification.\n")
    }
  }
  
}, error = function(e) {
  cat("\n**Error in sample results table:**\n")
  cat("Error message:", e$message, "\n")
  cat("This section will be skipped.\n\n")
})
```

```{r outlier-results, results='asis'}
if (isTRUE(analysis_config$enable_outlier_detection) && exists("outlier_flags") && nrow(outlier_flags) > 0) {
  cat(sprintf("\n\n### %s\n\n", tr("outlier_title", lang)))
  cat(tr("outlier_desc", lang, analysis_config$outlier_min_n %||% 3), "\n\n")

  n_outliers <- nrow(outlier_flags)
  n_groups <- length(unique(outlier_flags$Replicate))
  cat(tr("outlier_found", lang, n_outliers, n_groups), "\n\n")

  outlier_display <- outlier_flags %>%
    dplyr::select(Well, Replicate, method)
  names(outlier_display) <- c("Well", "Replicate Group", "Method")

  render_table(outlier_display, caption = tr("outlier_title", lang))
} else if (isTRUE(analysis_config$enable_outlier_detection)) {
  cat(sprintf("\n\n### %s\n\n", tr("outlier_title", lang)))
  cat(tr("outlier_none", lang), "\n\n")
}
```

```{r outlier-method-log, results='asis'}
if (exists("outlier_method_log") && nrow(outlier_method_log) > 0) {
  cat("\n**Outlier detection method per replicate group (Shapiro-Wilk pre-test enabled):**\n\n")
  for (i in seq_len(nrow(outlier_method_log))) {
    row <- outlier_method_log[i, ]
    cat(sprintf("- %s: %s (Shapiro-Wilk p = %.4f)\n", row$Replicate, row$method_used, row$shapiro_p))
  }
  cat("\n")
}
```

```{r ci-method-note, results='asis'}
if (isTRUE(analysis_config$ci_method == "bootstrap")) {
  cat(sprintf("\n*%s*\n\n", tr("ci_bootstrap_note", lang)))
  cat("*Replicate-level CIs computed via percentile bootstrap (1000 resamples).*\n\n")
} else {
  cat(sprintf("\n*%s*\n\n", tr("ci_tdist_note", lang)))
  cat("*Replicate-level CIs reflect measurement variability using Student's t-distribution (df = n-1).*\n\n")
}
cat(sprintf("*%s*\n\n", tr("ci_delta_method_note", lang)))
```

```{r ci-asymmetric-footnote, results='asis'}
cat(sprintf("\n*%s*\n\n", tr("ci_asymmetric_footnote", lang)))
```

```{r save-sample-stats, include=FALSE}
# BUG_008 fix: Update model_stats with mean_sample_cv now that replicate_stats exists
if (exists("model_stats") && exists("replicate_stats") &&
    !is.null(replicate_stats) && nrow(replicate_stats) > 0) {
  model_stats$mean_sample_cv <- mean(replicate_stats$cv_percent, na.rm = TRUE)
} else if (exists("model_stats")) {
  model_stats$mean_sample_cv <- NA_real_
}
# Re-save model_stats with the updated mean_sample_cv
if (exists("model_stats") && exists("write_json_safe")) {
  write_json_safe(model_stats, file.path(output_dir, "model_stats.json"))
}
```

## `r tr("detailed_summary", lang)`

<details>
<summary>Click to expand detailed results table</summary>

```{r summary-results-table}
if (exists("replicate_summary") && !is.null(replicate_summary) && nrow(replicate_summary) > 0) {
  
  # Format the summary table for display
  summary_display <- replicate_summary %>%
    dplyr::mutate(
      mean_concentration = if (isTRUE(is_elisa)) {
        format(as.numeric(mean_concentration), digits = 4, nsmall = 1, big.mark = ",")
      } else {
        format(as.numeric(mean_concentration), scientific = TRUE, digits = 3)
      },
      std_dev = if (isTRUE(is_elisa)) {
        ifelse(is.na(std_dev), "\u2014", format(as.numeric(std_dev), digits = 3, nsmall = 1))
      } else {
        ifelse(is.na(std_dev), "\u2014", format(as.numeric(std_dev), scientific = TRUE, digits = 2))
      },
      std_error = if (isTRUE(is_elisa)) {
        ifelse(is.na(std_error), "\u2014", format(as.numeric(std_error), digits = 3, nsmall = 1))
      } else {
        ifelse(is.na(std_error), "\u2014", format(as.numeric(std_error), scientific = TRUE, digits = 2))
      },
      ci_95 = if (isTRUE(is_elisa)) {
        paste0("[", format(as.numeric(ci_lower_95), digits = 3, nsmall = 1), " \u2013 ", 
               format(as.numeric(ci_upper_95), digits = 3, nsmall = 1), "]")
      } else {
        paste0("[", format(as.numeric(ci_lower_95), scientific = TRUE, digits = 2), " \u2013 ", 
               format(as.numeric(ci_upper_95), scientific = TRUE, digits = 2), "]")
      },
      cv_percent = ifelse(is.na(cv_percent), "\u2014", format(as.numeric(cv_percent), digits = 3, nsmall = 1)),
      range_flag_display = if ("range_flag" %in% names(.)) {
        ifelse(is.na(range_flag) | range_flag == "", "\u2705", range_flag)
      } else {
        "\u2705"
      },
      mean_bb0 = if ("mean_response" %in% names(.)) {
        ifelse(is.na(mean_response), "\u2014", paste0(round(as.numeric(mean_response), 2), "%"))
      } else {
        NULL
      }
    ) %>%
    dplyr::select(-ci_lower_95, -ci_upper_95,
                  -any_of(c("within_range", "range_flag", "mean_response"))) %>%
    dplyr::relocate(ci_95, .before = cv_percent) %>%
    dplyr::relocate(range_flag_display, .after = cv_percent) %>%
    { if ("mean_bb0" %in% names(.)) dplyr::relocate(., mean_bb0, .before = range_flag_display) else . }
  
  # Set column names
  conc_unit <- if (isTRUE(is_elisa)) paste0(assay_config$units %||% "pg/mL") else "mol/L"
  
  # Build column name vector dynamically based on available columns
  base_names <- c("Replicate Group", "Sample IDs", "Sample Type", "n", 
                  paste0("Mean (", conc_unit, ")"), 
                  paste0("SD (", conc_unit, ")"),
                  paste0("SE (", conc_unit, ")"),
                  "95% CI", "CV [%]")
  
  # Add %B/B0 and range flag columns
  if ("mean_bb0" %in% names(summary_display)) {
    base_names <- c(base_names, "%B/B0")
  }
  base_names <- c(base_names, "Range")
  
  # Add tissue concentration columns if present
  if ("TissueWeight_mg" %in% names(summary_display)) {
    summary_display <- summary_display %>%
      dplyr::mutate(
        TissueWeight_mg = ifelse(is.na(TissueWeight_mg), "\u2014",
                                  format(TissueWeight_mg, digits = 4, nsmall = 1)),
        concentration_pg_per_g = ifelse(is.na(concentration_pg_per_g), "\u2014",
                                         format(concentration_pg_per_g, digits = 4, nsmall = 1, big.mark = ","))
      )
    base_names <- c(base_names, "Tissue Mass [mg]", "Conc. (pg/g)")
  }
  
  # Remove extra columns that aren't in our name mapping
  summary_display <- summary_display %>%
    dplyr::select(-any_of(c("total_amount_pg")))
  
  names(summary_display) <- base_names[1:ncol(summary_display)]
  
  render_table(summary_display,
              caption = tr("detailed_caption", lang))
}
```

</details>

## `r tr("sample_variability", lang)`

```{r sample-boxplot, fig.width=10, fig.height=7, warning=FALSE, message=FALSE, fig.cap="Sample concentration estimates with confidence intervals"}
if (exists("sample_results") && !is.null(sample_results) && nrow(sample_results) > 0) {
  
  sample_plot_data <- sample_results %>%
    dplyr::filter(!is.na(estimated_concentration), is.finite(estimated_concentration)) %>%
    dplyr::mutate(
      Replicate_wrapped = stringr::str_wrap(Replicate, width = 15),
      within_range = mapply(classify_range, response_value, estimated_concentration,
                            MoreArgs = list(is_elisa_assay = is_elisa)),
      range_flag = mapply(flag_range, response_value, estimated_concentration,
                          MoreArgs = list(is_elisa_assay = is_elisa))
    )
  
  if (nrow(sample_plot_data) > 0) {
    
    # Create boxplot
    sample_box <- ggplot(sample_plot_data, 
                        aes(x = factor(Replicate_wrapped), 
                            y = estimated_concentration)) +
      geom_boxplot(outlier.shape = NA) +
      geom_jitter(aes(color = within_range,
                     text = paste0(
                       "Replicate: ", Replicate, "<br>",
                       "Sample ID: ", SampleID, "<br>",
                       "Concentration: ", 
                       if (isTRUE(is_elisa)) {
                         format(estimated_concentration, digits = 4, big.mark = ",")
                       } else {
                         scales::scientific(estimated_concentration, digits = 2)
                       }, "<br>",
                       "Range Status: ", within_range
                     )), 
                 width = 0.15, size = 2) +
      xlab("Replicate Group") +
      ylab(if (isTRUE(is_elisa)) {
        paste0("Estimated Concentration (", assay_config$units %||% "pg/mL", ")")
      } else {
        "Estimated Concentration (mol/L)"
      }) +
      theme_classic() +
      theme(
        legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "top"
      )
    
    # Add log scale if appropriate
    conc_range <- max(sample_plot_data$estimated_concentration, na.rm = TRUE) / 
                  min(sample_plot_data$estimated_concentration, na.rm = TRUE)
    
    if (!isTRUE(is_elisa) || conc_range > 100) {
      sample_box <- sample_box +
        scale_y_log10(labels = if (isTRUE(is_elisa)) {
          function(x) format(x, big.mark = ",")
        } else {
          scales::scientific_format(digits = 2)
        })
    }

    # Dynamic font sizing based on number of groups
    n_groups <- length(unique(sample_plot_data$Replicate_wrapped))
    x_font_size <- if (n_groups > 20) 7 else if (n_groups > 10) 9 else 11
    sample_box <- sample_box +
      theme(axis.text.x = element_text(angle = 45, hjust = 1, size = x_font_size))

    # For very large experiments, split into faceted panels
    if (n_groups > 30) {
      sample_plot_data$panel <- ceiling(as.numeric(factor(sample_plot_data$Replicate)) / 15)
      sample_box <- sample_box + facet_wrap(~panel, scales = "free_x", ncol = 1)
    }

    # Render plot - same pattern as working standard DRC
    if (knitr::is_html_output()) {
      plotly::ggplotly(sample_box, tooltip = "text")
    } else {
      print(sample_box)
    }
  }
}
```

## `r tr("heatmap_title", lang)`

<details>
<summary>Click to expand plate heatmap</summary>

```{r plate-heatmap, fig.width=10, fig.height=6, warning=FALSE, message=FALSE}
tryCatch({
  cat(tr("heatmap_desc", lang), "\n\n")

  # Reconstruct plate layout from long data
  heatmap_data <- data_long %>%
    dplyr::filter(!is.na(MeasurementValue)) %>%
    dplyr::mutate(
      Row = factor(Row, levels = LETTERS[1:8]),
      Column = as.numeric(gsub("[^0-9]", "", as.character(Column)))
    ) %>%
    dplyr::filter(!is.na(Row) & !is.na(Column))

  if (nrow(heatmap_data) > 0) {
    heatmap_plot <- ggplot(heatmap_data, aes(x = factor(Column), y = Row, fill = MeasurementValue)) +
      geom_tile(color = "white", size = 0.5) +
      geom_text(aes(label = sprintf("%.3f", MeasurementValue)), size = 2, color = "black") +
      scale_fill_gradient2(low = "#2166AC", mid = "#F7F7F7", high = "#B2182B",
                           midpoint = median(heatmap_data$MeasurementValue, na.rm = TRUE),
                           name = "Value") +
      scale_y_discrete(limits = rev(levels(factor(heatmap_data$Row, levels = LETTERS[1:8])))) +
      labs(x = "Column", y = "Row", title = tr("heatmap_title", lang)) +
      theme_minimal() +
      theme(
        panel.grid = element_blank(),
        axis.text = element_text(size = 10, face = "bold"),
        plot.title = element_text(hjust = 0.5)
      )

    print(heatmap_plot)
  }
}, error = function(e) {
  cat(sprintf("\n\n*Heatmap could not be generated: %s*\n\n", e$message))
})
```

</details>

```{r plate-positional-qc, results='asis'}
cat("\n\n### Plate Positional Quality\n\n")
if (exists("data_long") && "MeasurementValue" %in% names(data_long)) {
  plate_data <- data_long %>% filter(is.finite(MeasurementValue))
  if (nrow(plate_data) > 0) {
    plate_mean <- mean(plate_data$MeasurementValue, na.rm = TRUE)
    plate_sd <- sd(plate_data$MeasurementValue, na.rm = TRUE)
    row_stats <- plate_data %>%
      group_by(Row) %>%
      summarise(Mean = mean(MeasurementValue, na.rm = TRUE),
                SD = sd(MeasurementValue, na.rm = TRUE), .groups = "drop") %>%
      mutate(Flag = ifelse(abs(Mean - plate_mean) > 2 * plate_sd, "\u26a0", ""))
    flagged_rows <- sum(row_stats$Flag == "\u26a0")
    col_stats <- plate_data %>%
      group_by(Column) %>%
      summarise(Mean = mean(MeasurementValue, na.rm = TRUE),
                SD = sd(MeasurementValue, na.rm = TRUE), .groups = "drop") %>%
      mutate(Flag = ifelse(abs(Mean - plate_mean) > 2 * plate_sd, "\u26a0", ""))
    flagged_cols <- sum(col_stats$Flag == "\u26a0")
    if (flagged_rows > 0 || flagged_cols > 0) {
      cat(sprintf("**Warning:** %d row(s) and %d column(s) deviate >2 SD from plate mean, suggesting possible edge effects.\n\n", flagged_rows, flagged_cols))
    } else {
      cat("No significant row or column bias detected (all within 2 SD of plate mean).\n\n")
    }
    row_display <- row_stats
    row_display$Mean <- round(row_display$Mean, 3)
    row_display$SD <- round(row_display$SD, 3)
    names(row_display) <- c("Row", "Mean Response", "SD", "Flag")
    render_table(row_display, caption = "Response by Row")
  }
}
```

## `r tr("drc_with_samples", lang)`

```{r drc-with-samples, fig.width=12, fig.height=8, warning=FALSE, message=FALSE, fig.cap="Dose-response curve with sample concentration estimates overlaid on the fitted model"}
if (exists("sample_results") && !is.null(sample_results) && nrow(sample_results) > 0) {
  
  standards_plot_data <- data_long %>%
    dplyr::filter(!is.na(concentration), SampleType == "Standard")
  
  unknown_plot_data <- sample_results %>%
    dplyr::filter(!is.na(concentration_diluted), is.finite(concentration_diluted)) %>%
    dplyr::mutate(
      falls_within_range = mapply(classify_range, response_value, estimated_concentration,
                                  MoreArgs = list(is_elisa_assay = is_elisa))
    )
  
  if (nrow(unknown_plot_data) > 0) {
    
    combined_drc <- ggplot() +
      geom_line(data = model_fits, aes(x = conc, y = p), 
               color = "black", size = 1.2) +
      geom_point(data = standards_plot_data, 
                aes(x = concentration, y = .data[[response_var]], 
                    color = high_variability), 
                size = 3, alpha = 0.7) +
      geom_point(data = unknown_plot_data,
                aes(x = concentration_diluted, y = response_value,
                    shape = falls_within_range,
                    text = paste0(
                      "Sample: ", SampleID, "<br>",
                      "Replicate: ", Replicate, "<br>",
                      "Dilution: 1:", DilutionFactor, "<br>",
                      "Actual Concentration: ",
                      if (isTRUE(is_elisa)) {
                        format(estimated_concentration, digits = 4, big.mark = ",")
                      } else {
                        scales::scientific(estimated_concentration, digits = 2)
                      }, "<br>",
                      "Diluted Concentration: ",
                      if (isTRUE(is_elisa)) {
                        format(concentration_diluted, digits = 4, big.mark = ",")
                      } else {
                        scales::scientific(concentration_diluted, digits = 2)
                      }, "<br>",
                      labels$y_label, ": ", round(response_value, 2), "<br>",
                      if ("concentration_pg_per_g" %in% names(unknown_plot_data) &&
                          !is.na(concentration_pg_per_g)) {
                        paste0("Conc. (pg/g tissue): ",
                               format(concentration_pg_per_g, digits = 4, nsmall = 1, big.mark = ","), "<br>")
                      } else { "" },
                      "Range Status: ", falls_within_range
                    )),
                size = 3, color = "#E69F00") +
      scale_shape_manual(name = "Sample Status",
                        values = c("Within Range" = 16, "Out of Range" = 17, "Unknown" = 18)) +
      scale_color_manual(name = "Standard Variability",
                        values = c("Normal Variability" = "#0072B2", 
                                  "High Variability" = "#D55E00")) +
      theme_classic() +
      theme(
        legend.position = "top",
        legend.box = "vertical",
        axis.text.x = element_text(angle = 45, hjust = 1)
      ) +
      labs(x = labels$x_label, y = labels$y_label,
           title = tr("drc_combined_title", lang))
    
    if (isTRUE(is_elisa)) {
      combined_drc <- combined_drc + 
        scale_x_log10(
          labels = function(x) format(x, digits = 3, big.mark = ","),
          breaks = scales::trans_breaks("log10", function(x) 10^x, n = 6)
        )
    } else {
      combined_drc <- combined_drc + 
        scale_x_log10(
          limits = c(1e-12, 1e-5),
          breaks = 10^seq(-12, -5, by = 1),
          labels = scales::scientific_format(digits = 2)
        )
    }
    
    # Render plot - same pattern as working standard DRC
    if (knitr::is_html_output()) {
      plotly::ggplotly(combined_drc, tooltip = "text")
    } else {
      print(combined_drc)
    }
  }
}
```

## Summary

```{r report-summary, results='asis'}
assay_type_label <- toupper(assay_config$assay_type %||% "Unknown")
analyte_label <- assay_config$analyte %||% assay_config$toxin_standard_label %||% "Unknown"
cat(sprintf("- **Assay:** %s \u2014 %s\n", assay_type_label, analyte_label))

n_std_used <- if (exists("model_stats") && !is.null(model_stats)) model_stats$n_standards %||% NA_integer_ else NA_integer_
n_std_total <- if (exists("model_stats") && !is.null(model_stats)) model_stats$n_unique_concs %||% NA_integer_ else NA_integer_
if (!is.na(n_std_used) && !is.na(n_std_total)) {
  cat(sprintf("- **Standards:** %d used (after exclusions) out of %d\n", n_std_used, n_std_total))
}

weight_info <- if (exists("model_stats") && !is.null(model_stats)) model_stats$weight_method %||% "none" else "none"
cat(sprintf("- **Model:** 4PL logistic with %s weighting\n", weight_info))

if (exists("replicate_stats") && !is.null(replicate_stats) && nrow(replicate_stats) > 0) {
  n_total <- nrow(replicate_stats)
  n_quantified <- if ("within_range" %in% names(replicate_stats)) sum(replicate_stats$within_range == "Within Range", na.rm = TRUE) else NA_integer_
  n_flagged <- if (!is.na(n_quantified)) n_total - n_quantified else NA_integer_
  if (!is.na(n_quantified)) {
    cat(sprintf("- **Samples analyzed:** %d total, %d within range, %d flagged\n", n_total, n_quantified, n_flagged))
  } else {
    cat(sprintf("- **Samples analyzed:** %d total\n", n_total))
  }
}

if (exists("formal_lloq") && exists("formal_uloq") && !is.na(formal_lloq) && !is.na(formal_uloq)) {
  conc_unit <- if (isTRUE(is_elisa)) assay_config$units %||% "pg/mL" else "mol/L"
  lloq_fmt <- if (isTRUE(is_elisa)) format(formal_lloq, digits = 4, big.mark = ",") else format(formal_lloq, scientific = TRUE, digits = 3)
  uloq_fmt <- if (isTRUE(is_elisa)) format(formal_uloq, digits = 4, big.mark = ",") else format(formal_uloq, scientific = TRUE, digits = 3)
  cat(sprintf("- **LLOQ:** %s %s, **ULOQ:** %s %s\n", lloq_fmt, conc_unit, uloq_fmt, conc_unit))
}
```

## Interpretation & Recommendations

```{r report-interpretation, results='asis'}
qc_issues <- c()

r2_val <- if (exists("model_stats") && !is.null(model_stats)) model_stats$r_squared %||% NA_real_ else NA_real_
if (!is.na(r2_val) && is.finite(r2_val) && r2_val < 0.95) {
  qc_issues <- c(qc_issues, sprintf("Low R\u00b2 (%.3f)", r2_val))
}

hill_val <- if (exists("model_stats") && !is.null(model_stats)) model_stats$hill_slope %||% NA_real_ else NA_real_
if (!is.na(hill_val) && is.finite(hill_val) && (abs(hill_val) < 0.5 || abs(hill_val) > 3.0)) {
  qc_issues <- c(qc_issues, sprintf("Unusual Hill slope (%.2f)", hill_val))
}

if (exists("replicate_stats") && !is.null(replicate_stats) && "cv_percent" %in% names(replicate_stats)) {
  max_cv <- max(replicate_stats$cv_percent, na.rm = TRUE)
  if (is.finite(max_cv) && max_cv > 30) {
    qc_issues <- c(qc_issues, sprintf("High sample CV (max %.1f%%)", max_cv))
  }
}

if (length(qc_issues) == 0) {
  cat("All quality criteria met. Results are suitable for reporting.\n\n")
} else {
  cat(sprintf("**Quality warnings detected:** %s. Review flagged items before use.\n\n",
              paste(qc_issues, collapse = ", ")))
}
```

---

**`r tr("report_generated", lang)`** `r Sys.Date()`
**`r tr("contact", lang)`** kr.moeller@iaea.org
**`r tr("feedback", lang)`** [`r tr("online_form", lang)`](https://forms.office.com/e/q8eqJfp4QM)

*`r tr("automated_analysis", lang)`*