---
title: "Bioassay Analysis Report"
output:
  html_document:
    df_print: paged
    self_contained: true
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: "default"
    css: report_style.css
  word_document:
    toc: true
    fig_caption: true
    fig_width: 6
    fig_height: 4
    reference_docx: reference.docx
  pdf_document:
    toc: true
header-includes:
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \fancyfoot[C]{\thepage}
  - \fancyhead{}
  - \renewcommand{\headrulewidth}{0pt}
params:
  output_dir: NULL
  lang: "en"
  compact: false
---

<!--
================================================================================
Chunk pipeline (in execution order):

  setup                           - libraries, output_dir, i18n, source helpers
  load-data                       - read CSV, validate, compute %B/B0 for ELISA
  executive-summary               - top-of-report status box
  standards-table                 - display standard concentrations
  methods-section                 - methods paragraph with citations
  model-fitting                   - fit all selected weightings (LL.4 -> LL.3 ->
                                    interpolation fallback); defines:
                                    all_models, model_fit, R2, RMSE,
                                    conc_range, model_fits, ec20, ec80,
                                    classify_range(), flag_range()
  exec-summary-box                - rendered exec summary (needs model_fit)
  weight-comparison               - overlay plot of weightings (HTML)
  weighting-suitability           - Brown-Forsythe / variance ratio
  model-stability-assessment      - convergence / boundary diagnostics
  lloq-uloq                       - back-calculation accuracy -> formal LLOQ/ULOQ
  dose-response-plots             - main DRC plot
  model-stats                     - 4PL coefficient table
  save-model-stats                - write model_stats.json (partial)
  standard-backcalculation        - per-standard recovery table
  sample-analysis                 - MAIN SAMPLE QUANTIFICATION (defines:
                                    sample_results, replicate_stats,
                                    outlier_flags, replicate_summary,
                                    tissue-normalized columns if ELISA)
  qc-summary-card                 - merged severity-coded QC card
                                    (well counts + exclusions + outliers
                                    + high-CV groups in one banner)
  sample-results-table            - rendered sample table + CSV writes
  outlier-results                 - outlier detection report block
  ci-method-note                  - CI method footnote
  save-sample-stats               - updates model_stats.json with mean_sample_cv
  summary-results-table           - detailed replicate summary
  sample-boxplot                  - variability visualization
  plate-heatmap                   - plate heatmap
  plate-positional-qc             - row/column bias check
  drc-with-samples                - DRC plot with sample points overlaid
  parallelism-assessment          - optional multi-curve comparison
  report-summary                  - high-level summary bullets
  report-interpretation           - QC-profile-based interpretation
  tissue-normalization-traceability - ELISA + tissue only
  exclusion-audit                 - formal exclusion audit table

Key cross-chunk dependencies:
  - model-fitting defines variables used by almost all later chunks
  - sample-analysis must run before sample-results-table, outlier-results,
    save-sample-stats, summary-results-table, sample-boxplot
  - tissue_weights (ELISA) is loaded inside sample-analysis
================================================================================
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = TRUE, message = FALSE)

# For DOCX/PDF output, force static PNG figures at high resolution
if (!knitr::is_html_output()) {
  knitr::opts_chunk$set(
    dev = "png",
    dpi = 300,
    fig.width = 6,
    fig.height = 4,
    fig.retina = 1
  )
}

# Load required libraries.
# kableExtra guard: install if missing so the Rmd is self-healing in
# environments where global.R's install step was skipped or failed.
if (!requireNamespace("kableExtra", quietly = TRUE)) {
  .rmd_repos <- c("https://packagemanager.posit.co/cran/latest",
                  "https://cloud.r-project.org")
  install.packages("kableExtra", repos = .rmd_repos, quiet = TRUE)
  if (!requireNamespace("kableExtra", quietly = TRUE)) {
    if (!requireNamespace("remotes", quietly = TRUE))
      install.packages("remotes", repos = .rmd_repos, quiet = TRUE)
    remotes::install_github("haozhu233/kableExtra")
  }
  rm(.rmd_repos)
}
# DT: installed by global.R before render; guard against direct rmarkdown::render() calls.
.has_DT <- requireNamespace("DT", quietly = TRUE)
if (.has_DT) library(DT)
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(drc)
library(scales)
library(knitr)
library(kableExtra)
library(stringr)
library(ggtext)
library(glue)
library(htmltools)
library(tibble)
library(jsonlite)

# %||% operator is defined in global.R (loaded before rendering)

# H4: compact-mode flag for chunk gating below.
is_compact <- isTRUE(params$compact)

# Validate output directory
output_dir <- normalizePath(params$output_dir, winslash = "/", mustWork = TRUE)

# Load language setting
lang <- params$lang %||% "en"

# Get script directory FIRST (needed before sourcing any files).
# When this template is included via knitr::knit_child() (multiwavelength
# template), knitr::current_input() returns the preprocessed temp-file path
# in tempdir(), not the reports/ directory. Validate the candidate directory
# actually contains report_constants.R and fall back to getwd() (which knitr
# sets to knit_root_dir = reports/) when it doesn't.
script_dir <- tryCatch({
  rmd_path <- knitr::current_input()
  candidate <- if (!is.null(rmd_path) && nzchar(rmd_path) && rmd_path != ".") {
    normalizePath(dirname(rmd_path), winslash = "/")
  } else {
    NULL
  }
  if (!is.null(candidate) && file.exists(file.path(candidate, "report_constants.R"))) {
    candidate
  } else if (file.exists(file.path(getwd(), "report_constants.R"))) {
    normalizePath(getwd(), winslash = "/")
  } else if (dir.exists("reports")) {
    normalizePath("reports", winslash = "/")
  } else {
    possible_reports_dir <- file.path(dirname(params$output_dir), "reports")
    if (dir.exists(possible_reports_dir)) {
      normalizePath(possible_reports_dir, winslash = "/")
    } else {
      "."
    }
  }
}, error = function(e) {
  if (file.exists(file.path(getwd(), "report_constants.R"))) {
    normalizePath(getwd(), winslash = "/")
  } else if (dir.exists("reports")) {
    normalizePath("reports", winslash = "/")
  } else {
    "."
  }
})

# Source i18n translations (search multiple locations)
i18n_found <- FALSE
for (i18n_candidate in c(
  file.path(script_dir, "i18n.R"),
  file.path(dirname(script_dir), "i18n.R"),
  file.path(dirname(script_dir), "server", "i18n.R"),
  file.path(getwd(), "server", "i18n.R"),
  "../i18n.R",
  "../server/i18n.R",
  "i18n.R",
  file.path(dirname(output_dir), "i18n.R"),
  file.path(dirname(output_dir), "server", "i18n.R")
)) {
  if (file.exists(i18n_candidate)) {
    source(i18n_candidate)
    i18n_found <- TRUE
    break
  }
}
if (!i18n_found) {
  tr <- function(key, lang = "en", ...) paste0("[", key, "]")
}

# Check if all required function files exist
# M1.7: analysis_pipeline.R and report_sections.R are scaffolded empty now;
# subsequent M1.1-M1.6 extractions will populate them.
required_files <- c("report_constants.R", "report_functions.R",
                    "plot_functions.R", "analysis_pipeline.R",
                    "report_sections.R")
missing_files <- c()

for (file in required_files) {
  file_path <- file.path(script_dir, file)
  if (!file.exists(file_path)) {
    missing_files <- c(missing_files, file_path)
  }
}

if (length(missing_files) > 0) {
  stop("Missing required function files:\n", paste(missing_files, collapse = "\n"),
       "\n\nExpected in directory: ", script_dir,
       "\nActual files found: ", paste(list.files(script_dir, pattern = "\\.R$"), collapse = ", "),
       "\n\nMake sure all R files are in the 'reports/' folder.")
}

# Source function files
source(file.path(script_dir, "report_constants.R"))
source(file.path(script_dir, "report_functions.R"))
source(file.path(script_dir, "plot_functions.R"))
source(file.path(script_dir, "analysis_pipeline.R"))
source(file.path(script_dir, "report_sections.R"))

# AUDIT-020: Set a deterministic RNG seed for reproducible bootstrap CIs.
# The seed value is centralised in STATS_CONFIG (report_constants.R) so it
# can be inspected in the report metadata sidecar without hard-coding it here.
set.seed(STATS_CONFIG$report_seed)

# Load assay configuration
assay_config <- tryCatch({
  load_assay_config(output_dir)
}, error = function(e) {
  # Fallback to RBA defaults
  list(
    assay_type = "rba",
    detection_method = "radioligand",
    toxin_standard_label = "Saxitoxin"
  )
})

# Load analysis settings
analysis_config <- tryCatch({
  cfg_file <- file.path(output_dir, "analysis_config.json")
  if (file.exists(cfg_file)) {
    jsonlite::fromJSON(cfg_file)
  } else {
    list(regression_weight = "none", quant_range_min = 20, quant_range_max = 80,
         ci_method = "t_dist", enable_outlier_detection = FALSE, outlier_min_n = 3,
         normality_assumption = "assume")
  }
}, error = function(e) {
  list(regression_weight = "none", quant_range_min = 20, quant_range_max = 80,
       ci_method = "t_dist", enable_outlier_detection = FALSE, outlier_min_n = 3,
       normality_assumption = "assume")
})
# Ensure backward compatibility for configs saved without normality_assumption
if (is.null(analysis_config$normality_assumption)) {
  analysis_config$normality_assumption <- "assume"
}
# Resolve configurable CV threshold (app-level override or fallback to global constant)
cv_limit <- analysis_config$cv_limit %||% QC_THRESHOLDS$cv_limit

# Helper: format numbers with scientific notation for extreme values
fmt_smart <- function(x, digits = 4) {
  if (is.na(x) || !is.finite(x)) return("NA")
  if (abs(x) >= 1e5 || (abs(x) > 0 & abs(x) < 1e-4)) {
    formatC(x, format = "e", digits = digits)
  } else {
    formatC(x, format = "f", digits = digits)
  }
}
```

```{css details-style, echo=FALSE, eval=knitr::is_html_output()}
/* Report-only typography overrides. The report ships as a self-contained
   HTML for email/offline use; using a system font stack (rather than
   embedding webfonts via @import) keeps the file lean and renders
   identically without network access. */

body {
  font-family: -apple-system, "Segoe UI", "Helvetica Neue", Arial, sans-serif;
  color: #0f1722;
  font-size: 13px;
  line-height: 1.6;
}
.title { /* rmarkdown's auto title — system serif display heading */
  font-family: Georgia, "Times New Roman", serif;
  font-weight: 500;
  font-size: 32px;
  letter-spacing: -0.01em;
  color: #0f1722;
  margin-bottom: 4px;
}
h1, h2, h3, h4 {
  font-family: -apple-system, "Segoe UI", "Helvetica Neue", Arial, sans-serif;
  font-weight: 600;
  color: #0f1722;
  margin-top: 24px;
}
h1 { font-size: 22px; font-weight: 500; }
h2 { font-size: 17px; }
h3 { font-size: 14px; }
h4 { font-size: 13px; }
.mono, code, kbd, pre, tt {
  font-family: ui-monospace, "SF Mono", Menlo, Consolas, monospace;
}

/* Section <details> chevron + sentence-case heading (§6.7) */
details > summary {
  cursor: pointer;
  font-weight: 500;
  color: #2a3340;
  margin-bottom: 8px;
}
details > summary:hover { color: #1f3a5f; }
details {
  margin-bottom: 12px;
}

/* GUI refresh §6.6: kable / kableExtra table styling. Mono numerics,
   tabular numerals, small-caps mono header, surface-2 hover. */
table.kable, table.dataframe, .kable_wrapper table {
  border-collapse: collapse;
  width: 100%;
  font-size: 12px;
}
table.kable thead th, table.dataframe thead th, .kable_wrapper thead th {
  background: #f1f3f5;
  color: #5b6573;
  font-family: ui-monospace, "SF Mono", Menlo, Consolas, monospace;
  font-size: 11px;
  font-weight: 500;
  letter-spacing: 0.04em;
  text-transform: uppercase;
  padding: 6px 10px;
  border-bottom: 1px solid #e3e6ea;
  text-align: left;
}
table.kable td, table.dataframe td, .kable_wrapper td {
  padding: 6px 10px;
  border-bottom: 1px solid #e3e6ea;
  vertical-align: top;
}
/* Numeric-looking cells (mono + tabular) get right-aligned. heuristic:
   any <td> where rmarkdown's pandoc CSS class has been applied. */
table.kable td.r, table.dataframe td.r {
  font-family: ui-monospace, "SF Mono", Menlo, Consolas, monospace;
  font-variant-numeric: tabular-nums;
  text-align: right;
}
table.kable tbody tr:hover, table.dataframe tbody tr:hover {
  background: #f1f3f5;
}

/* GUI refresh §6.3: executive summary card — quiet 1 px line, no
   tinted gradient (the .Rmd helper renders inline styling for
   email-safe fallback, this overrides for live HTML).
   Exclude is-warn/is-fail so QC-alert cards keep their colour. */
.exec-summary-card:not(.is-warn):not(.is-fail), div[style*="linear-gradient(135deg, #E8F5E9"] {
  background: #ffffff !important;
  border: 1px solid #e3e6ea !important;
  border-left: 4px solid #009e73 !important;
  box-shadow: none !important;
}
```

```{r load-data, include=FALSE}
# Load and validate data
data_long_raw <- read.csv(file.path(output_dir, "long_data_output.csv"), stringsAsFactors = FALSE)

# Remove diagnostic control columns that confuse users
diagnostic_cols <- c("ControlValues_B0", "ControlValues_NSB", "ControlValues_Blank")
data_long <- data_long_raw %>%
  dplyr::select(-any_of(diagnostic_cols))

validate_input_data(data_long)

# Detect assay type from data if not specified in config
has_normalized <- "NormalizedValue" %in% names(data_long)
# Use assay_type from config as the authoritative source. The
# `|| has_normalized` fallback was added for legacy data without a config, but
# since normalize_assay_data() now adds NormalizedValue for RBA too, that
# fallback wrongly triggered for RBA. Restrict fallback to configs where
# assay_type is absent.
is_elisa <- isTRUE(assay_config$assay_type == "elisa") ||
            (has_normalized && is.null(assay_config$assay_type))

# For ELISA, calculate proper B/B0 values following Cayman protocol
if (isTRUE(is_elisa)) {
  # Step 1: Try B/B0 calculation (keep full result so we can read the attribute)
  .bb0_result <- tryCatch(
    calculate_elisa_bb0(data_long_raw, use_percent = TRUE),
    error = function(e) {
      warning("Could not calculate B/B0 values: ", e$message)
      NULL
    }
  )

  # Step 2: Capture control_summary NOW — before dplyr::select() strips the attribute
  control_summary <- if (!is.null(.bb0_result)) attr(.bb0_result, "control_summary") else NULL

  # Step 3: Build the final data_long
  data_long <- if (!is.null(.bb0_result)) {
    .bb0_result %>% dplyr::select(-any_of(diagnostic_cols))
  } else {
    warning("Falling back to existing NormalizedValue column")
    data_long$calculated_bb0 <- data_long$NormalizedValue
    data_long
  }

  response_var <- "calculated_bb0"

  # Update assay configuration
  if (!isTRUE(assay_config$assay_type == "elisa")) {
    assay_config$assay_type <- "elisa"
    assay_config$units <- "pg/mL"
    assay_config$analyte <- assay_config$analyte %||% "Analyte"
  }

} else {
  # RBA: Use existing response variable
  response_var <- if (has_normalized && "NormalizedValue" %in% names(data_long)) {
    "NormalizedValue"
  } else {
    "MeasurementValue"  
  }
}

# Get axis labels
labels <- get_axis_labels(assay_config, lang = lang)

# Load notes (optional)
notes_file <- file.path(output_dir, "notes.json")
notes_text <- "No notes added."
if (file.exists(notes_file)) {
  notes_obj <- jsonlite::fromJSON(notes_file, simplifyVector = FALSE)
  notes_text <- notes_obj$notes %||% "No notes added."
}

# Load sample processing configuration for ELISA
if (isTRUE(is_elisa)) {
  processing_config <- load_sample_processing_config(output_dir)
}
```


```{r report-meta, results='asis'}
# Item 12: replace the former <h4 class="author"> + inline bold lines with a
# muted two-column metadata header.  HTML gets the styled .bs-report-meta
# block; DOCX/PDF falls back to compact bold lines.
.drc_ver_meta <- tryCatch(as.character(packageVersion("drc")), error = function(e) "?")
.assay_lbl    <- paste0(toupper(assay_config$assay_type %||% ""),
                        " — ",
                        assay_config$analyte %||%
                          assay_config$toxin_standard_label %||% "")
.app_ver_meta <- REPORT_INFO$version %||% "?"

if (is_html_out()) {
  # Tiles (render_report_header) own: Analyst, Analysis Date, Assay Type, Wells.
  # Masthead keeps only fields not duplicated by tiles: Authors and Software.
  cat(paste0(
    '<header class="bs-report-meta">\n',
    '<div class="meta-row">',
      '<span class="label">Authors</span>',
      '<span class="value">Kristof Moeller (IAEA, Monaco)</span>',
    '</div>\n',
    '<div class="meta-row">',
      '<span class="label">Software</span>',
      sprintf('<span class="value">Bioassay Analysis App v%s · %s · drc v%s</span>',
              .app_ver_meta, R.version.string, .drc_ver_meta),
    '</div>\n',
    '</header>\n\n'
  ))
} else {
  cat(paste0(
    "**Author:** Kristof Moeller (IAEA, Monaco)  \n",
    sprintf("**Generated:** %s · %s  \n", format(Sys.Date(), "%Y-%m-%d"), Sys.info()[["user"]]),
    sprintf("**Assay:** %s  \n", .assay_lbl),
    sprintf("**Software:** Bioassay Analysis App v%s · %s · drc v%s  \n\n",
            .app_ver_meta, R.version.string, .drc_ver_meta)
  ))
}
```

```{r report-metadata-sidecar, include=FALSE}
# AUDIT-018: Write a machine-readable sidecar (report_metadata.json) alongside
# each rendered report so that the exact inputs, RNG seed, and R environment
# that produced the report are permanently recorded.
.run_uuid <- paste(sample(c(letters[1:6], 0:9), 32L, replace = TRUE), collapse = "")
.input_hash <- tryCatch(
  digest::digest(
    list(
      read.csv(file.path(output_dir, "long_data_output.csv")),
      jsonlite::read_json(file.path(output_dir, "model_stats.json"))
    ),
    algo = "sha256"
  ),
  error = function(e) NA_character_
)
.sinfo <- utils::sessionInfo()
.meta <- list(
  run_uuid    = .run_uuid,
  input_hash  = .input_hash,
  report_seed = STATS_CONFIG$report_seed,
  rendered_at = format(Sys.time(), "%Y-%m-%dT%H:%M:%S%z"),
  r_version   = paste(R.version$major, R.version$minor, sep = "."),
  packages    = sapply(.sinfo$otherPkgs, function(p) p$Version, USE.NAMES = TRUE)
)
jsonlite::write_json(.meta, file.path(output_dir, "report_metadata.json"),
                     pretty = TRUE, auto_unbox = TRUE)
rm(.run_uuid, .input_hash, .sinfo, .meta)
```

```{r dynamic-description, results='asis'}
# Dynamic description based on assay type
if (isTRUE(is_elisa)) {
  cat(tr("elisa_intro", lang, assay_config$analyte %||% "analyte"), "\n\n")
  cat("**", tr("elisa_method", lang), "**\n\n", sep = "")
} else {
  cat(tr("rba_intro", lang, assay_config$toxin_standard_label %||% "biotoxin"), "\n\n")
  cat("**", tr("rba_method", lang, assay_config$detection_method %||% "radioligand"), "**\n\n", sep = "")
}
```

```{r notes-section, results='asis'}
.has_notes <- nzchar(trimws(notes_text)) && notes_text != "No notes added."
if (.has_notes) {
  cat("## ", tr("analysis_notes", lang), "\n\n", sep = "")
  cat(notes_text, "\n\n")
}
```


```{r compact-body-class, results='asis', eval=is_compact}
if (is_html_out()) {
  cat('<script>document.addEventListener("DOMContentLoaded",function(){document.body.classList.add("compact");});</script>\n')
}
```

```{r model-fitting, include=FALSE}
# M1.2: fit_all_models() in reports/analysis_pipeline.R handles the
# modelling work (high-variability filtering, weight setup, H2 auto pick,
# LL.4 -> LL.3 -> interpolation fallback chain, EC20/EC80, prediction grid).
# The two closure helpers below stay in the chunk so they capture the
# chunk-local environment (where bb0_lower / bb0_upper / ec20 / ec80 now
# live after list2env) rather than the function's local scope, preserving
# the legacy lookup behaviour for later chunks that call them.
#
# H4 cache: render_reports() in report_pipeline.R renders the same template
# up to four times per run (compact + full × html + docx). All variants share
# the same fit, so we serialise the first run's result to output_dir and
# short-circuit subsequent renders. Pipeline clears the file at run start.
.fit_cache_path <- file.path(output_dir, ".fit_all_models_cache.rds")
if (file.exists(.fit_cache_path)) {
  .fit_res <- readRDS(.fit_cache_path)
} else {
  .fit_res <- fit_all_models(
    data_long       = data_long,
    response_var    = response_var,
    analysis_config = analysis_config,
    is_elisa        = is_elisa,
    lang            = lang,
    cv_limit        = cv_limit,
    STATS_CONFIG    = STATS_CONFIG
  )
  saveRDS(.fit_res, .fit_cache_path)
}
list2env(.fit_res, envir = environment())

# classify_range(): uses different reference axes depending on assay type.
#
# ELISA: classifies by %B/B0 response value. The kit manufacturer
#   (e.g., Cayman) defines the reliable quantification range in terms of
#   %B/B0 bounds (default 20-80%). The response axis is bounded [0, 100],
#   so %B/B0 is the natural reference.
#
# RBA: classifies by estimated concentration vs EC20/EC80 concentrations,
#   following the AOAC SMPR recommendation for radioligand binding assays.
#   The response axis (CPM) is unbounded, so concentration is the natural
#   reference.
#
# Both approaches identify samples falling outside the linear portion of the
# dose-response curve; they just use different axes to do so.
classify_range <- function(response_val, conc_val, is_elisa_assay) {
  if (isTRUE(is_elisa_assay)) {
    if (is.na(response_val)) return("Unknown")
    if (response_val >= bb0_lower && response_val <= bb0_upper) return("Within Range")
    return("Out of Range")
  } else {
    if (is.na(conc_val) || is.na(ec20) || is.na(ec80)) return("Unknown")
    if (conc_val >= ec20 && conc_val <= ec80) return("Within Range")
    return("Out of Range")
  }
}

# Standard curve bounds used by flag_range() to extend <LLOQ / >ULOQ to
# concentrations outside the range of fitted standards (extrapolated region).
.std_min_conc <- if (exists("standards_for_model") && nrow(standards_for_model) > 0)
  min(standards_for_model$concentration, na.rm = TRUE) else NA_real_
.std_max_conc <- if (exists("standards_for_model") && nrow(standards_for_model) > 0)
  max(standards_for_model$concentration, na.rm = TRUE) else NA_real_

# Helper function: get LLOQ/ULOQ flag for a sample.
# Both assay types flag concentrations outside the standard curve range as
# <LLOQ (below lowest standard) or >ULOQ (above highest standard), since
# such estimates are no more reliable than out-of-range ones.
flag_range <- function(response_val, conc_val, is_elisa_assay) {
  if (isTRUE(is_elisa_assay)) {
    if (is.na(response_val)) return("")
    if (response_val > bb0_upper) return("<LLOQ")
    if (response_val < bb0_lower) return(">ULOQ")
    # Standard-range coverage check (mirrors the RBA branch below)
    if (!is.na(conc_val)) {
      if (!is.na(.std_min_conc) && conc_val < .std_min_conc) return("<LLOQ")
      if (!is.na(.std_max_conc) && conc_val > .std_max_conc) return(">ULOQ")
    }
    return("")
  } else {
    if (is.na(conc_val) || is.na(ec20) || is.na(ec80)) return("")
    if (conc_val < ec20 || (!is.na(.std_min_conc) && conc_val < .std_min_conc)) return("<LLOQ")
    if (conc_val > ec80 || (!is.na(.std_max_conc) && conc_val > .std_max_conc)) return(">ULOQ")
    return("")
  }
}

# GUI refresh §6.2 / Q4(e): pre-compute LLOQ/ULOQ from
# determine_lloq_uloq() right after the fit so the KPI strip below
# can reference formal_lloq / formal_uloq. The lloq-uloq chunk
# further down still owns the rendering of the back-calculation
# table; it now reads .lloq_res from this chunk's environment
# instead of recomputing.
.lloq_res <- if (!drc_failed_completely && !is.null(model_fit)) {
  tryCatch(
    determine_lloq_uloq(
      standards_for_model = standards_for_model,
      model_fit           = model_fit,
      response_var        = response_var,
      is_elisa            = is_elisa
    ),
    error = function(e) NULL
  )
} else {
  NULL
}
formal_lloq <- if (!is.null(.lloq_res)) .lloq_res$formal_lloq else NA_real_
formal_uloq <- if (!is.null(.lloq_res)) .lloq_res$formal_uloq else NA_real_
```

```{r standard-backcalculation-compute, include=FALSE}
.recovery <- compute_standard_recovery(data_long       = data_long,
                                       model_fit       = model_fit,
                                       response_var    = response_var,
                                       is_elisa        = is_elisa)
standards_backcalc <- .recovery$standards_backcalc
recovery_summary   <- .recovery$recovery_summary
```

```{r sample-analysis, include=FALSE}
# M1.1: quantify_samples() in reports/analysis_pipeline.R does the per-well
# delta-method prediction, optional outlier detection (Shapiro->MAD/Dixon's Q/
# Grubbs), replicate-group layered uncertainty (HP5/HP6), within-range and
# LLOQ/ULOQ flagging, and (for ELISA + tissue_weights.json) pg/g tissue.
# classify_range and flag_range are passed in so the function calls them
# without re-creating the chunk-scoped closures.
#
# H4 cache: as for fit_all_models, all variants in a single render_reports()
# run share the same per-well quantification; serialise the first render's
# result and short-circuit the rest. Pipeline clears the file at run start.
.samp_cache_path <- file.path(output_dir, ".quantify_samples_cache.rds")
if (file.exists(.samp_cache_path)) {
  .samp_res <- readRDS(.samp_cache_path)
} else {
  .samp_res <- quantify_samples(
    data_long           = data_long,
    response_var        = response_var,
    model_fit           = model_fit,
    standards_for_model = standards_for_model,
    analysis_config     = analysis_config,
    STATS_CONFIG        = STATS_CONFIG,
    is_elisa            = is_elisa,
    output_dir          = output_dir,
    classify_range      = classify_range,
    flag_range          = flag_range
  )
  saveRDS(.samp_res, .samp_cache_path)
}
list2env(.samp_res, envir = environment())
```

```{r save-model-stats, include=FALSE}
# Save model statistics for multiwavelength executive summary
# standards_data: all unique defined concentrations (including high-variability ones
# excluded from fitting). Computed here so it's available in both compact and full mode
# (the standards-table display chunk is gated eval=!is_compact).
if (!exists("standards_data")) {
  standards_data <- data_long %>%
    dplyr::filter(SampleType == "Standard", !is.na(StandardConc)) %>%
    dplyr::arrange(desc(StandardConc)) %>%
    dplyr::distinct(StandardConc)
}
model_stats <- list(
  r_squared = R2,
  rmse = RMSE,
  ic50 = coef(model_fit)[4],
  hill_slope = coef(model_fit)[1],
  n_standards = nrow(standards_for_model),
  n_unique_concs = length(unique(standards_data$StandardConc)),
  weight_method = weight_desc
)

# Note: mean_sample_cv is computed later in save-sample-stats chunk
# after replicate_stats is available (BUG_008 fix)

# write_json_safe is defined in global.R (loaded before rendering)
write_json_safe(model_stats, file.path(output_dir, "model_stats.json"))

```

```{r save-sample-stats, include=FALSE}
# BUG_008 fix: Update model_stats with mean_sample_cv now that replicate_stats exists
if (exists("model_stats") && exists("replicate_stats") &&
    !is.null(replicate_stats) && nrow(replicate_stats) > 0) {
  model_stats$mean_sample_cv <- mean(replicate_stats$cv_percent, na.rm = TRUE)
} else if (exists("model_stats")) {
  model_stats$mean_sample_cv <- NA_real_
}
# Re-save model_stats with the updated mean_sample_cv
if (exists("model_stats") && exists("write_json_safe")) {
  write_json_safe(model_stats, file.path(output_dir, "model_stats.json"))
}
```

```{r report-header, results='asis'}
# Condensed metadata block (left) + KPI strip (right) in a single flex row.
# Replaces the former stacked pre-fit executive summary box and post-fit
# green-gradient overall-status box.
render_report_header(
  R2              = R2,
  RMSE            = RMSE,
  formal_lloq     = formal_lloq,
  formal_uloq     = formal_uloq,
  assay_config    = assay_config,
  is_elisa        = is_elisa,
  data_long       = data_long,
  lang            = lang,
  control_summary = if (exists("control_summary")) control_summary else NULL
)
```

```{r exec-summary-accuracy, results='asis'}
# M1.6: one-line "Mean recovery X% ± Y% across N levels" appended after
# the status box. render_overall_accuracy() returns invisibly when the
# standard-recovery summary was not produced (e.g. compact mode).
render_overall_accuracy(
  recovery_summary = if (exists("recovery_summary")) recovery_summary else NULL
)
```

```{r variability-info, results='asis'}
# M1.6: bullet list of standards excluded for high CV.
render_high_variability_info(high_var_standards = high_var_standards,
                             is_elisa           = is_elisa,
                             assay_config       = assay_config,
                             lang               = lang)
```

```{r qc-summary-card, results='asis'}
# Section A1 of the v2 refactor: merges the pre-analysis "Data Quality
# Overview" panel (well counts + exclusions) with the post-table "Quality
# Alert" emit (high-CV warning) into a single severity-coded card placed
# after sample-analysis runs so all signals are available. Uses the
# .exec-summary-card is-pass / is-warn classes defined in report_style.css.
.qc_total_wells    <- nrow(data_long)
.qc_std_wells      <- sum(data_long$SampleType == "Standard",                na.rm = TRUE)
.qc_sample_wells   <- sum(data_long$SampleType == "Sample",                  na.rm = TRUE)
.qc_control_wells  <- sum(data_long$SampleType %in% c("Blank", "NSB", "B0"), na.rm = TRUE)
.qc_excluded_wells <- sum(is.na(data_long$MeasurementValue) |
                          !is.finite(data_long$MeasurementValue))
.qc_outliers_n     <- if (exists("outlier_flags") && !is.null(outlier_flags))
                        nrow(outlier_flags) else 0L
.qc_high_cv_groups <- if (exists("replicate_stats") && !is.null(replicate_stats) &&
                          "cv_percent" %in% names(replicate_stats))
                        replicate_stats$Replicate[
                          !is.na(replicate_stats$cv_percent) &
                          replicate_stats$cv_percent > 30] else character(0)
.qc_high_cv_n      <- length(.qc_high_cv_groups)

.qc_severity <- if (.qc_excluded_wells == 0L && .qc_outliers_n == 0L &&
                    .qc_high_cv_n == 0L) "pass" else "warn"
.qc_title_key <- if (.qc_severity == "pass") "quality_pass" else "quality_alert"

if (is_html_out()) {
  cat(sprintf('<div class="exec-summary-card is-%s">\n', .qc_severity))
  cat(sprintf('  <strong>%s</strong>\n', tr(.qc_title_key, lang)))
  cat(sprintf('  <p>Total wells: %d &middot; Standards: %d &middot; Samples: %d &middot; Controls: %d</p>\n',
              .qc_total_wells, .qc_std_wells, .qc_sample_wells, .qc_control_wells))
  if (.qc_severity == "warn") {
    cat('  <ul>\n')
    if (.qc_excluded_wells > 0L) cat(sprintf('    <li>%d excluded wells (missing or non-finite)</li>\n',
                                              .qc_excluded_wells))
    if (.qc_outliers_n     > 0L) cat(sprintf('    <li>%d outliers flagged</li>\n',
                                              .qc_outliers_n))
    if (.qc_high_cv_n      > 0L) cat(sprintf('    <li>%s</li>\n',
                                              tr("high_cv_groups", lang,
                                                 paste(.qc_high_cv_groups, collapse = ", "))))
    cat('  </ul>\n')
    if (.qc_high_cv_n > 0L) {
      cat(sprintf('  <small>%s</small>\n', tr("check_preparation", lang)))
    }
  }
  cat('</div>\n')
} else {
  # DOCX / PDF: plain markdown fallback (no banner styling).
  cat("\n**", tr(.qc_title_key, lang), "**\n\n", sep = "")
  cat(sprintf("Total wells: %d | Standards: %d | Samples: %d | Controls: %d\n\n",
              .qc_total_wells, .qc_std_wells, .qc_sample_wells, .qc_control_wells))
  if (.qc_severity == "warn") {
    if (.qc_excluded_wells > 0L) cat(sprintf("- %d excluded wells (missing or non-finite)\n",
                                              .qc_excluded_wells))
    if (.qc_outliers_n     > 0L) cat(sprintf("- %d outliers flagged\n", .qc_outliers_n))
    if (.qc_high_cv_n      > 0L) cat("- ",
                                     tr("high_cv_groups", lang,
                                        paste(.qc_high_cv_groups, collapse = ", ")),
                                     "\n", sep = "")
    if (.qc_high_cv_n      > 0L) cat("- ", tr("check_preparation", lang), "\n", sep = "")
    cat("\n")
  }
}

rm(list = ls(pattern = "^\\.qc_"))
```

## `r tr("section_drc_samples", lang)`

```{r combined-drc, fig.width=12, fig.height=8, warning=FALSE, message=FALSE, fig.cap="Dose-response curve with sample concentration estimates overlaid on the fitted model"}
if (exists("sample_results") && !is.null(sample_results) && nrow(sample_results) > 0) {
  
  standards_plot_data <- data_long %>%
    dplyr::filter(!is.na(concentration), SampleType == "Standard") %>%
    # Task 3: pre-compute hover text for the standards layer so plotly's
    # tooltip="text" picks it up. Inline aes(text = paste0(...)) would also
    # work but a mutate keeps the response-column dynamic lookup
    # (.data[[response_var]]) outside the aesthetic mapping.
    dplyr::mutate(
      .response_for_hover = .data[[response_var]],
      .hover_text = paste0(
        "Standard: ", SampleID, "<br>",
        "Concentration: ",
        if (isTRUE(is_elisa)) {
          format(concentration, digits = 4, big.mark = ",")
        } else {
          scales::scientific(concentration, digits = 2)
        }, "<br>",
        labels$y_label, ": ", round(.response_for_hover, 2), "<br>",
        "Variability: ", high_variability
      )
    )

  unknown_plot_data <- sample_results %>%
    dplyr::filter(!is.na(concentration_diluted), is.finite(concentration_diluted)) %>%
    dplyr::mutate(
      falls_within_range = mapply(classify_range, response_value, estimated_concentration,
                                  MoreArgs = list(is_elisa_assay = is_elisa))
    )

  if (nrow(unknown_plot_data) > 0) {

    combined_drc <- ggplot() +
      geom_line(data = model_fits, aes(x = conc, y = p),
               color = "black", linewidth = 1.2) +
      geom_point(data = standards_plot_data,
                aes(x = concentration, y = .data[[response_var]],
                    color = high_variability,
                    text = .hover_text),
                size = 3, alpha = 0.7) +
      geom_point(data = unknown_plot_data,
                aes(x = concentration_diluted, y = response_value,
                    shape = falls_within_range,
                    text = paste0(
                      "Sample: ", SampleID, "<br>",
                      "Replicate: ", Replicate, "<br>",
                      "Dilution: 1:", DilutionFactor, "<br>",
                      "Actual Concentration: ",
                      if (isTRUE(is_elisa)) {
                        format(estimated_concentration, digits = 4, big.mark = ",")
                      } else {
                        scales::scientific(estimated_concentration, digits = 2)
                      }, "<br>",
                      "Diluted Concentration: ",
                      if (isTRUE(is_elisa)) {
                        format(concentration_diluted, digits = 4, big.mark = ",")
                      } else {
                        scales::scientific(concentration_diluted, digits = 2)
                      }, "<br>",
                      labels$y_label, ": ", round(response_value, 2), "<br>",
                      if ("concentration_pg_per_g" %in% names(unknown_plot_data) &&
                          !is.na(concentration_pg_per_g)) {
                        paste0("Conc. (pg/g tissue): ",
                               format(concentration_pg_per_g, digits = 4, nsmall = 1, big.mark = ","), "<br>")
                      } else { "" },
                      "Range Status: ", falls_within_range
                    )),
                size = 3, color = "#E69F00") +
      scale_shape_manual(name = "Sample Status",
                        values = c("Within Range" = 16, "Out of Range" = 17, "Unknown" = 18)) +
      # GUI refresh §6.5: keep variability cue on standards (still
      # useful) but adopt the Okabe-Ito hexes verbatim.
      scale_color_manual(name = tr("plot_std_variability", lang),
                        values = c("Normal Variability" = "#0072b2",
                                  "High Variability"   = "#d55e00"),
                        labels = c("Normal Variability" = tr("variability_normal", lang),
                                   "High Variability"   = tr("variability_high", lang))) +
      # GUI refresh §6.5: theme_minimal() with the design's gridline
      # color and a hairline panel border. Replaces the previous
      # theme_classic() + theme() chain with a single declarative
      # block.
      theme_minimal() +
      theme(
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "#e3e6ea", linewidth = 0.4),
        panel.border     = element_rect(color = "#e3e6ea", fill = NA, linewidth = 0.5),
        legend.position  = "top",
        legend.box       = "vertical",
        axis.text.x      = element_text(angle = 45, hjust = 1)
      ) +
      labs(x = labels$x_label, y = labels$y_label)
    
    if (isTRUE(is_elisa)) {
      combined_drc <- combined_drc + 
        scale_x_log10(
          labels = function(x) format(x, digits = 3, big.mark = ","),
          breaks = scales::trans_breaks("log10", function(x) 10^x, n = 6)
        )
    } else {
      combined_drc <- combined_drc + 
        scale_x_log10(
          limits = c(1e-12, 1e-5),
          breaks = 10^seq(-12, -5, by = 1),
          labels = scales::scientific_format(digits = 2)
        )
    }
    
    # Render plot - interactive for HTML, static for DOCX/PDF.
    # ggplotly names traces as "(level, N)" when multiple discrete aesthetics
    # interact. The fix must run on the raw plotly object inside render_plot()
    # before it is wrapped in tagList — pass it as plotly_postprocess.
    std_lbl   <- tr("plot_std_variability", lang)
    known_std <- c(tr("variability_normal", lang), tr("variability_high", lang))
    .fix_drc_legend <- function(px) {
      px$x$data <- lapply(px$x$data, function(trace) {
        nm    <- trace$name %||% ""
        clean <- sub("^\\((.+),\\s*\\d+\\)$", "\\1", nm)
        trace$name <- clean
        if (clean %in% known_std) {
          trace$legendgroup      <- "std_var"
          trace$legendgrouptitle <- list(text = std_lbl)
        } else if (clean %in% c("Within Range", "Out of Range", "Unknown")) {
          trace$legendgroup      <- "smp_status"
          trace$legendgrouptitle <- list(text = "Sample Status")
        }
        trace
      })
      px
    }
    p <- render_plot(combined_drc, tooltip = "text",
                     plotly_postprocess = .fix_drc_legend)

    if (is_html_out()) {
      htmltools::tagList(
        p,
        htmltools::tags$p(class = "bs-plot-caption",
                          htmltools::tags$em(tr("drc_legend_hint", lang)))
      )
    } else {
      p
    }
  }
}
```

```{r parallelism-assessment, results='asis', eval=!is_compact}
# Task C (HP7): Optional parallelism / relative potency assessment.
# Heading is emitted inside the chunk so it only appears when n_curves >= 2.
tryCatch({
  if (exists("model_fit") && !is.null(model_fit) && exists("data_long") && exists("response_var")) {
    parallelism_result <- assess_parallelism(data_long, model_fit, response_var)

    if (isTRUE(parallelism_result$applicable)) {
      cat("## ", tr("parallelism_title", lang), "\n\n", sep = "")
      cat("### Test Results\n\n")

      if (!is.na(parallelism_result$slope_test$p_value)) {
        if (isTRUE(parallelism_result$slope_test$parallel)) {
          cat(sprintf("**Slope comparison:** Slopes are not significantly different (t = %.3f, p = %.4f). Curves appear parallel.\n\n",
                      parallelism_result$slope_test$statistic,
                      parallelism_result$slope_test$p_value))
        } else {
          cat(sprintf("**Slope comparison:** Slopes differ significantly (t = %.3f, p = %.4f). Curves are **not** parallel.\n\n",
                      parallelism_result$slope_test$statistic,
                      parallelism_result$slope_test$p_value))
        }
      }

      if (!is.na(parallelism_result$potency_ratio$estimate)) {
        cat(sprintf("**Relative potency (EC50 ratio):** %.4f [95%% CI: %.4f \u2013 %.4f]\n\n",
                    parallelism_result$potency_ratio$estimate,
                    parallelism_result$potency_ratio$ci_lower,
                    parallelism_result$potency_ratio$ci_upper))
      }

      cat(sprintf("*%d curve groups were compared.*\n\n", parallelism_result$curves_found))
    }
    # When not applicable: emit nothing \u2014 section is omitted entirely
  }
}, error = function(e) {
  cat(sprintf("*Parallelism assessment could not be completed: %s*\n\n", e$message))
})
```

## `r tr("sample_results", lang)`

```{r sample-results-table, results='asis', warning=FALSE}
# .dagger_groups: accumulate CI-truncated group names for Exclusion Audit
.dagger_groups <- character(0)

tryCatch({

  if (exists("replicate_stats") && !is.null(replicate_stats) && nrow(replicate_stats) > 0) {

    cv_limit <- QC_THRESHOLDS$cv_limit %||% 30

    # ---- tissue weights ----
    has_tissue_weights <- FALSE
    if (exists("sample_results") && !is.null(sample_results) &&
        "TissueWeight_mg" %in% names(sample_results)) {
      has_tissue_weights <- isTRUE(tryCatch(
        any(!is.na(sample_results$TissueWeight_mg) & sample_results$TissueWeight_mg > 0,
            na.rm = TRUE),
        error = function(e) FALSE))
    }

    conc_col_name <- if (isTRUE(is_elisa)) "Mean Conc. (pg/mL)" else "Mean Conc. (mol/L)"

    # ---- numeric flag columns (kept for row-class callbacks) ----
    rs <- replicate_stats %>%
      dplyr::mutate(
        .cv_num   = as.numeric(cv_percent),
        .oor_flag = if ("within_range" %in% names(.)) within_range == "Out of Range" else FALSE,
        .cv_flag  = !is.na(.cv_num) & .cv_num > cv_limit
      )

    # ---- accumulate dagger groups ----
    if ("ci_truncated" %in% names(rs)) {
      .dagger_groups <<- rs$Replicate[!is.na(rs$ci_truncated) & rs$ci_truncated]
    }

    # ---- build display data frame ----
    base_cols <- c("Replicate", "sampleID", "n_replicates",
                   "mean_conc_formatted", "ci_formatted", "cv_formatted")

    # Status: two stacked pills \u2014 Cal. Range (top) + Quant. Range (bottom)
    .status_pills <- function(qs, rf) {
      qs_str <- if (!is.na(qs) && nzchar(qs)) as.character(qs) else "Unknown"
      rf_str <- if (!is.na(rf)) as.character(rf) else ""
      .pill  <- function(label, role)
        sprintf('<span class="bs-status-pill is-%s">%s</span>', role, label)

      if (is_html_out()) {
        # Cal. Range \u2014 calibration curve coverage
        cal_html <- if (qs_str %in% c("Interpolated", "Within Range") && rf_str == "") {
          .pill(tr("cal_in_range", lang), "pass")
        } else if (rf_str == "<LLOQ") {
          .pill(tr("cal_below", lang), "warn")
        } else if (rf_str == ">ULOQ") {
          .pill(tr("cal_above", lang), "warn")
        } else if (qs_str == "Not estimable") {
          .pill("Not\u00a0estimable", "fail")
        } else {
          .pill(qs_str, "warn")
        }

        # Quant. Range \u2014 EC20-EC80 / %B/B0 20-80% quantifiable window
        quant_html <- if (qs_str == "Not estimable") {
          .pill("N/A", "fail")
        } else if (rf_str == "<LLOQ") {
          .pill(tr("quant_below_lloq", lang), "warn")
        } else if (rf_str == ">ULOQ") {
          .pill(tr("quant_above_uloq", lang), "warn")
        } else {
          .pill(tr("quant_in_range", lang), "pass")
        }

        paste0(cal_html, '<br>', quant_html)
      } else {
        # DOCX/PDF plain text
        rf_part <- if (nzchar(rf_str)) rf_str else qs_str
        qr_part <- if (qs_str == "Not estimable") "N/A"
                   else if (rf_str == "<LLOQ")     "< LLOQ"
                   else if (rf_str == ">ULOQ")     "> ULOQ"
                   else                            "Within quantifiable"
        paste0(rf_part, " / ", qr_part)
      }
    }

    .ci_cell <- function(ci_fmt, cv_num) {
      if (!is_html_out()) return(ci_fmt)
      if (!is.na(cv_num) && cv_num > cv_limit) {
        paste0('<span class="ci-uninformative" title="CI uninformative \u2014 CV > ',
               cv_limit, '%">', ci_fmt, '</span>')
      } else {
        ci_fmt
      }
    }

    rs <- rs %>%
      dplyr::mutate(
        .qs    = if ("quant_status" %in% names(.)) as.character(quant_status) else "Unknown",
        .rf    = if ("range_flag"   %in% names(.)) ifelse(is.na(range_flag), "", as.character(range_flag)) else "",
        Status = mapply(.status_pills, .qs, .rf, SIMPLIFY = TRUE),
        ci_display = mapply(.ci_cell, ci_formatted, .cv_num, SIMPLIFY = TRUE)
      )

    results_display <- rs %>%
      dplyr::select(Replicate, sampleID, n_replicates,
                    mean_conc_formatted, ci_display, cv_formatted,
                    Status, .oor_flag, .cv_flag)

    names(results_display) <- c(
      tr("col_replicate_group", lang), tr("col_sample_ids", lang),
      tr("col_n", lang), conc_col_name,
      tr("col_ci", lang), tr("col_cv", lang),
      "Status",
      ".oor", ".cv"
    )

    # ---- tissue columns ----
    if (isTRUE(is_elisa) && isTRUE(has_tissue_weights)) {
      tissue_results <- tryCatch({
        sample_results %>%
          dplyr::filter(!is.na(TissueWeight_mg) & TissueWeight_mg > 0) %>%
          dplyr::group_by(Replicate) %>%
          dplyr::summarise(
            tissue_weight    = dplyr::first(TissueWeight_mg),
            mean_tissue_conc = mean(concentration_pg_per_g, na.rm = TRUE),
            .groups = "drop"
          ) %>%
          dplyr::mutate(
            tissue_weight_formatted = format(tissue_weight,    digits = 4, nsmall = 1),
            tissue_conc_formatted   = format(mean_tissue_conc, digits = 4, nsmall = 2, big.mark = ",")
          )
      }, error = function(e) NULL)

      if (!is.null(tissue_results) && nrow(tissue_results) > 0) {
        rg_col <- tr("col_replicate_group", lang)
        results_display <- results_display %>%
          dplyr::left_join(
            tissue_results %>% dplyr::select(Replicate, tissue_weight_formatted, tissue_conc_formatted),
            by = setNames("Replicate", rg_col)
          )
        names(results_display)[names(results_display) == "tissue_weight_formatted"] <- "Tissue Mass [mg]"
        names(results_display)[names(results_display) == "tissue_conc_formatted"]   <- "Conc. (pg/g tissue)"
      }
    }

    caption_text <- tr("sample_results_caption", lang, toupper(assay_config$assay_type))
    if (isTRUE(is_elisa) && isTRUE(has_tissue_weights)) {
      caption_text <- paste(caption_text, tr("with_tissue", lang))
    }

    # ---- compact mode: only flagged rows ----
    if (isTRUE(is_compact)) {
      results_display <- results_display %>%
        dplyr::filter(.data[[".oor"]] | .data[[".cv"]])
    }

    # ---- flag column indices (0-based for DataTables JS) ----
    oor_col_idx <- which(names(results_display) == ".oor") - 1L
    cv_col_idx  <- which(names(results_display) == ".cv")  - 1L

    if (isTRUE(.has_DT) && is_html_out() && !isTRUE(is_compact)) {
      # Convert logical flag cols to numeric for JS comparison
      results_display[[".oor"]] <- as.integer(results_display[[".oor"]])
      results_display[[".cv"]]  <- as.integer(results_display[[".cv"]])

      dt_obj <- DT::datatable(
        results_display,
        caption    = caption_text,
        escape     = FALSE,
        rownames   = FALSE,
        options    = list(
          pageLength = 25,
          dom        = "ftip",
          columnDefs = list(
            list(targets = c(oor_col_idx, cv_col_idx), visible = FALSE)
          ),
          rowCallback = DT::JS("function(row, data) { }")
        )
      )
      htmltools::tagList(dt_obj)
    } else {
      # DOCX / PDF or DT unavailable: plain kable (drop hidden flag columns)
      display_for_kable <- results_display %>%
        dplyr::select(-dplyr::any_of(c(".oor", ".cv")))
      render_table(display_for_kable, caption = caption_text)
    }

    # ---- save CSV files ----
    if (exists("sample_results_clean") && !is.null(sample_results_clean)) {
      write.csv(sample_results_clean,
                file.path(output_dir, "unknown_results.csv"), row.names = FALSE)
    }
    if (exists("replicate_summary") && !is.null(replicate_summary)) {
      write.csv(replicate_summary,
                file.path(output_dir, "unknown_results_summary.csv"), row.names = FALSE)
    }

    if (is_html_out()) {
      .fmt_code <- function(s) gsub("`([^`]+)`", "<code>\\1</code>", s)
      cat(paste0(
        '\n<p><strong>', tr("output_files_created", lang), '</strong></p>\n',
        '<ul class="bs-output-files">\n',
        '<li>', .fmt_code(tr("individual_results", lang)), '</li>\n',
        '<li>', .fmt_code(tr("summary_results",    lang)), '</li>\n',
        '</ul>\n\n'
      ))
    } else {
      cat("\n**", tr("output_files_created", lang), "**\n\n", sep = "")
      cat("- ", tr("individual_results", lang), "\n", sep = "")
      cat("- ", tr("summary_results", lang), "\n\n", sep = "")
    }

  } else {
    cat("*", tr("no_samples_quantified", lang), "*\n\n", sep = "")

    if (exists("samples_to_analyze") && !is.null(samples_to_analyze) &&
        nrow(samples_to_analyze) > 0) {
      cat("**Possible reasons:**\n")
      cat("- Response values outside the standard curve range\n")
      cat("- Model fitting issues\n")
      if (isTRUE(is_elisa)) {
        cat("- Invalid %B/B0 normalization\n- Control well hierarchy violations\n")
      }
      cat("- Mathematical prediction errors\n")
    } else {
      cat("**No sample data found for analysis.**\n")
      cat("Note: Only wells marked as 'Sample' are quantified.\n")
      cat("Blanks, NSB, TA, and Standards are excluded from quantification.\n")
    }
  }

}, error = function(e) {
  cat("\n**Error in sample results table:**\n")
  cat("Error message:", e$message, "\n")
  cat("This section will be skipped.\n\n")
})
```

```{r outlier-results, results='asis'}
if (isTRUE(analysis_config$enable_outlier_detection)) {
  section_start(tr("outlier_title", lang))

  # Phase 1.5: clarify that flagged != removed from the raw table
  cat("> ", tr("outlier_flagged_not_removed_note", lang), "\n\n", sep = "")

  if (exists("outlier_flags") && nrow(outlier_flags) > 0) {
    cat(tr("outlier_desc", lang, analysis_config$outlier_min_n %||% 3), "\n\n")

    n_outliers <- nrow(outlier_flags)
    n_groups <- length(unique(outlier_flags$Replicate))
    cat(tr("outlier_found", lang, n_outliers, n_groups), "\n\n")

    outlier_display <- outlier_flags %>%
      dplyr::select(Well, Replicate, method)
    names(outlier_display) <- c("Well", "Replicate Group", "Method")

    render_table(outlier_display, caption = tr("outlier_title", lang))
  } else {
    cat(tr("outlier_none", lang), "\n\n")
  }

  # Method log (Shapiro-Wilk pre-test details)
  if (exists("outlier_method_log") && nrow(outlier_method_log) > 0) {
    cat("\n**Outlier detection method per replicate group (Shapiro-Wilk pre-test enabled):**\n\n")
    for (i in seq_len(nrow(outlier_method_log))) {
      row <- outlier_method_log[i, ]
      cat(sprintf("- %s: %s (Shapiro-Wilk p = %.4f)\n", row$Replicate, row$method_used, row$shapiro_p))
    }
    cat("\n")
  }

  section_end()
}
```

```{r ci-method-note, results='asis'}
if (isTRUE(analysis_config$ci_method == "bootstrap")) {
  cat(sprintf("\n*%s*\n\n", tr("ci_bootstrap_note", lang)))
  cat(sprintf("*Replicate-level CIs computed via percentile bootstrap (%d resamples).*\n\n",
              STATS_CONFIG$bootstrap_iterations))
} else {
  cat(sprintf("\n*%s*\n\n", tr("ci_tdist_note", lang)))
  cat("*Replicate-level CIs reflect measurement variability using Student's t-distribution (df = n-1).*\n\n")
}
cat(sprintf("*%s*\n\n", tr("ci_delta_method_note", lang)))
```

```{r ci-asymmetric-footnote, results='asis'}
cat(sprintf("\n*%s*\n\n", tr("ci_asymmetric_footnote", lang)))
```

```{r table-legend, results='asis'}
render_table_legend()
```

`r section_open(tr("detailed_summary", lang), default_open = TRUE)`

```{r summary-results-table, results='asis', warning=FALSE}
if (exists("replicate_summary") && !is.null(replicate_summary) && nrow(replicate_summary) > 0) {

  .raw_cv <- if ("cv_percent" %in% names(replicate_summary)) as.numeric(replicate_summary$cv_percent) else rep(NA_real_, nrow(replicate_summary))

  # B4: pre-extract flag/status columns before the mutate to avoid data-masking issues
  .rf_col <- if ("range_flag"   %in% names(replicate_summary)) replicate_summary[["range_flag"]]   else rep(NA_character_, nrow(replicate_summary))
  .qs_col <- if ("quant_status" %in% names(replicate_summary)) replicate_summary[["quant_status"]] else rep(NA_character_, nrow(replicate_summary))

  # Section B1: SD and SE are different statistics (SD = within-group spread
  # of wells; SE = SD/sqrt(n) = uncertainty of the group mean) \u2014 render them
  # as two separate columns rather than fusing as "X +/- Y", which misleads
  # readers into thinking SE is the uncertainty *of* SD. A one-line legend
  # above the table explains both terms.
  summary_display <- replicate_summary %>%
    dplyr::mutate(
      mean_concentration = if (isTRUE(is_elisa)) {
        format(as.numeric(mean_concentration), digits = 4, nsmall = 1, big.mark = ",")
      } else {
        format(as.numeric(mean_concentration), scientific = TRUE, digits = 3)
      },
      sd_value = if (isTRUE(is_elisa)) {
        ifelse(is.na(std_dev),   "\u2014",
               format(as.numeric(std_dev),   digits = 3, nsmall = 1))
      } else {
        ifelse(is.na(std_dev),   "\u2014",
               format(as.numeric(std_dev),   scientific = TRUE, digits = 2))
      },
      se_value = if (isTRUE(is_elisa)) {
        ifelse(is.na(std_error), "\u2014",
               format(as.numeric(std_error), digits = 3, nsmall = 1))
      } else {
        ifelse(is.na(std_error), "\u2014",
               format(as.numeric(std_error), scientific = TRUE, digits = 2))
      },
      ci_95 = if (isTRUE(is_elisa)) {
        paste0("[", format(as.numeric(ci_lower_95), digits = 3, nsmall = 1), " \u2013 ",
               format(as.numeric(ci_upper_95),      digits = 3, nsmall = 1), "]")
      } else {
        paste0("[", format(as.numeric(ci_lower_95), scientific = TRUE, digits = 2), " \u2013 ",
               format(as.numeric(ci_upper_95),      scientific = TRUE, digits = 2), "]")
      },
      cv_percent = ifelse(is.na(cv_percent), "\u2014",
                          sprintf("%.1f%%", as.numeric(cv_percent))),
      status_combined = dplyr::case_when(
        !is.na(.rf_col) & .rf_col != ""                                        ~ .rf_col,
        !is.na(.qs_col) & !(.qs_col %in% c("Quantified", "Within Range", "")) ~ .qs_col,
        TRUE                                                                    ~ "OK"
      ),
      mean_bb0 = if ("mean_response" %in% names(.)) {
        ifelse(is.na(mean_response), "\u2014",
               paste0(round(as.numeric(mean_response), 2), "%"))
      } else {
        NULL
      }
    ) %>%
    # Drop CI raw bounds, raw SD / SE (replaced by formatted sd_value / se_value),
    # and internal cols.
    dplyr::select(-ci_lower_95, -ci_upper_95, -std_dev, -std_error,
                  -any_of(c("within_range", "range_flag", "quant_status", "mean_response",
                            "sample_type",
                            "ci_lower_model",      "ci_upper_model",
                            "ci_lower_replicate",  "ci_upper_replicate",
                            "ci_lower_raw",        "ci_upper_raw",
                            "ci_truncated"))) %>%
    dplyr::relocate(sd_value,        .after  = mean_concentration) %>%
    dplyr::relocate(se_value,        .after  = sd_value) %>%
    dplyr::relocate(ci_95,           .before = cv_percent) %>%
    dplyr::relocate(status_combined, .after  = cv_percent) %>%
    { if ("mean_bb0" %in% names(.)) dplyr::relocate(., mean_bb0, .before = status_combined) else . }

  # Item 18 (HTML only): merge Sample IDs + n as a small grey line inside the
  # Group cell; drop those columns so they don't appear as separate columns.
  # NOTE: analysis_pipeline.R renames Replicate\u2192replicate_group and
  #       sampleID\u2192sample_ids; use those actual column names here.
  if (is_html_out()) {
    .has_n   <- "n_replicates" %in% names(summary_display)
    .has_sid <- "sample_ids"   %in% names(summary_display)
    # Pre-extract outside mutate to avoid data-masking lookup issues.
    .sids    <- if (.has_sid) summary_display[["sample_ids"]]
                else          character(nrow(summary_display))
    summary_display <- summary_display %>%
      dplyr::mutate(
        replicate_group = if (.has_n) {
          paste0(replicate_group,
                 '<br><span class="bs-replicate-meta">',
                 .sids,
                 ' \u00b7 n=', n_replicates, '</span>')
        } else {
          paste0(replicate_group,
                 '<br><span class="bs-replicate-meta">',
                 .sids,
                 '</span>')
        }
      ) %>%
      dplyr::select(-any_of(c("sample_ids", "n_replicates")))
    rm(.has_n, .has_sid, .sids)

    # B4: convert status columns to two stacked bs-status-pill HTML spans.
    # .status_pills() is defined in the sample-results-table chunk (which runs
    # first); fall back to a single pill if somehow it is not in scope.
    if (exists(".status_pills", mode = "function")) {
      summary_display <- summary_display %>%
        dplyr::mutate(
          status_combined = mapply(.status_pills, .qs_col, .rf_col, SIMPLIFY = TRUE)
        )
    } else {
      summary_display <- summary_display %>%
        dplyr::mutate(
          status_combined = dplyr::case_when(
            status_combined == ">ULOQ" ~ '<span class="bs-status-pill is-warn">&gt;Cal.&nbsp;high</span>',
            status_combined == "<LLOQ" ~ '<span class="bs-status-pill is-warn">&lt;Cal.&nbsp;low</span>',
            status_combined == "OK"    ~ '<span class="bs-status-pill is-pass">OK</span>',
            TRUE                       ~ paste0('<span class="bs-status-pill is-warn">', status_combined, '</span>')
          )
        )
    }
  } else {
    # DOCX/PDF: drop n; keep sample_ids as its own column (no HTML tags)
    summary_display <- summary_display %>%
      dplyr::select(-any_of("n_replicates"))
  }

  # Build column names dynamically based on output format and available columns
  conc_unit <- if (isTRUE(is_elisa)) assay_config$units %||% "pg/mL" else "mol/L"

  if (is_html_out()) {
    base_names <- c("Group",
                    paste0("Mean (", conc_unit, ")"),
                    paste0("SD (", conc_unit, ")"), "SE",
                    "95% CI", "CV [%]")
  } else {
    base_names <- c("Group", "Sample IDs",
                    paste0("Mean (", conc_unit, ")"),
                    paste0("SD (", conc_unit, ")"), "SE",
                    "95% CI", "CV [%]")
  }

  if ("mean_bb0"        %in% names(summary_display)) base_names <- c(base_names, "%B/B0")
  if ("status_combined" %in% names(summary_display)) base_names <- c(base_names, "Status")

  # Tissue concentration columns if present
  if ("TissueWeight_mg" %in% names(summary_display)) {
    summary_display <- summary_display %>%
      dplyr::mutate(
        TissueWeight_mg = ifelse(
          is.na(TissueWeight_mg), "\u2014",
          format(TissueWeight_mg, digits = 4, nsmall = 1)),
        concentration_pg_per_g = ifelse(
          is.na(concentration_pg_per_g), "\u2014",
          format(concentration_pg_per_g, digits = 4, nsmall = 1, big.mark = ","))
      )
    base_names <- c(base_names, "Tissue Mass [mg]", "Conc. (pg/g)")
  }

  summary_display <- summary_display %>%
    dplyr::select(-any_of("total_amount_pg"))

  # C4: wrap ci_95 in .ci-uninformative span when CV exceeds threshold
  if (is_html_out()) {
    summary_display$ci_95 <- mapply(function(ci_fmt, cv_num) {
      if (!is.na(cv_num) && cv_num > 30)
        paste0('<span class="ci-uninformative" title="CI uninformative — CV > 30%">', ci_fmt, '</span>')
      else ci_fmt
    }, summary_display$ci_95, .raw_cv, SIMPLIFY = TRUE)
  }

  names(summary_display) <- base_names[1:ncol(summary_display)]

  .tbl <- render_table(summary_display,
                       caption = tr("detailed_caption", lang),
                       escape  = !is_html_out())

  if (is_html_out()) {
    # HTML branch: emit a one-line legend explaining SD vs SE, a client-side
    # filter input above the table, and wrap the table in a fixed-height
    # scroll container (sticky thead via CSS). The JS is guarded so
    # re-rendering doesn't redefine the function.
    cat('<aside class="bs-table-legend"><small>',
        tr("sd_se_legend", lang),
        '</small></aside>\n', sep = "")
    cat('<div class="bs-table-filter">',
        '<input type="text" class="bs-filter-input"',
        ' placeholder="', tr("filter_placeholder", lang), '"',
        ' aria-label="', tr("filter_aria_label", lang), '"',
        ' oninput="bsFilterTable(this)">',
        '</div>\n', sep = "")
    cat('<script>\n',
        'if (typeof window.bsFilterTable !== "function") {\n',
        '  window.bsFilterTable = function(input) {\n',
        '    var q = (input.value || "").toLowerCase();\n',
        '    var wrap = input.closest(".bs-table-filter");\n',
        '    if (!wrap) return;\n',
        '    var scroll = wrap.nextElementSibling;\n',
        '    if (!scroll) return;\n',
        '    var tbl = scroll.querySelector("table");\n',
        '    if (!tbl || !tbl.tBodies[0]) return;\n',
        '    var rows = tbl.tBodies[0].rows;\n',
        '    for (var i = 0; i < rows.length; i++) {\n',
        '      var text = rows[i].textContent.toLowerCase();\n',
        '      rows[i].style.display = (q === "" || text.indexOf(q) >= 0) ? "" : "none";\n',
        '    }\n',
        '  };\n',
        '}\n',
        '</script>\n', sep = "")
    cat('<script>\n',
        'if (typeof window.bsSortTable !== "function") {\n',
        '  window.bsSortTable = function(th) {\n',
        '    var tbl = th.closest("table");\n',
        '    if (!tbl || !tbl.tBodies[0]) return;\n',
        '    var col = Array.prototype.indexOf.call(th.parentNode.children, th);\n',
        '    var asc = th.getAttribute("data-sort-dir") !== "asc";\n',
        '    var rows = Array.from(tbl.tBodies[0].rows);\n',
        '    rows.sort(function(a, b) {\n',
        '      var va = (a.cells[col] || {}).textContent.trim();\n',
        '      var vb = (b.cells[col] || {}).textContent.trim();\n',
        '      var na = parseFloat(va.replace(/[^0-9.\\-eE+]/g, ""));\n',
        '      var nb = parseFloat(vb.replace(/[^0-9.\\-eE+]/g, ""));\n',
        '      if (!isNaN(na) && !isNaN(nb)) return asc ? na - nb : nb - na;\n',
        '      return asc ? va.localeCompare(vb) : vb.localeCompare(va);\n',
        '    });\n',
        '    rows.forEach(function(r) { tbl.tBodies[0].appendChild(r); });\n',
        '    th.parentNode.querySelectorAll("th").forEach(function(t) { t.removeAttribute("data-sort-dir"); });\n',
        '    th.setAttribute("data-sort-dir", asc ? "asc" : "desc");\n',
        '  };\n',
        '  document.addEventListener("DOMContentLoaded", function() {\n',
        '    document.querySelectorAll(".bs-table-scroll thead th").forEach(function(th) {\n',
        '      th.addEventListener("click", function() { window.bsSortTable(th); });\n',
        '    });\n',
        '  });\n',
        '}\n',
        '</script>\n', sep = "")
    cat('<div class="bs-table-scroll">\n')
    cat(as.character(.tbl))
    cat('\n</div>\n')
    invisible(NULL)
  } else {
    # DOCX/PDF: italic markdown legend above the plain table.
    cat(sprintf("*%s*\n\n", tr("sd_se_legend", lang)))
    .tbl
  }
}
```

`r section_close()`

```{r section-methods-open, results='asis', eval=!is_compact}
if (is_html_out()) {
  cat('<details>\n<summary><strong>', tr("methods_title", lang), '</strong></summary>\n\n')
} else {
  cat('## ', tr("methods_title", lang), '\n\n')
}
```

`r if (!is_compact) section_open(tr("std_concentrations", lang)) else ""`

```{r standards-table, eval=!is_compact}
# Extract and display standard concentrations
standards_data <- data_long %>%
  filter(SampleType == "Standard", !is.na(StandardConc)) %>%
  arrange(desc(StandardConc)) %>%
  distinct(StandardConc)

# Format concentrations based on assay type
if (isTRUE(is_elisa)) {
  # ELISA: Use regular number formatting (no scientific notation)
  standard_table <- data.frame(
    StandardNumber = seq_len(nrow(standards_data)),
    Concentration = format(standards_data$StandardConc, digits = 4, nsmall = 1, big.mark = ",")
  )
  unit_label <- paste0("Concentration (", assay_config$units %||% "pg/mL", ")")
} else {
  # RBA: Use scientific notation  
  standard_table <- data.frame(
    StandardNumber = seq_len(nrow(standards_data)),
    Concentration = format(standards_data$StandardConc, scientific = TRUE, digits = 3)
  )
  unit_label <- "Concentration (mol/L)"
}

render_table(
  standard_table, 
  caption = tr("std_concentrations_table", lang),
  col_names = c(tr("col_standards", lang), unit_label)
)
```

`r if (!is_compact) section_close() else ""`

```{r methods-section, results='asis', eval=!is_compact}
# Item 10: expanded Methods section covering model, weighting, LLOQ/ULOQ
# criteria, CI method, outlier rule, and software. Replaces the former single
# sentence. render_methods_section() lives in reports/report_sections.R.
render_methods_section(
  analysis_config = analysis_config,
  assay_config    = assay_config,
  is_elisa        = is_elisa,
  lang            = lang
)
```

```{r lloq-uloq, results='asis', warning=FALSE, eval=!is_compact}
# Formal LLOQ/ULOQ determination by back-calculated standard accuracy.
# M1.3: compute lives in determine_lloq_uloq() in reports/analysis_pipeline.R;
# this chunk handles localisation, formatting, and rendering.
# GUI refresh §6.2 / Q4(e): determine_lloq_uloq() is now called once
# in the model-fitting chunk above (so the KPI strip can read its
# results); this chunk reuses .lloq_res from that call. Falls back
# to recomputing if the upstream chunk failed.
tryCatch({
  section_start(tr("lloq_uloq_title", lang))
  cat(tr("lloq_uloq_desc", lang), "\n\n")

  if (!exists(".lloq_res") || is.null(.lloq_res)) {
    .lloq_res <- determine_lloq_uloq(
      standards_for_model = standards_for_model,
      model_fit           = model_fit,
      response_var        = response_var,
      is_elisa            = is_elisa
    )
  }
  backcalc_results <- .lloq_res$backcalc_results
  backcalc_summary <- .lloq_res$backcalc_summary
  passing_concs    <- .lloq_res$passing_concs

  # Add the localised accuracy label to the summary so the rendering code
  # below stays as-is.
  if (nrow(backcalc_summary) > 0) {
    backcalc_summary$accuracy <- ifelse(backcalc_summary$passes,
                                        tr("qc_green", lang),
                                        tr("qc_red", lang))
  }

  if (nrow(backcalc_summary) > 0) {
    # Display back-calculation table
    display_bc <- backcalc_summary %>%
      dplyr::select(nominal_conc, mean_backcalc, mean_recovery, cv_backcalc, accuracy)

    if (isTRUE(is_elisa)) {
      display_bc$nominal_conc <- format(display_bc$nominal_conc, digits = 4, big.mark = ",")
      display_bc$mean_backcalc <- format(display_bc$mean_backcalc, digits = 4, big.mark = ",")
    } else {
      display_bc$nominal_conc <- format(display_bc$nominal_conc, scientific = TRUE, digits = 3)
      display_bc$mean_backcalc <- format(display_bc$mean_backcalc, scientific = TRUE, digits = 3)
    }
    display_bc$mean_recovery <- paste0(round(as.numeric(backcalc_summary$mean_recovery), 1), "%")
    display_bc$cv_backcalc <- ifelse(is.na(backcalc_summary$cv_backcalc), "\u2014",
                                      paste0(round(backcalc_summary$cv_backcalc, 1), "%"))

    units_str <- if (isTRUE(is_elisa)) assay_config$units %||% "pg/mL" else "mol/L"
    names(display_bc) <- c(
      sprintf(tr("col_nominal", lang), units_str),
      sprintf(tr("col_backcalc", lang), units_str),
      tr("col_recovery", lang),
      "CV%",
      tr("col_accuracy", lang)
    )
    render_table(display_bc, caption = tr("backcalc_title", lang))

    # LLOQ/ULOQ already computed by determine_lloq_uloq(). Promote to the
    # outer chunk env (formal_lloq / formal_uloq are read by a later chunk
    # via exists() so we set them either way).
    if (length(passing_concs) >= 2) {
      formal_lloq <- .lloq_res$formal_lloq
      formal_uloq <- .lloq_res$formal_uloq
      cat(sprintf("\n\n**%s:** %s %s\n\n", tr("lloq_label", lang),
                  if (isTRUE(is_elisa)) format(formal_lloq, digits = 4, big.mark = ",") else format(formal_lloq, scientific = TRUE, digits = 3),
                  units_str))
      cat(sprintf("**%s:** %s %s\n\n", tr("uloq_label", lang),
                  if (isTRUE(is_elisa)) format(formal_uloq, digits = 4, big.mark = ",") else format(formal_uloq, scientific = TRUE, digits = 3),
                  units_str))
    } else {
      cat(tr("lloq_uloq_none", lang), "\n\n")
    }
  } else {
    cat(tr("lloq_uloq_none", lang), "\n\n")
  }

  section_end()
}, error = function(e) {
  cat(sprintf("\n\n*LLOQ/ULOQ determination failed: %s*\n\n", e$message))
})
```

```{r section-methods-close, results='asis', eval=!is_compact}
if (is_html_out()) cat('</details>\n\n')
```

```{r section-curve-diagnostics-open, results='asis', eval=!is_compact}
if (is_html_out()) {
  cat('<details>\n<summary><strong>', tr("section_curve_diagnostics", lang), '</strong></summary>\n\n')
} else {
  cat('## ', tr("section_curve_diagnostics", lang), '\n\n')
}
```

`r if (!is_compact) section_open(tr("four_pl_coefficients", lang), default_open = TRUE) else ""`

```{r model-stats, warning=FALSE, eval=!is_compact}
# Extract and display model coefficients with correct parameter order
tryCatch({
  
  coefficients <- summary(model_fit)$coefficients
  
  # Check if coefficients have expected structure
  if (is.matrix(coefficients) && nrow(coefficients) >= 4) {
    
    # DRC package LL.4() parameter order: b (Hill slope), c (Lower asymptote), d (Upper asymptote), e (IC50)
    actual_param_names <- c(tr("hill_slope", lang), tr("bottom", lang), tr("top", lang), tr("ic50", lang))
    
    # Ensure we don't exceed available rows
    n_params <- min(length(actual_param_names), nrow(coefficients))
    
    # Create results data frame with correct parameter mapping
    pval_col <- if ("Pr(>|t|)" %in% colnames(coefficients)) {
      coefficients[1:n_params, "Pr(>|t|)"]
    } else {
      rep(NA_real_, n_params)
    }
    drc_results <- data.frame(
      Parameter = actual_param_names[1:n_params],
      Value = coefficients[1:n_params, "Estimate"],
      StdError = if ("Std. Error" %in% colnames(coefficients)) {
        coefficients[1:n_params, "Std. Error"]
      } else {
        rep(NA_real_, n_params)
      },
      stringsAsFactors = FALSE
    )
    col_nm <- c(tr("col_parameter", lang), tr("col_estimate", lang), tr("col_std_error", lang))

    # Drop p-value column when all values are NA (e.g. 3-parameter fits)
    if (!all(is.na(pval_col))) {
      drc_results$PValue <- pval_col
      col_nm <- c(col_nm, tr("col_pvalue", lang))
    }

    render_table(
      drc_results,
      caption = tr("four_pl_coefficients", lang),
      col_names = col_nm
    )
    
  } else {
    cat("Model coefficients structure unexpected. Available coefficients:\n")
    print(coefficients)
  }
  
}, error = function(e) {
  cat("Error extracting model coefficients:", e$message, "\n")
  cat("Model summary:\n")
  print(summary(model_fit))
})
```

```{r model-fit-stats-display, results='asis', eval=!is_compact}
if (is_html_out()) cat('</details>\n\n')
cat(sprintf("\n**%s**\n\n", tr("model_fit_stats", lang)))
cat(sprintf("- **R\u00b2** = %.3f\n", R2))
cat(sprintf("- **RMSE** = %.3f %s\n", RMSE, labels$y_unit))
n_uniq  <- length(unique(standards_for_model$StandardConc))
n_reps  <- round(nrow(standards_for_model) / max(n_uniq, 1))
n_total <- nrow(standards_for_model)
n_def   <- length(unique(standards_data$StandardConc))
cat(sprintf("- **%s** = %d %s \u00d7 %d replicates = %d wells (of %d defined levels)\n\n",
            tr("standards_used", lang), n_uniq, tr("std_levels", lang), n_reps, n_total, n_def))
```

```{r weighting-suitability, results='asis', warning=FALSE, eval=!is_compact}
# H2: document the auto-weighting decision (if taken) at the top of the
# weighting-suitability section, before the full diagnostic block below.
if (!is.null(auto_weighting_result)) {
  weight_labels <- list(none = tr("weight_unweighted", lang), inv_y = "1/Y", inv_y2 = "1/Y²")
  chosen_label <- weight_labels[[auto_weighting_result$chosen_weight]] %||%
                  auto_weighting_result$chosen_weight
  if (isTRUE(auto_weighting_result$fallback)) {
    cat("\n\n", tr("weight_auto_fallback", lang), "\n\n", sep = "")
  } else {
    cat("\n\n", tr("weight_auto_decision", lang,
                    chosen_label,
                    auto_weighting_result$test_name %||% "Brown-Forsythe",
                    auto_weighting_result$statistic %||% NA_real_,
                    auto_weighting_result$p_value %||% NA_real_),
        "\n\n", sep = "")
  }
  if (isTRUE(auto_weighting_result$degenerate)) {
    cat("\n\n> **Warning:** The heteroscedasticity test was numerically degenerate",
        "(at least one replicate group has zero within-group variance — F → ∞).",
        "The auto-weight selection used the variance ratio as a fallback.",
        "Inspect your standards for all-identical replicates at any concentration level.",
        "\n\n", sep = " ")
  }
}

# Formal heteroscedasticity diagnostic (assess_heteroscedasticity)
tryCatch({
  # Determine which model to assess: prefer the unweighted model, otherwise
  # use the first available model as a reference
  hetero_model <- NULL
  hetero_label <- NULL

  if ("none" %in% names(all_models) && !is.null(all_models[["none"]]$model)) {
    hetero_model <- all_models[["none"]]$model
    hetero_label <- "Unweighted (LL.4/LL.3)"
  } else if (!drc_failed_completely) {
    # No unweighted model was fitted; use primary model as proxy
    hetero_model <- all_models[[primary_key]]$model
    hetero_label <- all_models[[primary_key]]$label
  }

  if (!is.null(hetero_model)) {
    hetero_result <- assess_heteroscedasticity(
      model = hetero_model,
      data_long = standards_for_model,
      response_var = response_var
    )

    cat(sprintf("\n\n### %s\n\n", tr("weighting_suitability_title", lang)))

    cat(sprintf("**Test:** %s\n\n", hetero_result$test_name %||% "N/A"))

    if (!isTRUE(hetero_result$degenerate) && !is.na(hetero_result$statistic)) {
      cat(sprintf("**Statistic:** %s", fmt_smart(hetero_result$statistic, 3)))
      if (!is.na(hetero_result$p_value)) {
        cat(sprintf(" \u2003 **p-value:** %s", fmt_smart(hetero_result$p_value, 4)))
      }
      cat("\n\n")
    }

    if (!is.na(hetero_result$variance_ratio)) {
      cat(sprintf("**Variance ratio** (max/min across concentration levels): %.1f\n\n",
                  hetero_result$variance_ratio))
    }

    cat(sprintf("**Interpretation:** %s\n\n", hetero_result$interpretation))
    cat(sprintf("**Recommendation:** %s\n\n", hetero_result$recommendation))

    cat("*Note: This diagnostic is advisory only and does not change the selected model.*\n\n")
  }
}, error = function(e) {
  cat(sprintf("\n\n*Weighting suitability assessment could not be performed: %s*\n\n", e$message))
})
```

```{r weight-comparison, results='asis', fig.width=12, fig.height=7, warning=FALSE, message=FALSE, eval=!is_compact}
# Show comparison of multiple weighting methods if the user selected more than one
if (multi_weight_mode && length(all_models) > 1) {
  section_start(tr("weight_comparison_title", lang))
  cat(tr("weight_comparison_desc", lang), "\n\n")

  # --- Comparison table: R², RMSE, AIC, parameters (+ stability if available) ---
  comp_rows <- lapply(names(all_models), function(k) {
    m <- all_models[[k]]
    coefs <- m$coefs  # b (Hill), c (Bottom), d (Top), e (IC50)
    if (is.null(coefs)) {
      row <- data.frame(
        Method = m$label,
        R2 = NA, RMSE = NA, AIC = NA,
        Hill = NA, Bottom = NA, Top = NA, IC50 = NA,
        stringsAsFactors = FALSE
      )
    } else {
      row <- data.frame(
        Method = m$label,
        R2 = round(m$R2, 5),
        RMSE = round(m$RMSE, 3),
        AIC = if (!is.null(m$AIC) && !is.na(m$AIC)) round(m$AIC, 1) else NA,
        Hill = round(abs(coefs[1]), 4),
        Bottom = round(coefs[2], 2),
        Top = round(coefs[3], 2),
        IC50 = if (isTRUE(is_elisa)) format(round(coefs[4], 2), big.mark = ",") else format(coefs[4], scientific = TRUE, digits = 3),
        stringsAsFactors = FALSE
      )
    }
    # Append stability grade if already assessed
    if (!is.null(m$stability)) {
      grade_icon <- function(g) switch(g,
        good = "\U0001F7E2", acceptable = "\U0001F7E1",
        unstable = "\U0001F534", failed = "\u274C", "\u2753")
      row$Stability <- paste(grade_icon(m$stability$stability_grade),
                              tools::toTitleCase(m$stability$stability_grade))
    }
    row
  })
  comp_df <- do.call(rbind, comp_rows)
  col_names <- c(tr("qc_metric", lang), "R\u00B2", "RMSE", "AIC",
                  tr("hill_slope", lang), tr("bottom", lang), tr("top", lang), tr("ic50", lang))
  if ("Stability" %in% names(comp_df)) {
    col_names <- c(col_names, "Stability")
  }
  names(comp_df) <- col_names

  render_table(comp_df, caption = tr("weight_comparison_title", lang))

  # --- AIC recommendation note ---
  aic_vals <- sapply(all_models, function(m) if (!is.null(m$AIC)) m$AIC else NA_real_)
  if (any(!is.na(aic_vals))) {
    best_key <- names(which.min(aic_vals))
    best_label <- all_models[[best_key]]$label
    best_aic <- min(aic_vals, na.rm = TRUE)
    cat(sprintf("\n\n**AIC recommends: %s** (lowest AIC = %.1f). ", best_label, best_aic))
    cat(tr("weight_comparison_primary", lang, all_models[[primary_key]]$label), "\n\n")
  }

  # --- Overlay DRC curves ---
  curve_colors <- c("none" = "#000000", "inv_y" = "#0072B2", "inv_y2" = "#D55E00")
  curve_data_list <- list()
  for (k in names(all_models)) {
    if (is.null(all_models[[k]]$model)) next
    pred_df <- data.frame(conc = conc_range)
    pred_df$p <- predict(all_models[[k]]$model, newdata = pred_df)
    pred_df$method <- all_models[[k]]$label
    curve_data_list[[k]] <- pred_df
  }
  if (length(curve_data_list) > 0) {
  curve_data <- do.call(rbind, curve_data_list)
  plotted_keys <- names(curve_data_list)

  standards_plot <- data_long %>%
    dplyr::filter(!is.na(concentration), SampleType == "Standard")

  cmp_plot <- ggplot() +
    geom_point(data = standards_plot,
               aes(x = concentration, y = .data[[response_var]]),
               size = 3, alpha = 0.6, color = "grey40") +
    geom_line(data = curve_data,
              aes(x = conc, y = p, color = method),
              linewidth = 1.1) +
    scale_color_manual(
      name = tr("col_weighting_label", lang),
      values = setNames(
        curve_colors[plotted_keys],
        sapply(all_models[plotted_keys], function(m) m$label)
      )
    ) +
    theme_classic() +
    theme(legend.position = "top",
          axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x = labels$x_label, y = labels$y_label)

  if (isTRUE(is_elisa)) {
    cmp_plot <- cmp_plot +
      scale_x_log10(
        labels = function(x) format(x, digits = 3, big.mark = ","),
        breaks = scales::trans_breaks("log10", function(x) 10^x, n = 6)
      )
  } else {
    cmp_plot <- cmp_plot +
      scale_x_log10(
        limits = c(1e-12, 1e-5),
        breaks = 10^seq(-12, -5, by = 1),
        labels = scales::scientific_format(digits = 2)
      )
  }

  } # end if (length(curve_data_list) > 0)
}
```

```{r weight-comparison-plot, fig.width=12, fig.height=7, warning=FALSE, message=FALSE, eval=!is_compact}
# Separate chunk so render_plot() return value is visible to knitr
if (exists("cmp_plot") && multi_weight_mode && length(all_models) > 1) {
  render_plot(cmp_plot)
}
```

```{r weight-comparison-close, results='asis', eval=!is_compact}
if (multi_weight_mode && length(all_models) > 1) {
  cat(sprintf("\n\n*%s*\n\n",
      tr("weight_comparison_primary", lang, all_models[[primary_key]]$label)))
  section_end()
}

# Simple heteroscedasticity diagnostic for unweighted model
if ("none" %in% names(all_models) && !is.null(all_models[["none"]]$model)) {
  unw_fit <- all_models[["none"]]$model
  unw_resid <- residuals(unw_fit)
  conc_levels <- standards_for_model$StandardConc
  var_by_level <- tapply(unw_resid^2, conc_levels, mean, na.rm = TRUE)
  var_ratio <- max(var_by_level, na.rm = TRUE) / max(min(var_by_level[var_by_level > 0], na.rm = TRUE), 1e-10)
  if (var_ratio > 3) {
    cat(sprintf("\n\n> **Note:** Residual variance ratio across concentration levels = %.1f (>3), suggesting heteroscedasticity. Weighted regression (1/Y or 1/Y²) is recommended.\n\n", var_ratio))
  }
}
```

```{r model-stability-assessment, results='asis', warning=FALSE, eval=!is_compact}
# Convergence and stability diagnostics for all fitted models
tryCatch({
  if (!drc_failed_completely && length(all_models) > 0) {

    # Assess stability for each fitted model
    stability_results <- lapply(names(all_models), function(k) {
      assess_model_stability(all_models[[k]], all_models[[k]]$label)
    })
    names(stability_results) <- names(all_models)

    emit_heading(tr("model_stability_title", lang), 4)

    # Build summary table
    grade_icon <- function(g) {
      switch(g,
        good       = "\U0001F7E2 Good",
        acceptable = "\U0001F7E1 Acceptable",
        unstable   = "\U0001F534 Unstable",
        failed     = "\u274C Failed",
        "\u2753 Unknown"
      )
    }

    stability_df <- data.frame(
      Model = sapply(stability_results, function(s) s$model_name),
      Method = sapply(stability_results, function(s) s$fit_method %||% "N/A"),
      Converged = sapply(stability_results, function(s) if (s$converged) tr("yes_label", lang) else tr("no_label", lang)),
      SE_Available = sapply(stability_results, function(s) if (s$se_available) tr("yes_label", lang) else tr("no_label", lang)),
      Boundary = sapply(stability_results, function(s) if (s$boundary_flag) tr("yes_label", lang) else tr("no_label", lang)),
      Grade = sapply(stability_results, function(s) grade_icon(s$stability_grade)),
      stringsAsFactors = FALSE
    )
    names(stability_df) <- c(tr("col_model", lang), tr("col_fit_method", lang),
                              tr("col_converged", lang), tr("col_se_available", lang),
                              tr("col_boundary_hit", lang), tr("col_stability_grade", lang))

    render_table(stability_df, caption = tr("model_stability_title", lang))

    # Print detailed warnings / suspicious parameters per model
    for (k in names(stability_results)) {
      sr <- stability_results[[k]]
      has_details <- length(sr$suspicious_params) > 0 || length(sr$warnings) > 0

      if (has_details) {
        cat(sprintf("\n**%s:**\n\n", sr$model_name))
        for (w in sr$warnings) {
          cat(sprintf("- %s\n", w))
        }
        for (sp in sr$suspicious_params) {
          cat(sprintf("- %s\n", sp))
        }
        cat("\n")
      }
    }

    # Store stability info for downstream use (e.g., model comparison table)
    # Attach to all_models entries
    for (k in names(stability_results)) {
      all_models[[k]]$stability <- stability_results[[k]]
    }
  }
}, error = function(e) {
  cat(sprintf("\n\n*Model stability assessment could not be performed: %s*\n\n", e$message))
})
```

```{r standard-backcalculation-render, results='asis', warning=FALSE, eval=!is_compact}
section_start(tr("std_backcalc_title", lang))
# Format for display
if (isTRUE(is_elisa)) {
  recovery_display <- recovery_summary %>%
    dplyr::mutate(
      `Nominal Conc.` = format(nominal, digits = 4, nsmall = 1, big.mark = ","),
      `Back-Calculated` = format(mean_backcalc, digits = 4, nsmall = 1, big.mark = ","),
      `SD` = ifelse(is.na(sd_backcalc), "—", format(sd_backcalc, digits = 3, nsmall = 1)),
      `Recovery (%)` = ifelse(is.na(mean_recovery), "—", format(mean_recovery, digits = 4, nsmall = 1))
    ) %>%
    dplyr::select(`Nominal Conc.`, n, `Back-Calculated`, SD, `Recovery (%)`)
} else {
  recovery_display <- recovery_summary %>%
    dplyr::mutate(
      `Nominal Conc.` = format(nominal, scientific = TRUE, digits = 3),
      `Back-Calculated` = format(mean_backcalc, scientific = TRUE, digits = 3),
      `SD` = ifelse(is.na(sd_backcalc), "—", format(sd_backcalc, scientific = TRUE, digits = 2)),
      `Recovery (%)` = ifelse(is.na(mean_recovery), "—", format(mean_recovery, digits = 4, nsmall = 1))
    ) %>%
    dplyr::select(`Nominal Conc.`, n, `Back-Calculated`, SD, `Recovery (%)`)
}

conc_unit <- if (isTRUE(is_elisa)) paste0(assay_config$units %||% "pg/mL") else "mol/L"
names(recovery_display) <- c(sprintf(tr("col_nominal", lang), conc_unit), tr("col_n", lang),
                              sprintf(tr("col_backcalc", lang), conc_unit),
                              paste0("SD (", conc_unit, ")"),
                              tr("col_recovery", lang))

render_table(recovery_display,
            caption = tr("std_backcalc_caption", lang))

# Report overall recovery statistics
mean_overall_recovery <- mean(recovery_summary$mean_recovery, na.rm = TRUE)
cat(sprintf("\n**%s**\n", tr("overall_recovery", lang, mean_overall_recovery)))
if (mean_overall_recovery >= 80 && mean_overall_recovery <= 120) {
  cat(tr("recovery_acceptable", lang), "\n\n")
} else {
  cat(tr("recovery_outside", lang), "\n\n")
}
section_end()
```


```{r qc-traffic-light, results='asis', warning=FALSE, eval=!is_compact}
# M1.6: traffic-light QC summary card (R^2 / Hill / max CV / mean recovery).
# Compute + presentation are inseparable here so the helper does both;
# wrap whole block in a tryCatch that prints a one-line "could not be
# generated" message instead of crashing the report.
render_qc_traffic_light_section(model_fit          = model_fit,
                                standards_for_model = standards_for_model,
                                response_var       = response_var,
                                replicate_stats    = if (exists("replicate_stats")) replicate_stats else NULL,
                                R2                 = R2,
                                weight_desc        = weight_desc,
                                assay_config       = assay_config,
                                lang               = lang)
```

```{r section-curve-diagnostics-close, results='asis', eval=!is_compact}
if (is_html_out()) cat('</details>\n\n')
```

```{r section-plate-open, results='asis', eval=!is_compact}
if (is_html_out()) {
  cat('<details>\n<summary><strong>', tr("section_plate_diagnostics", lang), '</strong></summary>\n\n')
} else {
  cat('## ', tr("section_plate_diagnostics", lang), '\n\n')
}
```

```{r sample-boxplot, fig.width=10, fig.height=7, results='asis', warning=FALSE, message=FALSE, eval=!is_compact}
if (exists("sample_results") && !is.null(sample_results) && nrow(sample_results) > 0) {
  
  sample_plot_data <- sample_results %>%
    dplyr::filter(!is.na(estimated_concentration), is.finite(estimated_concentration)) %>%
    dplyr::mutate(
      Replicate_wrapped = stringr::str_wrap(Replicate, width = 15),
      within_range = mapply(classify_range, response_value, estimated_concentration,
                            MoreArgs = list(is_elisa_assay = is_elisa)),
      range_flag = mapply(flag_range, response_value, estimated_concentration,
                          MoreArgs = list(is_elisa_assay = is_elisa))
    )
  
  if (nrow(sample_plot_data) > 0) {

    # Dynamic font sizing based on number of groups
    n_groups <- length(unique(sample_plot_data$Replicate_wrapped))
    x_font_size <- if (n_groups > 20) 7 else if (n_groups > 10) 9 else 11

    # Create boxplot
    sample_box <- ggplot(sample_plot_data,
                        aes(x = factor(Replicate_wrapped),
                            y = estimated_concentration)) +
      geom_boxplot(outlier.shape = NA) +
      geom_jitter(aes(color = within_range,
                     text = paste0(
                       "Replicate: ", Replicate, "<br>",
                       "Sample ID: ", SampleID, "<br>",
                       "Concentration: ",
                       if (isTRUE(is_elisa)) {
                         format(estimated_concentration, digits = 4, big.mark = ",")
                       } else {
                         scales::scientific(estimated_concentration, digits = 2)
                       }, "<br>",
                       "Range Status: ", within_range
                     )),
                 width = 0.15, size = 2) +
      xlab(tr("col_replicate_group", lang)) +
      ylab(if (isTRUE(is_elisa)) {
        paste0("Estimated Concentration (", assay_config$units %||% "pg/mL", ")")
      } else {
        "Estimated Concentration (mol/L)"
      }) +
      theme_classic() +
      theme(
        axis.text.x = element_text(angle = 45, hjust = 1, size = x_font_size),
        axis.title.x = element_text(margin = margin(t = 12)),
        axis.title.y = element_text(margin = margin(r = 12)),
        legend.position = "top",
        panel.spacing.y = unit(1.5, "lines"),
        strip.text = element_blank()
      ) +
      labs(color = "Within linear range?")

    # Add log scale if appropriate
    conc_range <- max(sample_plot_data$estimated_concentration, na.rm = TRUE) /
                  min(sample_plot_data$estimated_concentration, na.rm = TRUE)

    if (!isTRUE(is_elisa) || conc_range > 100) {
      sample_box <- sample_box +
        scale_y_log10(labels = if (isTRUE(is_elisa)) {
          function(x) format(x, big.mark = ",")
        } else {
          scales::scientific_format(digits = 2)
        })
    }

    n_unique_replicates <- length(unique(sample_plot_data$Replicate))
    .box_subtitle <- sprintf("Replicate groups, n=%d each", round(nrow(sample_plot_data) / n_unique_replicates))
    sample_box <- sample_box +
      labs(subtitle = .box_subtitle)

    # Section wrapper + description — open by default (details open)
    if (is_html_out()) {
      cat('<details open>\n<summary><strong>',
          tr("sample_variability", lang),
          '</strong></summary>\n\n', sep = "")
    } else {
      section_start("Sample variability visualization")
    }
    cat(sprintf("*%s*\n\n", tr("sample_variability_desc", lang)))
  }
}

```{r sample-boxplot-render, fig.width=10, fig.height=5, warning=FALSE, message=FALSE, eval=!is_compact}
# Separate chunk so render_plot() return value is visible to knitr
if (exists("sample_box") && !is.null(sample_box)) {
  fig_h_px <- 480L
  render_plot(sample_box, tooltip = "text", height = fig_h_px,
              plotly_layout = list(margin = list(b = 120, l = 80)))
}
```

```{r sample-boxplot-close, results='asis', eval=!is_compact}
if (exists("sample_box") && !is.null(sample_box)) {
  if (is_html_out()) cat("</details>\n\n") else section_end()
}
```

```{r plate-heatmap, fig.width=10, fig.height=6, results='asis', warning=FALSE, message=FALSE, eval=!is_compact}
tryCatch({
  cat(tr("heatmap_desc", lang), "\n\n")

  # Enrich with sample-group labels for the hover tooltip
  .group_lookup <- data_long %>%
    dplyr::distinct(Row, Column, SampleType, SampleID) %>%
    dplyr::mutate(Column = as.numeric(gsub("[^0-9]", "", as.character(Column))))

  heatmap_data <- data_long %>%
    dplyr::filter(!is.na(MeasurementValue)) %>%
    dplyr::mutate(
      Row    = factor(Row, levels = LETTERS[1:8]),
      Column = as.numeric(gsub("[^0-9]", "", as.character(Column)))
    ) %>%
    dplyr::filter(!is.na(Row) & !is.na(Column)) %>%
    dplyr::left_join(.group_lookup, by = c("Row", "Column", "SampleType", "SampleID")) %>%
    dplyr::mutate(
      well_id   = paste0(as.character(Row), Column),
      bb0_label = if (isTRUE(is_elisa) && "calculated_bb0" %in% names(.)) {
        paste0(round(calculated_bb0, 1), "% B/B0")
      } else { "" },
      hover_txt = paste0(
        "Well: ",   well_id,          "<br>",
        "Raw OD: ", round(MeasurementValue, 4), "<br>",
        ifelse(nzchar(bb0_label), paste0(bb0_label, "<br>"), ""),
        "Group: ",  SampleType,       "<br>",
        "ID: ",     SampleID
      )
    )

  if (nrow(heatmap_data) > 0) {
    if (is_html_out()) {
      # Interactive plotly heatmap for HTML
      hm_plotly <- plot_ly(
        data        = heatmap_data,
        x           = ~factor(Column, levels = 1:12),
        y           = ~factor(as.character(Row), levels = rev(LETTERS[1:8])),
        z           = ~MeasurementValue,
        text        = ~hover_txt,
        hoverinfo   = "text",
        type        = "heatmap",
        colorscale  = list(c(0, "#2166AC"), c(0.5, "#F7F7F7"), c(1, "#B2182B")),
        colorbar    = list(title = "Value"),
        showscale   = TRUE
      ) %>%
        layout(
          xaxis = list(title = "Column", tickvals = 1:12),
          yaxis = list(title = "Row"),
          margin = list(l = 50, r = 20, t = 40, b = 50)
        )
      htmltools::tagList(hm_plotly)
    } else {
      # Static ggplot for DOCX/PDF
      heatmap_plot <- ggplot(heatmap_data,
                             aes(x = factor(Column), y = Row, fill = MeasurementValue)) +
        geom_tile(color = "white", linewidth = 0.5) +
        geom_text(aes(label = sprintf("%.3f", MeasurementValue)), size = 3.5, color = "black") +
        scale_fill_gradient2(low = "#2166AC", mid = "#F7F7F7", high = "#B2182B",
                             midpoint = median(heatmap_data$MeasurementValue, na.rm = TRUE),
                             name = "Value") +
        scale_y_discrete(limits = rev(levels(factor(heatmap_data$Row, levels = LETTERS[1:8])))) +
        labs(x = "Column", y = "Row", title = tr("heatmap_title", lang)) +
        theme_minimal() +
        theme(panel.grid = element_blank(),
              axis.text  = element_text(size = 10, face = "bold"),
              plot.title = element_text(hjust = 0.5))
      print(heatmap_plot)
    }
  }
}, error = function(e) {
  cat("\n\n*Heatmap unavailable for this run.*\n\n")
})
```

```{r plate-positional-qc, results='asis', eval=!is_compact}
section_start("Plate Positional Quality")
positional <- if (exists("data_long")) assess_plate_positional(data_long) else NULL
if (!is.null(positional)) {
  if (positional$flagged_rows > 0 || positional$flagged_cols > 0) {
    cat(sprintf("**Warning:** %d row(s) and %d column(s) deviate >2 SD from plate mean, suggesting possible edge effects.\n\n",
                positional$flagged_rows, positional$flagged_cols))
  } else {
    cat("No significant row or column bias detected (all within 2 SD of plate mean).\n\n")
  }
  row_display <- positional$row_stats
  row_display$Mean <- round(row_display$Mean, 3)
  row_display$SD   <- round(row_display$SD,   3)
  # Empty-flag cells: show a pass pill for HTML, blank for DOCX/PDF
  if (is_html_out())
    row_display$Flag <- ifelse(row_display$Flag == "",
                               '<span class="bs-status-pill is-pass">OK</span>',
                               row_display$Flag)
  names(row_display) <- c("Row", "Mean Response", "SD", "Flag")
  render_table(row_display, caption = "Response by Row", escape = !is_html_out())
}
section_end()
```

```{r section-plate-close, results='asis', eval=!is_compact}
if (is_html_out()) cat('</details>\n\n')
```

```{r section-interpretation-open, results='asis'}
if (is_html_out()) {
  cat('<details open>\n<summary><strong>Interpretation &amp; Recommendations</strong></summary>\n\n')
} else {
  cat('## Interpretation & Recommendations\n\n')
}
```

```{r report-interpretation, results='asis'}
tryCatch({
  qc_issues <- c()

  # Use assay-specific QC profile for thresholds (M7)
  interp_qc_profile <- get_qc_profile(assay_config$assay_type %||% "rba")

  r2_val <- if (exists("model_stats") && !is.null(model_stats)) model_stats$r_squared %||% NA_real_ else NA_real_
  if (!is.na(r2_val) && is.finite(r2_val) && r2_val < interp_qc_profile$r2_threshold) {
    qc_issues <- c(qc_issues, sprintf("Low R\u00b2 (%.3f < %.2f threshold)", r2_val, interp_qc_profile$r2_threshold))
  }

  hill_val <- if (exists("model_stats") && !is.null(model_stats)) model_stats$hill_slope %||% NA_real_ else NA_real_
  if (!is.na(hill_val) && is.finite(hill_val)) {
    abs_hill <- abs(hill_val)
    abs_range <- sort(abs(interp_qc_profile$hill_slope_range))
    if (abs_hill < abs_range[1] || abs_hill > abs_range[2]) {
      qc_issues <- c(qc_issues, sprintf("Unusual Hill slope (%.2f; expected |slope| in %.1f to %.1f)",
                                          hill_val, abs_range[1], abs_range[2]))
    }
  }

  if (exists("replicate_stats") && !is.null(replicate_stats) && "cv_percent" %in% names(replicate_stats)) {
    max_cv <- max(replicate_stats$cv_percent, na.rm = TRUE)
    if (is.finite(max_cv) && max_cv > interp_qc_profile$cv_limit) {
      qc_issues <- c(qc_issues, sprintf("High sample CV (max %.1f%% > %d%% limit)", max_cv, interp_qc_profile$cv_limit))
    }
  }

  if (length(qc_issues) == 0) {
    cat("All quality criteria met. Results are suitable for reporting.\n\n")
  } else {
    cat(sprintf("**Quality warnings detected:** %s. Review flagged items before use.\n\n",
                paste(qc_issues, collapse = ", ")))
  }

  # M6: Quantification status counts
  if (exists("sample_results") && !is.null(sample_results) && nrow(sample_results) > 0 &&
      "quantification_status" %in% names(sample_results)) {
    status_counts <- table(sample_results$quantification_status)
    cat("**Quantification status breakdown (per well):**\n\n")
    for (s in names(status_counts)) {
      cat(sprintf("- %s: %d\n", s, status_counts[[s]]))
    }
    cat("\n")
  }

  cat(sprintf("\n*QC profile: %s (R\u00b2 >= %.2f, CV <= %d%%, Hill slope [%.1f, %.1f], Recovery [%d%%, %d%%])*\n\n",
              toupper(assay_config$assay_type %||% "rba"),
              interp_qc_profile$r2_threshold,
              interp_qc_profile$cv_limit,
              interp_qc_profile$hill_slope_range[1],
              interp_qc_profile$hill_slope_range[2],
              interp_qc_profile$recovery_range[1],
              interp_qc_profile$recovery_range[2]))
}, error = function(e) {
  cat(sprintf("\n\n*Interpretation section failed: %s*\n\n", e$message))
})
```

```{r tissue-normalization-traceability, results='asis', eval=!is_compact}
# M8: Tissue Normalization Traceability (ELISA + tissue weights only)
render_tissue_normalization_section(
  is_elisa          = isTRUE(is_elisa),
  tissue_weights    = if (exists("tissue_weights")) tissue_weights else NULL,
  processing_config = if (exists("processing_config")) processing_config else list(),
  replicate_stats   = if (exists("replicate_stats")) replicate_stats else NULL,
  lang              = lang
)
```

```{r section-interpretation-close, results='asis'}
if (is_html_out()) cat('</details>\n\n')
```

```{r section-exclusion-open, results='asis', eval=!is_compact}
if (is_html_out()) {
  cat('<details>\n<summary><strong>', tr("exclusion_audit_caption", lang), '</strong></summary>\n\n')
} else {
  cat('## ', tr("exclusion_audit_caption", lang), '\n\n')
}
```

```{r exclusion-audit, results='asis', eval=!is_compact}
# M5: Formal Exclusion Audit Table
render_exclusion_audit_section(
  data_long          = if (exists("data_long")) data_long else NULL,
  outlier_flags      = if (exists("outlier_flags")) outlier_flags else NULL,
  high_var_standards = if (exists("high_var_standards")) high_var_standards else NULL,
  lang               = lang
)

# Dagger list: CI lower bounds truncated to 0 — deferred from sample-results-table
if (exists(".dagger_groups") && length(.dagger_groups) > 0) {
  if (is_html_out()) {
    cat(sprintf(
      "\n<p><small>† CI lower bound truncated to 0 (negative original) for: %s</small></p>\n\n",
      paste(.dagger_groups, collapse = ", ")
    ))
  } else {
    cat(sprintf(
      "\n*† CI lower bound truncated to 0 (negative original) for: %s*\n\n",
      paste(.dagger_groups, collapse = ", ")
    ))
  }
}
```

```{r section-exclusion-close, results='asis', eval=!is_compact}
if (is_html_out()) cat('</details>\n\n')
```

```{r section-abbreviations-open, results='asis', eval=!is_compact}
if (is_html_out()) {
  cat('<details>\n<summary><strong>Abbreviations</strong></summary>\n\n')
} else {
  cat('## Abbreviations\n\n')
}
```

```{r abbreviations, results='asis', eval=!is_compact}
# Item 11: collapsed Abbreviations block above the report footer.
render_abbreviations_block()
```

```{r section-abbreviations-close, results='asis', eval=!is_compact}
if (is_html_out()) cat('</details>\n\n')
```

---

**`r tr("report_generated", lang)`** `r Sys.Date()`
**`r tr("contact", lang)`** kr.moeller@iaea.org
**`r tr("feedback", lang)`** [`r tr("online_form", lang)`](https://forms.office.com/e/q8eqJfp4QM)

*`r tr("automated_analysis", lang)`*
